---
title: "NHP Glucose, Body Weight, and Body Comp"
author: "Eric Shiroma, John Nam"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
---

## Study Description 

- This is an investigation into the longitudinal trajectories of glucose, body weight, and body composition in NHP (non-human primate colonies). These results will be used to compare similar trajectories in mice and humans.

***

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}

library (survival)
library (broom)
library (ggplot2)
library (ggthemes)
library (extrafont)
library (arsenal)
library (psych)
library (survey)
library (lubridate)
library (kableExtra)
library (gridExtra)
library (ggpubr)
library (grid)
library (tinytex)
library (haven)
library (tidyverse)
library (finalfit)
library (survival)
library (survminer)
library (BSDA)
library (readxl)
library (mgcv)

knitr::opts_chunk$set(autodep = TRUE) 
knitr::opts_chunk$set(warning = FALSE) 
knitr::opts_chunk$set(message = FALSE) 
knitr::opts_chunk$set(echo = FALSE) 
knitr::opts_chunk$set(cache = TRUE) 

```
## Analysis 
#### function to coalesce columns with missing values

```{r functions, warning=FALSE, message=FALSE, echo=FALSE}

coalesce_all_columns <- function(df) {
  return(coalesce(!!! as.list(df)))
}

```

#### data read in (nia colony)
  - input files from NHP_NIA_raw folder 
    - Copy of #2 - CR Study_FBG Data.xlsx (1/2/2019) 
    - Copy of #4 - Other Rhesus_FBG Data.xlsx (4/9/2019) 
    - Copy of NHP-NIA_average FBG_042319.xlsx (4/23/2019) 
    - Copy of NHP-NIA_LT CR dexa data_Michel.xlsx (10/15/2019) 
    - NIA_mortality_list2_update032020_v2.xlsx (3/20/20) 

```{r nhp_data_nia, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE} 

#pull data from excel tabs
#CR_FBG <- read_excel("raw files/NHP_NIA_raw/Copy of #2 - CR Study_FBG Data.xlsx", skip = 2) 
#other_FBG <- read_excel("raw files/NHP_NIA_raw/Copy of #4  - Other Rhesus_FBG Data.xlsx") 
#CR_dexa <- read_excel("raw files/NHP_NIA_raw/Copy of NHP-NIA_LT CR dexa data_Michel.xlsx", sheet = "initial #dataset") 
#surv_grp <- read_excel("raw files/NHP_NIA_raw/NIA_mortality_list2_update032020_v2.xlsx") 

CR_FBG <- read_excel("/Users/shiromaej/Box/ISG Code Check/data/nhp/NHP_NIA_raw/Copy of #2 - CR Study_FBG Data.xlsx", skip = 2) 
other_FBG <- read_excel("/Users/shiromaej/Box/ISG Code Check/data/nhp/NHP_NIA_raw/Copy of #4  - Other Rhesus_FBG Data.xlsx") 
CR_dexa <- read_excel("/Users/shiromaej/Box/ISG Code Check/data/nhp/NHP_NIA_raw/Copy of NHP-NIA_LT CR dexa data_Michel.xlsx", sheet = "initial dataset") 
surv_grp <- read_excel("/Users/shiromaej/Box/ISG Code Check/data/nhp/NHP_NIA_raw/NIA_mortality_list2_update032020_v2.xlsx") 


```

#### glucose cleaning (nia colony)
- CR study (CR_FBG2)
- other rhesus (other_FBG2)

```{r nhp_gluc_nia, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE} 

#format glucose datasets
CR_FBG2 <- CR_FBG %>% 
  rename(idno = ID, sex = Sex, diet = Diet, agegrp = AgeGrp, date = `Date of Measure`) %>% 
  mutate(idno = toupper(idno)) 
CR_FBG2$date <- as.Date(CR_FBG2$date)

other_FBG2 <- other_FBG %>% 
  rename(idno = ID, sex = Sex, diet = Diet, agegrp = `Age Group`, date = `Date of Measurement`, age = `Age @ \r\nMeasure (mos)`, gluc = `Fasting Blood Glucose`, source = Source) %>% 
  mutate(idno = toupper(idno)) 
other_FBG2$date <- as.Date(other_FBG2$date) 

```

#### body composition cleaning (nia colony)
- CR study (CR_dexa4)
  - calculate body weight from fat and % fat 
  - normalize units to kg 

```{r nhp_bc_nia, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE} 

#format dexa datset
CR_dexa2 <- CR_dexa %>% 
  select(`Animal ID`, Sex, `Age Group`, Diet, `Total Body Measure`, `Measurement Date`, `Measure Value`, `Age at Measurement`) %>% 
  rename(idno = `Animal ID`, sex = Sex, agegrp = `Age Group`, diet = Diet, measure = `Total Body Measure`,  date = `Measurement Date`, value = `Measure Value`, age_m = `Age at Measurement`) %>% 
  mutate(idno = toupper(idno)) 
CR_dexa2$date = as.Date(CR_dexa2$date)

#spread by measure from long to wide format
CR_dexa2_fat <- CR_dexa2 %>% 
  filter(measure == "Fat") %>% 
  rename(fat = value) %>% 
  select(-measure)

CR_dexa2_lbm <- CR_dexa2 %>% 
  filter(measure == "LBM") %>% 
  rename(lbm = value)  %>% 
  select(-measure)

CR_dexa2_regfat <- CR_dexa2 %>% 
  filter(measure == "Region % fat") %>% 
  rename(reg_percent_fat = value)  %>% 
  select(-measure)

CR_dexa2_tisfat <- CR_dexa2 %>% 
  filter(measure == "Tissue % fat") %>% 
  rename(tis_percent_fat = value)  %>% 
  select(-measure)

CR_dexa3 <- CR_dexa2_fat %>% 
  left_join(CR_dexa2_lbm) %>% 
  left_join(CR_dexa2_regfat) %>%
  left_join(CR_dexa2_tisfat) 

CR_dexa3$fat <- as.numeric(CR_dexa3$fat)
CR_dexa3$lbm <- as.numeric(CR_dexa3$lbm)
CR_dexa3$reg_percent_fat <- as.numeric(CR_dexa3$reg_percent_fat)
CR_dexa3$tis_percent_fat <- as.numeric(CR_dexa3$tis_percent_fat)

#normalize to kg 
CR_dexa4 <- CR_dexa3 %>% 
  group_by(idno, date) %>% 
  ungroup() %>% 
  mutate(bw = fat/(tis_percent_fat/100)) %>% 
  mutate(bw = bw/1000, 
         fat = fat/1000, 
         lbm = lbm/1000)
```

#### survival census cleaning (nia colony)

```{r nhp_data_nia_surv, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE} 

#format survival dataset
surv_grp2 <- surv_grp %>% 
  select(idno, sex, age_grp, diet, dob, dod, death_age, death_euth, group, cause) %>% 
  rename(age_death = death_age, death_type = death_euth) %>% 
  mutate(idno = toupper(idno)) %>% 
  mutate(idno = ifelse(idno == "CH159*", "CH159", 
                ifelse(idno == "47Z*", "47Z",
                ifelse(idno == "O3*", "O3",
                ifelse(idno == "CH77*", "CH77", idno))))) %>% 
  select(-dob) 
surv_grp2$dod <- as.Date(as.numeric(surv_grp2$dod), origin = "1899-12-30") 
surv_grp2$age_death <- as.numeric(surv_grp2$age_death) 

#convert age at death to months 
surv_nia <- surv_grp2 %>% 
  mutate(age_death = age_death*12) %>% 
  group_by(idno) %>% 
  filter(!is.na(idno)) 

```

#### merging datasets for census creation (nia colony)
- other_FBG2
- CR_dexa2
- surv_nia

```{r nhp_census_nia, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE} 

#comprehensive list of all monkeys created using full join (5 from body composition then another 8 from survival)
census_wgaps <- other_FBG2 %>% 
  full_join(CR_dexa2) %>% 
  full_join(surv_nia) %>%
  mutate(strain = "nia") %>% 
  select(idno, sex, source, diet, strain) 

#census creation
census_nia <- census_wgaps %>% 
  group_by(idno) %>% 
  summarise_all(coalesce_all_columns) 

```

#### data read in (wisconsin colony)
  - input files from NHP_WIS_raw folder
    - Copy of Control Glucose_.xlsx (4/28/2019)
    - Copy of WIS-NHP DXA data_Michel.xlsx (4/29/2019)
    - Euth.Nat.xlsx (1/17/2020)
    - GlucoseDate.DOBDOD.xlsx (12/2/2019)

```{r nhp_data_wis, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE}

#FBG <- read_excel("raw files/NHP_WIS_raw/Copy of Control Glucose_.xlsx", sheet = "Basal Glucose") 
#IVGTT <- read_excel("raw files/NHP_WIS_raw/Copy of Control Glucose_.xlsx", sheet = "Glucose from IVGTT") 
#DEXA <- read_excel("raw files/NHP_WIS_raw/Copy of WIS-NHP DXA data_Michel.xlsx", sheet = "raw data") 
#surv_cat <- read_excel("raw files/NHP_WIS_raw/Euth.Nat.xlsx") 
#surv_date <- read_excel("raw files/NHP_WIS_raw/GlucoseDate.DOBDOD.xlsx")

FBG <- read_excel("/Users/shiromaej/Box/ISG Code Check/data/nhp/NHP_WIS_raw/Copy of Control Glucose_.xlsx", sheet = "Basal Glucose") 
IVGTT <- read_excel("/Users/shiromaej/Box/ISG Code Check/data/nhp/NHP_WIS_raw/Copy of Control Glucose_.xlsx", sheet = "Glucose from IVGTT") 
DEXA <- read_excel("/Users/shiromaej/Box/ISG Code Check/data/nhp/NHP_WIS_raw/Copy of WIS-NHP DXA data_Michel.xlsx", sheet = "raw data") 
surv_cat <- read_excel("/Users/shiromaej/Box/ISG Code Check/data/nhp/NHP_WIS_raw/Euth.Nat.xlsx") 
surv_date <- read_excel("/Users/shiromaej/Box/ISG Code Check/data/nhp/NHP_WIS_raw/GlucoseDate.DOBDOD.xlsx")

```
  
#### glucose creation (wisconsin colony)
- using provided age since dates are not provided in glucose file
- pulled in sex from IVGTT as some are missing from glucose file 
- converted age to months
- converted measures into kg 
  
```{r nhp_gluc_bc_wis, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE} 

#format glucose
FBG2 <- FBG %>% 
  rename(idno = I.D., sex = Sex, age = `Age (years)`, gluc = `Fasted Glucose (mg/dl)`) %>% 
  mutate(idno = toupper(idno)) 
FBG2$age = as.numeric(FBG2$age*12) 

#format IVGTT: only used to pull in missing sex for glucose monkeys 
IVGTT2 <- IVGTT %>% 
  select(idno = ID, sex = Sex) %>% 
  mutate(idno = toupper(idno)) %>% 
  distinct()

#format body composition
DEXA2 <- DEXA %>% 
  select(I.D., Group, Diet, Birthdate, `Proc Date`, `Age (years)`, `Scale Wt`, Fat, Lean, BMC, `Fat (%)-DXA`) %>% 
  rename(idno = I.D., group = Group, diet = Diet, dob = Birthdate, date = `Proc Date`, age_yr = `Age (years)`, bw = `Scale Wt`, fat = Fat, lean = Lean, bmc = BMC) %>% 
  mutate(idno = toupper(idno)) 

#convert age to months, measures to kg
DEXA2 <- DEXA2 %>% 
  mutate(age_m = as.numeric(age_yr)*12) %>% 
  select(idno, diet, age_m, bw, fat, lean) %>% 
  mutate(bw = as.numeric(bw)/1000, fat = as.numeric(fat)/1000, lean = as.numeric(lean)/1000)

```

#### survival census creation (wisconsin colony)

```{r nhp_surv_wis, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE} 

## survival formatting and removal of invalid idnos
surv_cat2 <- surv_cat %>% 
  select(ID, Sex, `Age at Death`, Classification, `Nat/Euth`, Diabetes) %>% 
  rename(idno = ID, sex = Sex, age_death = `Age at Death`, age_related = Classification, death_type = `Nat/Euth`, diabetes = Diabetes) %>% 
  mutate(idno = toupper(idno)) %>% 
  mutate(idno = ifelse(idno %in% c("RED TEXT = DIABETIC", "GREEN TEXT = GLUCOREGULATORY IMPAIRMENT (NOT YET DIABETIC)"), NA, idno))

surv_date2 <- surv_date %>% 
  select(I.D., Sex, DOD, DOB) %>% 
  rename(idno = I.D., sex = Sex, dod = DOD, dob = DOB) %>% 
  mutate(idno = toupper(idno)) 
surv_date2$dob <- as.Date(surv_date2$dob) 
surv_date2$dod <- as.Date(surv_date2$dod) 

#merging survival files and removing empty rows
surv_wis <- surv_cat2 %>% 
  full_join(surv_date2) %>% 
  filter(!is.na(idno)) %>%
  mutate(age_death = age_death*12) 

```

#### merging datasets for census creation (wisconsin colony) 
- DEXA2 
- surv_wis 

```{r nhp_census_wis, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE} 

## census creation
census_wgaps <- FBG2 %>% 
  full_join(IVGTT2) %>%
  full_join(DEXA2) %>% 
  full_join(surv_wis) %>% 
  mutate(strain = "wis") %>% 
  select(idno, sex, diet, dob, strain, diabetes) 

#census removal of empty rows 
census_wis <- census_wgaps %>% 
  group_by(idno) %>% 
  summarise_all(coalesce_all_columns) %>% 
  filter(!is.na(idno)) 

```

#### merging datasets for glucose creation (all colonies)
- other_FBG2 (nia)
- FBG2 (wis)

```{r nhp_gluc_all, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE}

#merging glucose files
gluc_all <- other_FBG2 %>% 
  full_join(FBG2) %>% 
  select(idno, age, date, gluc) %>% 
  rename(age_wk = age) 

#fill in gaps created due to merged datasets missing different columns of data
glucose <- gluc_all %>% 
  group_by(idno, age_wk) %>% 
  summarise_all(coalesce_all_columns) %>% 
  ungroup() 

```

#### merging datasets for body composition creation (all colonies)
- CR_dexa4 (nia)
- DEXA2 (wis)

```{r nhp_bodycomp_all, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE}

#standardize sex and diet values 
bodycomp_all <- CR_dexa4 %>% 
  full_join(DEXA2) %>% 
  mutate(
    sex = ifelse(sex %in% c("F", "Female"), "F", 
          ifelse(sex %in% c("M", "Male"), "M", NA)), 
    diet = ifelse(diet %in% c("CON", "C"), "CON", 
          ifelse(diet %in% c("CR", "DR"), "CR", NA))
  ) %>% 
  select(idno, date, age_m, bw, fat, lbm) %>% 
  rename(age_wk = age_m) 

#fill in gaps created due to merged datasets missing different columns of data
bodycomp <- bodycomp_all %>% 
  group_by(idno, age_wk) %>% 
  summarise_all(coalesce_all_columns) 

```

#### merging datasets for census creation (all colonies)
- census_nia (nia)
- census_wis (wis)
- set all monkeys missing diabetes information as "Non-Diabetic"

```{r nhp_census_all, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE}

#standardize sex and diet values 
census_all <- census_wis %>% 
  full_join(census_nia) %>% 
  mutate( 
    sex = ifelse(sex %in% c("F", "Female"), "F", 
          ifelse(sex %in% c("M", "Male"), "M", NA)), 
    diet = ifelse(diet %in% c("CON", "C"), "CON", 
          ifelse(diet %in% c("CR", "DR"), "CR", NA)) 
  ) 

#create diabetes variable
census <- census_all %>% 
  group_by(idno) %>% 
  summarise_all(coalesce_all_columns) %>% 
  mutate(diabetes2 = ifelse(diabetes == 2, "Diabetic", 
                     ifelse(diabetes == 1, "Non-Diabetic", 
                            "Non-Diabetic"))) %>% 
  mutate(diabetes2 = ifelse(is.na(diabetes), "Non-Diabetic", diabetes2)) 

``` 

#### merging datasets for census creation (all colonies)
- surv_nia (nia) 
- surv_wis (wis) 

```{r nhp_surv_all, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE}

#standardize sex and diet values 
surv_all <- surv_wis %>% 
  full_join(surv_nia) %>% 
  mutate( 
    sex = ifelse(sex %in% c("female", "Female"), "F", 
          ifelse(sex %in% c("male", "Male"), "M", NA)), 
    diet = ifelse(diet %in% c("C"), "CON", 
          ifelse(diet %in% c("E"), "CR", NA))
  ) %>% 
  select(idno, age_death, dob, dod, death_type, age_grp, cause) %>% 
  mutate(death_type = ifelse(death_type %in% c("euth", "Euth", "eut"), "euth", 
                      ifelse(death_type %in% c("dead", "Nat"), "dead", NA))) %>% 
  rename(cod = cause) 

#fill in gaps created due to merged datasets missing different columns of data
surv <- surv_all %>% 
  group_by(idno) %>% 
  summarise_all(coalesce_all_columns) 


```

#### merging datasets into a main file (all colonies)
- glucose 
- bodycomp 
- census 

```{r nhp_main_all, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE, include=FALSE}

main_all <- glucose %>% 
  full_join(bodycomp) %>% 
  full_join(census, by = "idno") %>% 
  ungroup() %>% 
  mutate(
    age_censor = ifelse(idno %in% c("464", "9G3", "9GI", "CH159", "47Z"), 0, 1),
    health_censor = ifelse(idno %in% c("87L", "CH48"), 0, 1), 
    diet_censor = ifelse(diet != "CR" | is.na(diet), 1, 0)
  )

#354 
total <- n_distinct(main_all$idno) 
#91
cr <- n_distinct(main_all$idno[main_all$diet_censor == 0]) 
#263
non_cr <- n_distinct(main_all$idno[main_all$diet_censor == 1]) 
#259
non_cr_age_rest <- n_distinct(main_all$idno[main_all$diet_censor == 1 & main_all$age_censor == 1]) 
#258
non_cr_age_rest_health <- n_distinct(main_all$idno[main_all$diet_censor == 1 & main_all$age_censor == 1 & main_all$health_censor == 1]) 
#254
non_cr_age_rest_health_valid <- n_distinct(main_all$idno[main_all$diet_censor == 1 & main_all$age_censor == 1 & main_all$health_censor == 1 & (!is.na(main_all$gluc) | !is.na(main_all$bw) | !is.na(main_all$fat)) ] ) 

```

#### study bounded to non-caloric restriction diet monkeys
- total monkeys: `r total` 
  - monkeys on caloric restriction diet: `r cr` 
  - monkeys on control diet: `r non_cr` 

#### restriction of monkeys with young deaths before 10 years of age
- "464", "9G3", "9GI", "CH159", "47Z"
  - monkeys after young death censor: `r non_cr_age_rest`
    - one already removed for diet

#### restriction of monkeys removed from study for health reasons
- "87L" (recycled 12/96), "CH48" (removed for cachexia 1/90)
  - monkeys after health censor: `r non_cr_age_rest_health`
    - one already removed for diet
    
#### excluding monkeys with no valid metabolic index values
- "833", "O3", "O89", "OO5"
  - monkeys after valid value censor: `r non_cr_age_rest_health_valid`

```{r nhp_ratios, warning=FALSE, message=FALSE, echo=FALSE,  include=FALSE}

#mark for removal 
main_all2 <- main_all %>% 
  filter(diet_censor == 1,
         health_censor == 1,
         age_censor == 1,
         (!is.na(gluc) | !is.na(bw) | !is.na(fat)))

main_all2_n <- n_distinct(main_all2$idno)

## sexstrain tables 
subgroup <- main_all2 %>% 
  select(idno, sex, strain, diabetes2) %>% 
  distinct() 

sextable <- table(subgroup$sex)
straintable <- table(subgroup$strain)
diabetestable <- table(subgroup$diabetes2)

nhp_sex_ratio = format(round(((sextable[1] / (sextable[1] + sextable[2]))*100), 2), nsmall = 2)
nhp_strain_ratio = format(round(((straintable[2] / (straintable[1] + straintable[2]))*100), 2), nsmall = 2)
nhp_diabetes_ratio = format(round(((diabetestable[1] / (diabetestable[1] + diabetestable[2]))*100), 2), nsmall = 2)

```

#### population characteristics 
- `r main_all2_n` monkeys with valid indices 
- `r nhp_sex_ratio` % female 
- `r nhp_strain_ratio` % wis 
- `r nhp_diabetes_ratio` % diabetic 

#### age characteristics

```{r nhp_age_desc, warning=FALSE, message=FALSE, echo=FALSE}

main_all2_desc <- main_all2 %>%
group_by(idno, sex, strain) %>%
summarise(
  age_wk = min(age_wk, na.rm=T)
) %>% 
  mutate(age_wk = ifelse(age_wk == "Inf", NA, age_wk)) 

main_all2_desc2_all <- format(round(psych::describe(main_all2_desc$age_wk), 2), nsmall=2) %>% 
  mutate_each(funs(as.numeric(.))) %>%
  mutate(group1 = "all") %>% 
  select(-c("vars")) 
main_all2_desc_sex <- as.data.frame(describeBy(main_all2_desc$age_wk, main_all2_desc$sex, mat=TRUE, digits = 2)) %>% 
  select(-c("item", "vars")) %>% 
  mutate(n = as.numeric(n))
main_all2_desc_strain <- as.data.frame(describeBy(main_all2_desc$age_wk, main_all2_desc$strain, mat=TRUE, digits = 2)) %>% 
  select(-c("item", "vars")) %>% 
  mutate(n = as.numeric(n))

main_all2_desc2 <- bind_rows(main_all2_desc2_all, main_all2_desc_sex, main_all2_desc_strain) %>% 
  select(group1, everything()) 

```

```{r nhp_age_desc_kable, warning=FALSE, message=FALSE, echo=FALSE}

kable(main_all2_desc2, caption = "Age Characteristics",row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### frequencies of measures

```{r nhp_follow_up, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}

n_asmt <- main_all2 %>% 
  group_by(idno, sex) %>% 
  summarise_all(funs( 
    n = sum(!is.na(.)), 
    sd = format(round(sd(!is.na(.)), 2), nsmall=2) 
  )) 

gluc_asmt <- format(round(as.data.frame(psych::describe(n_asmt$gluc_n)), 2), nsmall=2) %>% 
  mutate(indices = "glucose")
bw_asmt <- format(round(as.data.frame(psych::describe(n_asmt$bw_n)), 2), nsmall=2) %>% 
  mutate(indices = "body weight")
fat_asmt <- format(round(as.data.frame(psych::describe(n_asmt$fat_n)), 2), nsmall=2) %>% 
  mutate(indices = "fat")

all_asmt <- bind_rows(gluc_asmt, bw_asmt, fat_asmt) %>% 
  select(indices, everything(), -vars)

nhp_gluc_asmt <- gluc_asmt
nhp_bw_asmt <- bw_asmt
nhp_fat_asmt <- fat_asmt

```

```{r asmt_kable, warning=FALSE, message=FALSE, echo=FALSE}

kable(all_asmt, caption = "Assessment Frequencies", row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### census_c16_death creation
  - merge 
    - census
    - surv
    - main_all2
  - restrict to non-caloric restriction diet monkeys (from original census and survival file)
  - restrict to monkeys that lived past 10 years (from original census and survival file)
    - "464", "9G3", "9GI", "CH159", "47Z"
  - excluding monkeys with no valid metabolic index values
    - "833", "O3", "O89", "OO5"
  - create all cause death censor category
  
***

```{r nhp_censor_df, warning=FALSE, message=FALSE, echo=FALSE}

#354 monkeys 
census_c16_death <- census %>% 
  left_join(surv) %>% 
  left_join(main_all2) %>% 
  filter(diet_censor == 1, age_censor == 1) %>% #258
  group_by(idno) %>% 
  summarise(age_first = min(age_wk, na.rm=T), 
            age_last = max(age_wk, na.rm=T), 
            lastvisit = max(date, na.rm=T), 
            firstvisit = min(date, na.rm=T) 
            ) %>% 
  #recode monkeys without any age_wks to NA 
  mutate(age_last = ifelse(age_last == Inf, NA,
                    ifelse(age_last == -Inf, NA, age_last))) %>% 
  distinct() %>% 
  left_join(surv, by = c("idno")) %>% 
  left_join(census, by ="idno") %>% 
  mutate( 
    dead_censor = ifelse(is.na(age_death), 0, 1),
    le_wk = ifelse(!is.na(age_death), age_death, 
                ifelse(!is.na(age_last), age_last, NA)), 
    fu_age_wk = le_wk - age_first
  ) %>%
  filter(!(idno %in% c("833", "O3", "O89", "OO5"))) %>% 
  select(idno, dead_censor, fu_age_wk, le_wk, cod, age_first, age_last, age_death, death_type, lastvisit, firstvisit, diabetes2, sex, strain)

## 
dctable1 <- as.data.frame(table(census_c16_death$dead_censor)) 
fu_time_desc <- format(round(as.data.frame(psych::describe(as.numeric(census_c16_death$fu_age_wk))), 2), nsmall=2) 

```

#### death frequency

```{r death_cat, warning=FALSE, message=FALSE, echo=FALSE} 
kable(dctable1, caption = "Death Frequency", col.names = c("Dead = 1", "Frequency"), row.names = FALSE, align = "c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

***

#### follow up time

```{r followup_cat, warning=FALSE, message=FALSE, echo=FALSE} 
kable(fu_time_desc, caption = "Follow Up Time", row.names = FALSE, align = "c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE) 
```

*** 

#### life expectancy

```{r nhp_life_expectancy, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}

create_LE <- function(data, criteria, filter){

life_exp <- data %>% 
  ungroup() %>% 
  filter(!!filter) %>% 
  summarise( 
            total = n_distinct(idno), 
            dead = n_distinct(idno[!!criteria]), 
            life_exp_wk = round(mean(le_wk[!!criteria]), 2), 
            sd = round(sd(as.numeric(le_wk[!!criteria])), 2), 
            se = round(sd(as.numeric(le_wk[!!criteria])) / sqrt(dead),2),
            lcl = round(life_exp_wk - 1.96 * se, 2), 
            ucl = round(life_exp_wk + 1.96 * se, 2) 
  ) 
life_exp$life_exp_wk <- as.numeric(life_exp$life_exp_wk) 

return(life_exp)

}

life_exp <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(sex %in% c("M", "F"))) %>% 
  mutate(group = "all") 
life_exp_f <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(sex %in% c("F"))) %>% 
  mutate(group = "female") 
life_exp_m <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(sex %in% c("M"))) %>% 
  mutate(group = "male") 
life_exp_nia <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(strain %in% c("nia"))) %>% 
  mutate(group = "nia") 
life_exp_wis <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(strain %in% c("wis"))) %>% 
  mutate(group = "wis") 
life_exp_diab <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(diabetes2 %in% c("Diabetic"))) %>% 
  mutate(group = "diabetes") 
life_exp_nondiab <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(diabetes2 %in% c("Non-Diabetic"))) %>% 
  mutate(group = "diabetes") 

life_exp2 <- bind_rows(life_exp, life_exp_f, life_exp_m, life_exp_nia, life_exp_wis) %>% 
  select(group, everything())
  
## for isg template
nhp_life_exp <- life_exp 
nhp_life_exp_f <- life_exp_f
nhp_life_exp_m <- life_exp_m 
nhp_life_exp_nia <- life_exp_nia 
nhp_life_exp_wis <- life_exp_wis 
nhp_life_exp_diab <- life_exp_diab 
nhp_life_exp_nondiab <- life_exp_nondiab

```

```{r nhp_life_exp_kable, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}
kable(life_exp2[,c(1:5)], caption = "Life Expectancy", col.names = c("Sex", "Total", "Dead", "Life Expectancy <br> (months)", "sd"), row.names = TRUE, align = "c", escape = FALSE) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left") 
```

***

#### age at peak index value creation 

```{r nhp_life_peaks, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}

#functions to create peaks for different groups
lifespan_peaks <- function(metric, data, le, filter){

#set boundary conditions for peak to lowest and highest life expectancy timepoints
low_bound <- floor(le$life_exp_wk*0.25)
high_bound <- floor(le$life_exp_wk*1.25)

#subset arguments
sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)") 
metric2 <- paste0("!is.na(", metric, ")") 

#subset for non-missing data
df <- base::subset(data, eval(parse(text = metric2))) 

#subset data by xlimit and subgroup restraints
peaks_df <- subset(df, eval(parse(text = sub))) %>% 
  arrange(age_wk) %>% 
  filter(!!filter) 

#gam model fit using geom_smooth parameters
gam_fit <- mgcv::gam(eval(parse(text = metric)) ~ s(age_wk, bs = "cs"), method = "REML", data = peaks_df) 

#create fitted values using gam model from dataframe age values
predict1 <- as.data.frame(predict.gam(gam_fit, peaks_df, se.fit=TRUE)) 

#create n and sd by age group
fit2 <- peaks_df %>% 
  bind_cols(predict1) %>% 
  select(age_wk, fit, se.fit) %>% 
  filter(age_wk >= low_bound, age_wk <= high_bound) %>%
  group_by(age_wk) %>% 
  mutate( 
    n = n(), 
    sd = se.fit*sqrt(n) 
  ) %>% 
  ungroup() 

#create low bound peak object
low_bound_peak <- fit2 %>% 
  filter(age_wk == min(age_wk)) %>% 
  mutate(le = (age_wk/le$life_exp_wk)*100) %>% 
  select(age_wk, le, fit, se.fit, n, sd) %>% 
  distinct()

#create high bound peak object
high_bound_peak <- fit2 %>% 
  filter(age_wk == max(age_wk)) %>% 
  mutate(le = (age_wk/le$life_exp_wk)*100) %>% 
  select(age_wk, le, fit, se.fit, n, sd) %>% 
  distinct()

#pulling peak age and value
infl <- NULL
infl$logic <- c(NA, diff(diff(fit2$fit)) < 0, NA)
infl <- as.data.frame(infl)

#finding age of the first point significantly different from peak for the lower range
#setting to lower bound if no values are present
peak <- fit2 %>% 
  bind_cols(infl) %>% 
  filter(infl$logic == "TRUE") %>% 
  group_by(age_wk, n, sd, se.fit) %>%
  summarise(
    fit = max(fit)
  )  %>% 
  ungroup() %>% 
  filter(fit == max(fit, na.rm=T)) %>% 
  mutate(le = (age_wk/le$life_exp_wk)*100)

#finding age of the first point significantly different from peak for the lower range
#setting to lower bound if no values are present
fit_diff_min <- tryCatch(
fit2 %>% 
  filter(age_wk < peak$age_wk) %>% 
  mutate(fit_diff = fit - peak$fit) %>% 
  mutate(age_diff = peak$age_wk - age_wk) %>% 
  mutate(pval = tsum.test(fit_diff, sd, n, mu = 0, conf.level = 0.95)$p.value) %>% 
  filter(pval <= 0.05) %>% 
  filter( 
    age_diff == min(age_diff) 
    ) %>% 
  distinct() %>% 
  mutate(le = (age_wk/le$life_exp_wk)*100) %>% 
  select(age_wk, le, fit), 
error=function(e) low_bound_peak)

#finding age of the first point significantly different from peak for the higher range
#setting to higher bound if no values are present
fit_diff_max <- tryCatch(
fit2 %>% 
  filter(age_wk > peak$age_wk) %>% 
  mutate(fit_diff = fit - peak$fit) %>% 
  mutate(age_diff = age_wk - peak$age_wk) %>% 
  mutate(pval = tsum.test(fit_diff, sd, n, mu = 0, conf.level = 0.95)$p.value) %>% 
  filter(pval <= 0.05) %>% 
  filter( 
    age_diff == min(age_diff) 
    ) %>% 
  distinct() %>% 
  mutate(le = (age_wk/le$life_exp_wk)*100) %>% 
  select(age_wk, le, fit),
error=function(e) high_bound_peak)

#setting to lower bound if no values are present
if(dim(fit_diff_min)[1] == 0){
  fit_diff_min <- low_bound_peak
}else{
}

#setting to higher bound if no values are present
if(dim(fit_diff_max)[1] == 0){
  fit_diff_max <- high_bound_peak
}else{
}

#putting all outputs into list 
all_peaks <- list(peak, fit_diff_min, fit_diff_max, high_bound_peak, low_bound_peak) 
names(all_peaks) <- c("all", "min", "max", "highbound", "lowbound")

return(all_peaks) 

} 

gluc_peak <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp, filter = quote(sex %in% c("M", "F"))) 
bw_peak <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp, filter = quote(sex %in% c("M", "F"))) 
fat_peak <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp, filter = quote(sex %in% c("M", "F"))) 

gluc_peak_male <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp_m, filter = quote(sex %in% c("M"))) 
bw_peak_male <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp_m, filter = quote(sex %in% c("M"))) 
fat_peak_male <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp_m, filter = quote(sex %in% c("M"))) 

gluc_peak_female <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp_f, filter = quote(sex %in% c("F"))) 
bw_peak_female <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp_f, filter = quote(sex %in% c("F"))) 
fat_peak_female <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp_f, filter = quote(sex %in% c("F"))) 

gluc_peak_nia <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp_nia, filter = quote(strain %in% c("nia"))) 
bw_peak_nia <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp_nia, filter = quote(strain %in% c("nia"))) 
fat_peak_nia <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp_nia, filter = quote(strain %in% c("nia"))) 

gluc_peak_wis <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp_wis, filter = quote(strain %in% c("wis"))) 
bw_peak_wis <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp_wis, filter = quote(strain %in% c("wis"))) 
fat_peak_wis <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp_wis, filter = quote(strain %in% c("wis"))) 

gluc_peak_diab <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp_diab, filter = quote(diabetes2 %in% c("Diabetic")))
bw_peak_diab <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp_diab, filter = quote(diabetes2 %in% c("Diabetic")))
fat_peak_diab <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp_diab, filter = quote(diabetes2 %in% c("Diabetic")))

gluc_peak_nondiab <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp_nondiab, filter = quote(diabetes2 %in% c("Non-Diabetic"))) 
bw_peak_nondiab <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp_nondiab, filter = quote(diabetes2 %in% c("Non-Diabetic"))) 
fat_peak_nondiab <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp_nondiab, filter = quote(diabetes2 %in% c("Non-Diabetic"))) 

## 
gluc_peak_per <- (gluc_peak$all$age_wk / as.numeric(life_exp$life_exp_wk)) 
bw_peak_per <- (bw_peak$all$age_wk / as.numeric(life_exp$life_exp_wk)) 
fat_peak_per <- (fat_peak$all$age_wk / as.numeric(life_exp$life_exp_wk)) 

gluc_peak_per_min <- (gluc_peak$min$age_wk / as.numeric(life_exp$life_exp_wk)) 
bw_peak_per_min <- (bw_peak$min$age_wk / as.numeric(life_exp$life_exp_wk)) 
fat_peak_per_min <- (fat_peak$min$age_wk / as.numeric(life_exp$life_exp_wk)) 

gluc_peak_per_max <- (gluc_peak$max$age_wk / as.numeric(life_exp$life_exp_wk)) 
bw_peak_per_max <- (bw_peak$max$age_wk / as.numeric(life_exp$life_exp_wk)) 
fat_peak_per_max <- (fat_peak$max$age_wk / as.numeric(life_exp$life_exp_wk)) 
##
gluc_peak_per_m <- (gluc_peak_male$all$age_wk / as.numeric(life_exp$life_exp_wk)) 
bw_peak_per_m <- (bw_peak_male$all$age_wk / as.numeric(life_exp$life_exp_wk)) 
fat_peak_per_m <- (fat_peak_male$all$age_wk / as.numeric(life_exp$life_exp_wk)) 

gluc_peak_per_m_min <- (gluc_peak_male$min$age_wk / as.numeric(life_exp$life_exp_wk)) 
bw_peak_per_m_min <- (bw_peak_male$min$age_wk / as.numeric(life_exp$life_exp_wk)) 
fat_peak_per_m_min <- (fat_peak_male$min$age_wk / as.numeric(life_exp$life_exp_wk)) 

gluc_peak_per_m_max <- (gluc_peak_male$max$age_wk / as.numeric(life_exp$life_exp_wk)) 
bw_peak_per_m_max <- (bw_peak_male$max$age_wk / as.numeric(life_exp$life_exp_wk)) 
fat_peak_per_m_max <- (fat_peak_male$max$age_wk / as.numeric(life_exp$life_exp_wk)) 

gluc_peak_per_f <- (gluc_peak_female$all$age_wk / as.numeric(life_exp$life_exp_wk)) 
bw_peak_per_f <- (bw_peak_female$all$age_wk / as.numeric(life_exp$life_exp_wk)) 
fat_peak_per_f <- (fat_peak_female$all$age_wk / as.numeric(life_exp$life_exp_wk)) 

gluc_peak_per_f_min <- (gluc_peak_female$min$age_wk / as.numeric(life_exp$life_exp_wk)) 
bw_peak_per_f_min <- (bw_peak_female$min$age_wk / as.numeric(life_exp$life_exp_wk)) 
fat_peak_per_f_min <- (fat_peak_female$min$age_wk / as.numeric(life_exp$life_exp_wk)) 

gluc_peak_per_f_max <- (gluc_peak_female$max$age_wk / as.numeric(life_exp$life_exp_wk)) 
bw_peak_per_f_max <- (bw_peak_female$max$age_wk / as.numeric(life_exp$life_exp_wk)) 
fat_peak_per_f_max <- (fat_peak_female$max$age_wk / as.numeric(life_exp$life_exp_wk)) 

```

#### main_cat2 creation 
- merge 
  - main_all2 
  - census_c16_death 
- create categories by survival length 
  - needed for survival bias sensitivity 
  
```{r nhp_main_cat, warning=FALSE, message=FALSE, echo=FALSE}
 
create_main_cat <- function(life_exp){
  
main_cat <- main_all2 %>% 
  left_join(census_c16_death) %>% 
  ungroup() %>% 
  mutate( 
    
    whenvisit = 
                ifelse((age_wk >= (life_exp$life_exp_wk*0)) & (age_wk < (life_exp$life_exp_wk*0.125)), 0,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.125)) & (age_wk < (life_exp$life_exp_wk*0.375)), 25,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.375)) & (age_wk < (life_exp$life_exp_wk*0.625)), 50,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.625)) & (age_wk < (life_exp$life_exp_wk*0.875)), 75,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.875)) & (age_wk < (life_exp$life_exp_wk*1.125)), 100,
                ifelse((age_wk >= (life_exp$life_exp_wk*1.125)), 125, NA)))))), 
    whendead =  
                ifelse((le_wk >= (life_exp$life_exp_wk*0)) & (le_wk < (life_exp$life_exp_wk*0.125)), 0,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.125)) & (le_wk < (life_exp$life_exp_wk*0.375)), 25,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.375)) & (le_wk < (life_exp$life_exp_wk*0.625)), 50,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.625)) & (le_wk < (life_exp$life_exp_wk*0.875)), 75,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.875)) & (le_wk < (life_exp$life_exp_wk*1.125)), 100,
                ifelse((le_wk >= (life_exp$life_exp_wk*1.125)), 125, NA)))))), 

    
    gluc_25 = ifelse(whenvisit == 25, gluc, NA),
    gluc_50 = ifelse(whenvisit == 50, gluc, NA),
    gluc_75 = ifelse(whenvisit == 75, gluc, NA),
    gluc_100 = ifelse(whenvisit == 100, gluc, NA),
    gluc_125 = ifelse(whenvisit == 125, gluc, NA), 
    
    bw_25 = ifelse(whenvisit == 25, bw, NA),
    bw_50 = ifelse(whenvisit == 50, bw, NA),
    bw_75 = ifelse(whenvisit == 75, bw, NA),
    bw_100 = ifelse(whenvisit == 100, bw, NA),
    bw_125 = ifelse(whenvisit == 125, bw, NA),
    
    fat_25 = ifelse(whenvisit == 25, fat, NA),
    fat_50 = ifelse(whenvisit == 50, fat, NA),
    fat_75 = ifelse(whenvisit == 75, fat, NA),
    fat_100 = ifelse(whenvisit == 100, fat, NA), 
    fat_125 = ifelse(whenvisit == 125, fat, NA), 
    
    #survival specific categories
    dgluc_25 = ifelse(whendead >= 25 & whenvisit == 25, gluc, NA),
    dgluc_50 = ifelse(whendead >= 50 & whenvisit == 50, gluc, NA),
    dgluc_75 = ifelse(whendead >= 75 & whenvisit == 75, gluc, NA),
    dgluc_100 = ifelse(whendead >= 100 & whenvisit == 100, gluc, NA),
    dgluc_125 = ifelse(whendead >= 125 & whenvisit == 125, gluc, NA),
    
    dbw_25 = ifelse(whendead >= 25 & whenvisit == 25, bw, NA),
    dbw_50 = ifelse(whendead >= 50 & whenvisit == 50, bw, NA),
    dbw_75 = ifelse(whendead >= 75 & whenvisit == 75, bw, NA),
    dbw_100 = ifelse(whendead >= 100 & whenvisit == 100, bw, NA),
    dbw_125 = ifelse(whendead >= 125 & whenvisit == 125, bw, NA),
    
    dfat_25 = ifelse(whendead >= 25 & whenvisit == 25, fat, NA),
    dfat_50 = ifelse(whendead >= 50 & whenvisit == 50, fat, NA),
    dfat_75 = ifelse(whendead >= 75 & whenvisit == 75, fat, NA),
    dfat_100 = ifelse(whendead >= 100 & whenvisit == 100, fat, NA),
    dfat_125 = ifelse(whendead >= 125 & whenvisit == 125, fat, NA)
    
  ) %>% 
  
  group_by(idno, whendead) %>% 
  summarise( 

    gluc_25_m = mean(gluc_25, na.rm = TRUE),
    gluc_50_m = mean(gluc_50, na.rm = TRUE),
    gluc_75_m = mean(gluc_75, na.rm = TRUE),
    gluc_100_m = mean(gluc_100, na.rm = TRUE),
    gluc_125_m = mean(gluc_125, na.rm = TRUE),
    
    bw_25_m = mean(bw_25, na.rm = TRUE),
    bw_50_m = mean(bw_50, na.rm = TRUE),
    bw_75_m = mean(bw_75, na.rm = TRUE),
    bw_100_m = mean(bw_100, na.rm = TRUE),
    bw_125_m = mean(bw_125, na.rm = TRUE),
    
    fat_25_m = mean(fat_25, na.rm = TRUE),
    fat_50_m = mean(fat_50, na.rm = TRUE),
    fat_75_m = mean(fat_75, na.rm = TRUE),
    fat_100_m = mean(fat_100, na.rm = TRUE),
    fat_125_m = mean(fat_125, na.rm = TRUE), 
    
    #survival specific categories
    dgluc_25_m = mean(dgluc_25, na.rm = TRUE),
    dgluc_50_m = mean(dgluc_50, na.rm = TRUE),
    dgluc_75_m = mean(dgluc_75, na.rm = TRUE),
    dgluc_100_m = mean(dgluc_100, na.rm = TRUE),
    dgluc_125_m = mean(dgluc_125, na.rm = TRUE),
    
    dbw_25_m = mean(dbw_25, na.rm = TRUE),
    dbw_50_m = mean(dbw_50, na.rm = TRUE),
    dbw_75_m = mean(dbw_75, na.rm = TRUE),
    dbw_100_m = mean(dbw_100, na.rm = TRUE),
    dbw_125_m = mean(dbw_125, na.rm = TRUE),
    
    dfat_25_m = mean(dfat_25, na.rm = TRUE),
    dfat_50_m = mean(dfat_50, na.rm = TRUE),
    dfat_75_m = mean(dfat_75, na.rm = TRUE),
    dfat_100_m = mean(dfat_100, na.rm = TRUE),
    dfat_125_m = mean(dfat_125, na.rm = TRUE)
  ) 

main_cat2 <- main_cat %>% 
  
  mutate(
    glucd_25_m = gluc_25_m - gluc_50_m, 
    glucd_50_m = gluc_50_m - gluc_50_m, 
    glucd_75_m = gluc_75_m - gluc_50_m, 
    glucd_100_m = gluc_100_m - gluc_50_m,  
    glucd_125_m = gluc_125_m - gluc_50_m, 
    
    bwd_25_m = bw_25_m - bw_50_m, 
    bwd_50_m = bw_50_m - bw_50_m, 
    bwd_75_m = bw_75_m - bw_50_m, 
    bwd_100_m = bw_100_m - bw_50_m,  
    bwd_125_m = bw_125_m - bw_50_m, 
    
    fatd_25_m = fat_25_m - fat_50_m, 
    fatd_50_m = fat_50_m - fat_50_m, 
    fatd_75_m = fat_75_m - fat_50_m, 
    fatd_100_m = fat_100_m - fat_50_m,  
    fatd_125_m = fat_125_m - fat_50_m, 
    
    glucp_25_m = glucd_25_m / gluc_50_m *100, 
    glucp_50_m = glucd_50_m / gluc_50_m *100, 
    glucp_75_m = glucd_75_m / gluc_50_m *100, 
    glucp_100_m = glucd_100_m / gluc_50_m *100, 
    glucp_125_m = glucd_125_m / gluc_50_m *100, 
    
    bwp_25_m = bwd_25_m / bw_50_m *100, 
    bwp_50_m = bwd_50_m / bw_50_m *100, 
    bwp_75_m = bwd_75_m / bw_50_m *100, 
    bwp_100_m = bwd_100_m / bw_50_m *100, 
    bwp_125_m = bwd_125_m / bw_50_m *100, 
    
    fatp_25_m = fatd_25_m / fat_50_m *100, 
    fatp_50_m = fatd_50_m / fat_50_m *100, 
    fatp_75_m = fatd_75_m / fat_50_m *100, 
    fatp_100_m = fatd_100_m / fat_50_m *100, 
    fatp_125_m = fatd_125_m / fat_50_m *100, 
    
##
    cfgluc_25_m = gluc_25_m, 
    cfgluc_50_m = ifelse(whendead >= 50 & is.na(gluc_50_m) & !is.na(gluc_25_m), gluc_25_m, gluc_50_m), 
    cfgluc_75_m = ifelse(whendead >= 75 & is.na(gluc_75_m) & !is.na(gluc_50_m), gluc_50_m, 
                  ifelse(whendead >= 75 & is.na(gluc_75_m) & !is.na(gluc_25_m), gluc_25_m, gluc_75_m)), 
    cfgluc_100_m = ifelse(whendead >= 100 & is.na(gluc_100_m) & !is.na(gluc_75_m), gluc_75_m, 
                   ifelse(whendead >= 100 & is.na(gluc_100_m) & !is.na(gluc_50_m), gluc_50_m, 
                   ifelse(whendead >= 100 & is.na(gluc_100_m) & !is.na(gluc_25_m), gluc_25_m, gluc_100_m))), 
    cfgluc_125_m = ifelse(whendead >= 125 & is.na(gluc_125_m) & !is.na(gluc_100_m), gluc_100_m, 
                   ifelse(whendead >= 125 & is.na(gluc_125_m) & !is.na(gluc_75_m), gluc_75_m,  
                   ifelse(whendead >= 125 & is.na(gluc_125_m) & !is.na(gluc_50_m), gluc_50_m, 
                   ifelse(whendead >= 125 & is.na(gluc_125_m) & !is.na(gluc_25_m), gluc_25_m, gluc_125_m)))), 

    cfbw_25_m = bw_25_m, 
    cfbw_50_m = ifelse(whendead >= 50 & is.na(bw_50_m) & !is.na(bw_25_m), bw_25_m, bw_50_m), 
    cfbw_75_m = ifelse(whendead >= 75 & is.na(bw_75_m) & !is.na(bw_50_m), bw_50_m, 
                  ifelse(whendead >= 75 & is.na(bw_75_m) & !is.na(bw_25_m), bw_25_m, bw_75_m)), 
    cfbw_100_m = ifelse(whendead >= 100 & is.na(bw_100_m) & !is.na(bw_75_m), bw_75_m, 
                   ifelse(whendead >= 100 & is.na(bw_100_m) & !is.na(bw_50_m), bw_50_m, 
                   ifelse(whendead >= 100 & is.na(bw_100_m) & !is.na(bw_25_m), bw_25_m, bw_100_m))), 
    cfbw_125_m = ifelse(whendead >= 125 & is.na(bw_125_m) & !is.na(bw_100_m), bw_100_m, 
                   ifelse(whendead >= 125 & is.na(bw_125_m) & !is.na(bw_75_m), bw_75_m,  
                   ifelse(whendead >= 125 & is.na(bw_125_m) & !is.na(bw_50_m), bw_50_m, 
                   ifelse(whendead >= 125 & is.na(bw_125_m) & !is.na(bw_25_m), bw_25_m, bw_125_m)))), 

    cffat_25_m = fat_25_m, 
    cffat_50_m = ifelse(whendead >= 50 & is.na(fat_50_m) & !is.na(fat_25_m), fat_25_m, fat_50_m), 
    cffat_75_m = ifelse(whendead >= 75 & is.na(fat_75_m) & !is.na(fat_50_m), fat_50_m, 
                  ifelse(whendead >= 75 & is.na(fat_75_m) & !is.na(fat_25_m), fat_25_m, fat_75_m)), 
    cffat_100_m = ifelse(whendead >= 100 & is.na(fat_100_m) & !is.na(fat_75_m), fat_75_m, 
                   ifelse(whendead >= 100 & is.na(fat_100_m) & !is.na(fat_50_m), fat_50_m, 
                   ifelse(whendead >= 100 & is.na(fat_100_m) & !is.na(fat_25_m), fat_25_m, fat_100_m))), 
    cffat_125_m = ifelse(whendead >= 125 & is.na(fat_125_m) & !is.na(fat_100_m), fat_100_m, 
                   ifelse(whendead >= 125 & is.na(fat_125_m) & !is.na(fat_75_m), fat_75_m,  
                   ifelse(whendead >= 125 & is.na(fat_125_m) & !is.na(fat_50_m), fat_50_m, 
                   ifelse(whendead >= 125 & is.na(fat_125_m) & !is.na(fat_25_m), fat_25_m, fat_125_m)))), 

  ) %>% 
  ungroup() 

return(main_cat2)

} 

## 
main_cat2 <- create_main_cat(life_exp) 
main_cat2_f <- create_main_cat(life_exp_f) 
main_cat2_m <- create_main_cat(life_exp_m) 

##for isg template
nhp_main_cat2 <- main_cat2 

``` 

#### main_all3 creation
- merge
  - main_all2
  - census_c16_death
  - main_cat2

```{r nhp_main_all3, warning=FALSE, message=FALSE, echo=FALSE}

main_all3 <- main_all2 %>% 
  left_join(census_c16_death) %>% 
  left_join(main_cat2) %>% 
  
  mutate( 

    whenvisit = as.factor( 
                ifelse((age_wk >= (life_exp$life_exp_wk*0)) & (age_wk < (life_exp$life_exp_wk*0.125)), 0,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.125)) & (age_wk < (life_exp$life_exp_wk*0.375)), 25,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.375)) & (age_wk < (life_exp$life_exp_wk*0.625)), 50,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.625)) & (age_wk < (life_exp$life_exp_wk*0.875)), 75,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.875)) & (age_wk < (life_exp$life_exp_wk*1.125)), 100,
                ifelse((age_wk >= (life_exp$life_exp_wk*1.125)), 125, NA))))))
                ), 

    whendead =  as.factor( 
                ifelse((le_wk >= (life_exp$life_exp_wk*0)) & (le_wk < (life_exp$life_exp_wk*0.125)), 0,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.125)) & (le_wk < (life_exp$life_exp_wk*0.375)), 25,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.375)) & (le_wk < (life_exp$life_exp_wk*0.625)), 50,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.625)) & (le_wk < (life_exp$life_exp_wk*0.875)), 75,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.875)) & (le_wk < (life_exp$life_exp_wk*1.125)), 100,
                ifelse((le_wk >= (life_exp$life_exp_wk*1.125)), 125, NA)))))))
  ) %>% 
  group_by(idno) %>% 
  
  mutate(
    
    age_25 = mean(age_wk[whenvisit == 25], na.rm=T), 
    age_50 = mean(age_wk[whenvisit == 50], na.rm=T), 
    age_75 = mean(age_wk[whenvisit == 75], na.rm=T), 
    age_100 = mean(age_wk[whenvisit == 100], na.rm=T), 
    age_125 = mean(age_wk[whenvisit == 125], na.rm=T) 
    
  ) %>% 
  ungroup() %>% 
  select(-dead_censor, -le_wk, -cod) 

```
 
## Results 
#### distribution figure with trend line
- generalized additive model for main indices
- linear model with quadratic fit for lifespan groups (survival bias analyses)
 
```{r nhp_ss_dist_funct, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

## subset to exclude missing sex 
LE_breaks <- c(life_exp$life_exp_wk*0.25 + 20, life_exp$life_exp_wk*0.50 + 20, life_exp$life_exp_wk*0.75 + 20, life_exp$life_exp_wk*1.00 + 27, life_exp$life_exp_wk*1.25 + 27) 

LE_labels <- c("25%", "50%", "75%", "100%", "125%") 
surv_color <-c("green", "yellow", "red", "purple", "gray")

fcn_plotdist_smooth <- function(metric, x, y, by, xlabel, ylabel, title, xlow, xhigh, peak1, peak2) { 

  metric2 <- paste0("!is.na(", metric, ")") 
  metric3 <- paste0("df$", metric) 
  df <- base::subset(main_all3, eval(parse(text = metric2))) 
  sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)") 
  z <- Inf

  ggplot(subset(df, eval(parse(text = sub))), aes_string(x = x, y = y, colour = by)) + 
    #background 
    annotate("rect", xmin = life_exp$life_exp_wk*1.25 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*1.25 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.2) + 
    annotate("rect", xmin = life_exp$life_exp_wk*1 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*1 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "purple", alpha = 0.2) + 
    annotate("rect", xmin = life_exp$life_exp_wk*0.75 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.75 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*0.50 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.50 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "orange", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*0.25 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.25 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "green", alpha = 0.2) +
    
    #dots
    geom_point(alpha=0.5, position = "jitter", size = 1, aes_string(colour = by), shape=1) +
    #LE lines
    geom_vline(xintercept = life_exp$life_exp_wk *1.25, colour = "black") +
    geom_vline(xintercept = life_exp$life_exp_wk, colour = "purple") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.75, colour = "red") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.5, colour = "orange") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.25, colour = "green") +
    
    #trendlines 
    geom_smooth(aes_string(group = by, linetype = "solid"), linetype = "solid", color = "black", size = 1.8) +
    geom_smooth(aes(linetype = "solid")) + 

    #peaks
    geom_point(aes(x = peak1$all$age_wk, y = peak1$all$fit), size=2, color="black", fill = "white", alpha=0.7, shape=23, stroke=2) +
    geom_point(aes(x = peak2$all$age_wk, y = peak2$all$fit), size=2, color="black", fill = "white", alpha=0.7, shape=23, stroke=2) +

    xlab(xlabel) + 
    ylab(ylabel) + 
    xlim(xlow, xhigh) + 
    
    #second axis
    scale_x_continuous(limits = c(xlow, xhigh), sec.axis = sec_axis(~.*1, breaks = LE_breaks, labels = LE_labels, name="Percent Life Expectancy")) + 
    theme(axis.text.x.top = element_text(angle = 0, margin = margin(t=-2), size = 7, face = "bold", color = "black", vjust = -5), 
        axis.title.x.top = element_text(size = 10, vjust = -1, margin = margin(t=-2)),
        axis.line.x.top = element_line(size = 0, colour = "white"), 
        axis.ticks.x.top = element_blank()) + 
    scale_linetype_manual(labels = c("All"), name = "", values = c("solid", "dashed", "solid"), guide = FALSE)
} 

fcn_plotdist_lm <- function(metric, x, y, by, xlabel, ylabel, title, xlow, xhigh) { 

  metric2 <- paste0("!is.na(", metric, ")") 
  metric3 <- paste0("df$", metric) 
  main_all3 <- main_all3 %>% filter(as.numeric(whenvisit) <= as.numeric(whendead)) %>% filter(as.numeric(whendead) > 0) 
  df <- base::subset(main_all3, eval(parse(text = metric2))) 
  sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)") 
  z <- Inf

  ggplot(subset(df, eval(parse(text = sub))), aes_string(x = x, y = y, colour = by)) + 
    #background 
    annotate("rect", xmin = life_exp$life_exp_wk*1.25 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*1.25 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*1 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*1 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "purple", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*0.75 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.75 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*0.50 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.50 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "orange", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*0.25 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.25 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "green", alpha = 0.2) +
    
    #LE lines
    geom_vline(xintercept = life_exp$life_exp_wk *1.25, colour = "black") +
    geom_vline(xintercept = life_exp$life_exp_wk, colour = "purple") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.75, colour = "red") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.5, colour = "orange") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.25, colour = "green") +
    
    #trendlines 
    geom_smooth(aes_string(group = by, linetype = "solid"), method = "lm", formula = y ~ poly(x, 3), 
                linetype = "solid", color = "black", size = 1.8, se = FALSE) +
    geom_smooth(aes(linetype = "solid"), method = "lm", formula = y ~ poly(x, 3), size = 1, se = FALSE) + 
    
    #labels
    xlab(xlabel) + 
    ylab(ylabel) +
    xlim(xlow, xhigh) + 
    
    #second axis
    scale_x_continuous(limits = c(xlow, xhigh), sec.axis = sec_axis(~.*1, breaks = LE_breaks, labels = LE_labels, name="Percent Life Expectancy")) + 
    theme(axis.text.x.top = element_text(angle = 0, margin = margin(t=-2), size = 7, face = "bold", color = "black", vjust = -5), 
        axis.title.x.top = element_text(size = 10, vjust = -1, margin = margin(t=-2)),
        axis.line.x.top = element_line(size = 0, colour = "white"), 
        axis.ticks.x.top = element_blank()) + 
    scale_linetype_manual(labels = c("All"), name = "", values = c("solid", "dashed", "solid"), guide = FALSE) + 
    scale_color_manual(values = surv_color, name = "Survival \n(% Life Expectancy)")

} 

``` 
 
```{r nhp_ss_dist_plots, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

cbPalette_dot <- c("#f8766d", "#00b0f6", "black") 

#sex
nhp_fig_i1 <- fcn_plotdist_smooth(metric = "gluc", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "gluc", ylabel = "Glucose (mg/dL)", by = "sex", peak1 = gluc_peak_male, peak2 = gluc_peak_female) + 
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)
nhp_fig_i2 <- fcn_plotdist_smooth(metric = "bw", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "bw", ylabel = "Body Weight (kg)", by = "sex", peak1 = bw_peak_male, peak2 = bw_peak_female) + 
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)
nhp_fig_i3 <- fcn_plotdist_smooth(metric = "fat", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "fat", ylabel = "Body Fat (kg)", by = "sex", peak1 = fat_peak_male, peak2 = fat_peak_female) + 
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)
#strain
nhp_fig_j1 <- fcn_plotdist_smooth(metric = "gluc", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "gluc", ylabel = "Glucose (mg/dL)", by = "strain", peak1 = gluc_peak_nia, peak2 = gluc_peak_wis) + 
  scale_color_manual(labels = c("NIA", "WIS"), name = "Colony", values = cbPalette_dot)
nhp_fig_j2 <- fcn_plotdist_smooth(metric = "bw", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "bw", ylabel = "Body Weight (kg)", by = "strain", peak1 = bw_peak_nia, peak2 = bw_peak_wis) + 
  scale_color_manual(labels = c("NIA", "WIS"), name = "Colony", values = cbPalette_dot)
nhp_fig_j3 <- fcn_plotdist_smooth(metric = "fat", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "fat", ylabel = "Body Fat (kg)", by = "strain", peak1 = fat_peak_nia, peak2 = fat_peak_wis) + 
  scale_color_manual(labels = c("NIA", "WIS"), name = "Colony", values = cbPalette_dot)
#diabetes
nhp_fig_diab1 <- fcn_plotdist_smooth(metric = "gluc", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "gluc", ylabel = "Glucose (mg/dL)", by = "diabetes2", peak1 = gluc_peak_diab, peak2 = gluc_peak_nondiab) + 
  scale_color_manual(labels = c("Diabetic", "Non-Diabetic"), name = "Health Status", values = cbPalette_dot)
nhp_fig_diab2 <- fcn_plotdist_smooth(metric = "bw", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "bw", ylabel = "Body Weight (kg)", by = "diabetes2", peak1 = bw_peak_diab, peak2 = bw_peak_nondiab) + 
  scale_color_manual(labels = c("Diabetic", "Non-Diabetic"), name = "Health Status", values = cbPalette_dot)
nhp_fig_diab3 <- fcn_plotdist_smooth(metric = "fat", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "fat", ylabel = "Body Fat (kg)", by = "diabetes2", peak1 = fat_peak_diab, peak2 = fat_peak_nondiab) + 
  scale_color_manual(labels = c("Diabetic", "Non-Diabetic"), name = "Health Status", values = cbPalette_dot)
#death
nhp_fig_d1 <- fcn_plotdist_lm(metric = "gluc", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "gluc", ylabel = "Glucose (mg/dL)", by = "whendead")
nhp_fig_d2 <- fcn_plotdist_lm(metric = "bw", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "bw", ylabel = "Body Weight (kg)", by = "whendead") 
nhp_fig_d3 <- fcn_plotdist_lm(metric = "fat", x = "age_wk", xlabel = "Age (months)", xlow = 0, xhigh = 445, 
                        y = "fat", ylabel = "Body Fat (kg)", by = "whendead") 

## 
nhp_fig_i13 <- ggarrange(nhp_fig_i1, nhp_fig_i2, nhp_fig_i3,
                      ncol = 2, nrow = 2,
                     common.legend = TRUE,
                     legend = "right")

nhp_fig_j13 <- ggarrange(nhp_fig_j1, nhp_fig_j2, nhp_fig_j3,
                      ncol = 2, nrow = 2,
                     common.legend = TRUE,
                     legend = "right")

nhp_fig_diab13 <- ggarrange(nhp_fig_diab1, nhp_fig_diab2, nhp_fig_diab3,
                      ncol = 2, nrow = 2,
                     common.legend = TRUE,
                     legend = "right")

nhp_fig_d13 <- ggarrange(nhp_fig_d1, nhp_fig_d2, nhp_fig_d3,
                      ncol = 2, nrow = 2,
                     common.legend = TRUE,
                     legend = "right")

``` 

```{r nhp_sex_dist_fig, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

annotate_figure(nhp_fig_i13, 
                top = text_grob("Trends in Metabolic indices over the Lifespan by Sex"), 
                bottom = text_grob("Note: \nHorizontal lines indicate percent of life expectancy: green (25%), orange (50%), red (75%), purple (100%), and black (125%) \nDiamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines")) 
                ) 
``` 

*** 

```{r nhp_strain_dist_fig, warning=FALSE, message=FALSE, echo=FALSE}

annotate_figure(nhp_fig_j13, 
                top = text_grob("Trends in Metabolic indices over the Lifespan by Colony"), 
                bottom = text_grob("Note: \nHorizontal lines indicate percent of life expectancy: green (25%), orange (50%), red (75%), purple (100%), and black (125%) \nDiamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines")) 
                ) 
``` 

*** 

```{r nhp_diab_dist_fig, warning=FALSE, message=FALSE, echo=FALSE}

annotate_figure(nhp_fig_diab13, 
                top = text_grob("Trends in Metabolic indices over the Lifespan by Health Status"), 
                bottom = text_grob("Note: \nHorizontal lines indicate percent of life expectancy: green (25%), orange (50%), red (75%), purple (100%), and black (125%) \nDiamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines")) 
                ) 
``` 

*** 

#### pre and post-peak change creation 

```{r nhp_phase_creation, warning=FALSE, message=FALSE, echo=FALSE}

## assuming peak indices value is fixed (sd = 0) for rate standard deviation calculations

all_gluc <- NULL
f_gluc <- NULL
m_gluc <- NULL

#all
all_gluc$pretime <- (gluc_peak$all$le - gluc_peak$lowbound$le) 
all_gluc$prepeak_diff <- (gluc_peak$all$fit - gluc_peak$lowbound$fit)
all_gluc$prepeak_rate <- ((all_gluc$prepeak_diff / gluc_peak$lowbound$fit)/all_gluc$pretime)*5*100
all_gluc$prepeak_diff_lcl <- all_gluc$prepeak_diff - 1.96*gluc_peak$lowbound$se.fit
all_gluc$prepeak_diff_ucl <- all_gluc$prepeak_diff + 1.96*gluc_peak$lowbound$se.fit
all_gluc$prepeak_rate_lcl <- ((all_gluc$prepeak_diff_lcl / gluc_peak$lowbound$fit)/all_gluc$pretime)*5*100
all_gluc$prepeak_rate_ucl <- ((all_gluc$prepeak_diff_ucl / gluc_peak$lowbound$fit)/all_gluc$pretime)*5*100
all_gluc$prepeak_rate_se <- (all_gluc$prepeak_rate_ucl - all_gluc$prepeak_rate)/1.96
all_gluc$prepeak_rate_n <- gluc_peak$lowbound$n
all_gluc$prepeak_rate_sd <- all_gluc$prepeak_rate_se*sqrt(all_gluc$prepeak_rate_n) 

#female
f_gluc$pretime <- (gluc_peak_female$all$le - gluc_peak_female$lowbound$le) 
f_gluc$prepeak_diff <- (gluc_peak_female$all$fit - gluc_peak_female$lowbound$fit)
f_gluc$prepeak_rate <- ((f_gluc$prepeak_diff / gluc_peak_female$lowbound$fit)/f_gluc$pretime)*5*100
f_gluc$prepeak_diff_lcl <- f_gluc$prepeak_diff - 1.96*gluc_peak_female$lowbound$se.fit
f_gluc$prepeak_diff_ucl <- f_gluc$prepeak_diff + 1.96*gluc_peak_female$lowbound$se.fit
f_gluc$prepeak_rate_lcl <- ((f_gluc$prepeak_diff_lcl / gluc_peak_female$lowbound$fit)/f_gluc$pretime)*5*100
f_gluc$prepeak_rate_ucl <- ((f_gluc$prepeak_diff_ucl / gluc_peak_female$lowbound$fit)/f_gluc$pretime)*5*100
f_gluc$prepeak_rate_se <- (f_gluc$prepeak_rate_ucl - f_gluc$prepeak_rate)/1.96
f_gluc$prepeak_rate_n <- gluc_peak_female$lowbound$n
f_gluc$prepeak_rate_sd <- f_gluc$prepeak_rate_se*sqrt(f_gluc$prepeak_rate_n) 

#male
m_gluc$pretime <- (gluc_peak_male$all$le - gluc_peak_male$lowbound$le) 
m_gluc$prepeak_diff <- (gluc_peak_male$all$fit - gluc_peak_male$lowbound$fit)
m_gluc$prepeak_rate <- ((m_gluc$prepeak_diff / gluc_peak_male$lowbound$fit)/m_gluc$pretime)*5*100
m_gluc$prepeak_diff_lcl <- m_gluc$prepeak_diff - 1.96*gluc_peak_male$lowbound$se.fit
m_gluc$prepeak_diff_ucl <- m_gluc$prepeak_diff + 1.96*gluc_peak_male$lowbound$se.fit
m_gluc$prepeak_rate_lcl <- ((m_gluc$prepeak_diff_lcl / gluc_peak_male$lowbound$fit)/m_gluc$pretime)*5*100
m_gluc$prepeak_rate_ucl <- ((m_gluc$prepeak_diff_ucl / gluc_peak_male$lowbound$fit)/m_gluc$pretime)*5*100
m_gluc$prepeak_rate_se <- (m_gluc$prepeak_rate_ucl - m_gluc$prepeak_rate)/1.96
m_gluc$prepeak_rate_n <- gluc_peak_male$lowbound$n
m_gluc$prepeak_rate_sd <- m_gluc$prepeak_rate_se*sqrt(m_gluc$prepeak_rate_n) 

#all
all_gluc$posttime <- (gluc_peak$highbound$le - gluc_peak$all$le) 
all_gluc$postpeak_diff <- (gluc_peak$highbound$fit - gluc_peak$all$fit)
all_gluc$postpeak_rate <- ((all_gluc$postpeak_diff / gluc_peak$all$fit)/all_gluc$posttime)*5*100
all_gluc$postpeak_diff_lcl <- all_gluc$postpeak_diff - 1.96*gluc_peak$highbound$se.fit
all_gluc$postpeak_diff_ucl <- all_gluc$postpeak_diff + 1.96*gluc_peak$highbound$se.fit
all_gluc$postpeak_rate_lcl <- ((all_gluc$postpeak_diff_lcl / gluc_peak$all$fit)/all_gluc$posttime)*5*100
all_gluc$postpeak_rate_ucl <- ((all_gluc$postpeak_diff_ucl / gluc_peak$all$fit)/all_gluc$posttime)*5*100
all_gluc$postpeak_rate_se <- (all_gluc$postpeak_rate_ucl - all_gluc$postpeak_rate)/1.96
all_gluc$postpeak_rate_n <- gluc_peak$highbound$n 
all_gluc$postpeak_rate_sd <- all_gluc$postpeak_rate_se*sqrt(all_gluc$postpeak_rate_n) 

#female
f_gluc$posttime <- (gluc_peak_female$highbound$le - gluc_peak_female$all$le) 
f_gluc$postpeak_diff <- (gluc_peak_female$highbound$fit - gluc_peak_female$all$fit) 
f_gluc$postpeak_rate <- ((f_gluc$postpeak_diff / gluc_peak_female$all$fit)/f_gluc$posttime)*5*100 
f_gluc$postpeak_diff_lcl <- f_gluc$postpeak_diff - 1.96*gluc_peak_female$highbound$se.fit 
f_gluc$postpeak_diff_ucl <- f_gluc$postpeak_diff + 1.96*gluc_peak_female$highbound$se.fit 
f_gluc$postpeak_rate_lcl <- ((f_gluc$postpeak_diff_lcl / gluc_peak_female$all$fit)/f_gluc$posttime)*5*100 
f_gluc$postpeak_rate_ucl <- ((f_gluc$postpeak_diff_ucl / gluc_peak_female$all$fit)/f_gluc$posttime)*5*100 
f_gluc$postpeak_rate_se <- (f_gluc$postpeak_rate_ucl - f_gluc$postpeak_rate)/1.96
f_gluc$postpeak_rate_n <- gluc_peak_female$highbound$n 
f_gluc$postpeak_rate_sd <- f_gluc$postpeak_rate_se*sqrt(f_gluc$postpeak_rate_n) 

#male
m_gluc$posttime <- (gluc_peak_male$highbound$le - gluc_peak_male$all$le) 
m_gluc$postpeak_diff <- (gluc_peak_male$highbound$fit - gluc_peak_male$all$fit) 
m_gluc$postpeak_rate <- ((m_gluc$postpeak_diff / gluc_peak_male$all$fit)/m_gluc$posttime)*5*100 
m_gluc$postpeak_diff_lcl <- m_gluc$postpeak_diff - 1.96*gluc_peak_male$highbound$se.fit 
m_gluc$postpeak_diff_ucl <- m_gluc$postpeak_diff + 1.96*gluc_peak_male$highbound$se.fit 
m_gluc$postpeak_rate_lcl <- ((m_gluc$postpeak_diff_lcl / gluc_peak_male$all$fit)/m_gluc$posttime)*5*100 
m_gluc$postpeak_rate_ucl <- ((m_gluc$postpeak_diff_ucl / gluc_peak_male$all$fit)/m_gluc$posttime)*5*100 
m_gluc$postpeak_rate_se <- (m_gluc$postpeak_rate_ucl - m_gluc$postpeak_rate)/1.96
m_gluc$postpeak_rate_n <- gluc_peak_male$highbound$n 
m_gluc$postpeak_rate_sd <- m_gluc$postpeak_rate_se*sqrt(m_gluc$postpeak_rate_n) 

gluc2 <- as.data.frame(bind_rows(all_gluc, f_gluc, m_gluc)) %>% 
  mutate(group = c("all", "female", "male"), 
         index = c("gluc", "gluc", "gluc")) 

all_bw <- NULL
f_bw <- NULL
m_bw <- NULL

#all
all_bw$pretime <- (bw_peak$all$le - bw_peak$lowbound$le) 
all_bw$prepeak_diff <- (bw_peak$all$fit - bw_peak$lowbound$fit)
all_bw$prepeak_rate <- ((all_bw$prepeak_diff / bw_peak$lowbound$fit)/all_bw$pretime)*5*100
all_bw$prepeak_diff_lcl <- all_bw$prepeak_diff - 1.96*bw_peak$lowbound$se.fit
all_bw$prepeak_diff_ucl <- all_bw$prepeak_diff + 1.96*bw_peak$lowbound$se.fit

all_bw$prepeak_rate_lcl <- ((all_bw$prepeak_diff_lcl / bw_peak$lowbound$fit)/all_bw$pretime)*5*100
all_bw$prepeak_rate_ucl <- ((all_bw$prepeak_diff_ucl / bw_peak$lowbound$fit)/all_bw$pretime)*5*100

all_bw$prepeak_rate_se <- (all_bw$prepeak_rate_ucl - all_bw$prepeak_rate)/1.96
all_bw$prepeak_rate_n <- bw_peak$lowbound$n
all_bw$prepeak_rate_sd <- all_bw$prepeak_rate_se*sqrt(all_bw$prepeak_rate_n) 

#female
f_bw$pretime <- (bw_peak_female$all$le - bw_peak_female$lowbound$le) 
f_bw$prepeak_diff <- (bw_peak_female$all$fit - bw_peak_female$lowbound$fit)
f_bw$prepeak_rate <- ((f_bw$prepeak_diff / bw_peak_female$lowbound$fit)/f_bw$pretime)*5*100
f_bw$prepeak_diff_lcl <- f_bw$prepeak_diff - 1.96*bw_peak_female$lowbound$se.fit
f_bw$prepeak_diff_ucl <- f_bw$prepeak_diff + 1.96*bw_peak_female$lowbound$se.fit
f_bw$prepeak_rate_lcl <- ((f_bw$prepeak_diff_lcl / bw_peak_female$lowbound$fit)/f_bw$pretime)*5*100
f_bw$prepeak_rate_ucl <- ((f_bw$prepeak_diff_ucl / bw_peak_female$lowbound$fit)/f_bw$pretime)*5*100
f_bw$prepeak_rate_se <- (f_bw$prepeak_rate_ucl - f_bw$prepeak_rate)/1.96
f_bw$prepeak_rate_n <- bw_peak_female$lowbound$n
f_bw$prepeak_rate_sd <- f_bw$prepeak_rate_se*sqrt(f_bw$prepeak_rate_n) 

#male
m_bw$pretime <- (bw_peak_male$all$le - bw_peak_male$lowbound$le) 
m_bw$prepeak_diff <- (bw_peak_male$all$fit - bw_peak_male$lowbound$fit)
m_bw$prepeak_rate <- ((m_bw$prepeak_diff / bw_peak_male$lowbound$fit)/m_bw$pretime)*5*100
m_bw$prepeak_diff_lcl <- m_bw$prepeak_diff - 1.96*bw_peak_male$lowbound$se.fit
m_bw$prepeak_diff_ucl <- m_bw$prepeak_diff + 1.96*bw_peak_male$lowbound$se.fit
m_bw$prepeak_rate_lcl <- ((m_bw$prepeak_diff_lcl / bw_peak_male$lowbound$fit)/m_bw$pretime)*5*100
m_bw$prepeak_rate_ucl <- ((m_bw$prepeak_diff_ucl / bw_peak_male$lowbound$fit)/m_bw$pretime)*5*100
m_bw$prepeak_rate_se <- (m_bw$prepeak_rate_ucl - m_bw$prepeak_rate)/1.96
m_bw$prepeak_rate_n <- bw_peak_male$lowbound$n
m_bw$prepeak_rate_sd <- m_bw$prepeak_rate_se*sqrt(m_bw$prepeak_rate_n) 

#all
all_bw$posttime <- (bw_peak$highbound$le - bw_peak$all$le) 
all_bw$postpeak_diff <- (bw_peak$highbound$fit - bw_peak$all$fit)
all_bw$postpeak_rate <- ((all_bw$postpeak_diff / bw_peak$all$fit)/all_bw$posttime)*5*100
all_bw$postpeak_diff_lcl <- all_bw$postpeak_diff - 1.96*bw_peak$highbound$se.fit
all_bw$postpeak_diff_ucl <- all_bw$postpeak_diff + 1.96*bw_peak$highbound$se.fit
all_bw$postpeak_rate_lcl <- ((all_bw$postpeak_diff_lcl / bw_peak$all$fit)/all_bw$posttime)*5*100
all_bw$postpeak_rate_ucl <- ((all_bw$postpeak_diff_ucl / bw_peak$all$fit)/all_bw$posttime)*5*100
all_bw$postpeak_rate_se <- (all_bw$postpeak_rate_ucl - all_bw$postpeak_rate)/1.96
all_bw$postpeak_rate_n <- bw_peak$highbound$n 
all_bw$postpeak_rate_sd <- all_bw$postpeak_rate_se*sqrt(all_bw$postpeak_rate_n) 

#female
f_bw$posttime <- (bw_peak_female$highbound$le - bw_peak_female$all$le) 
f_bw$postpeak_diff <- (bw_peak_female$highbound$fit - bw_peak_female$all$fit) 
f_bw$postpeak_rate <- ((f_bw$postpeak_diff / bw_peak_female$all$fit)/f_bw$posttime)*5*100 
f_bw$postpeak_diff_lcl <- f_bw$postpeak_diff - 1.96*bw_peak_female$highbound$se.fit 
f_bw$postpeak_diff_ucl <- f_bw$postpeak_diff + 1.96*bw_peak_female$highbound$se.fit 
f_bw$postpeak_rate_lcl <- ((f_bw$postpeak_diff_lcl / bw_peak_female$all$fit)/f_bw$posttime)*5*100 
f_bw$postpeak_rate_ucl <- ((f_bw$postpeak_diff_ucl / bw_peak_female$all$fit)/f_bw$posttime)*5*100 
f_bw$postpeak_rate_se <- (f_bw$postpeak_rate_ucl - f_bw$postpeak_rate)/1.96
f_bw$postpeak_rate_n <- bw_peak_female$highbound$n 
f_bw$postpeak_rate_sd <- f_bw$postpeak_rate_se*sqrt(f_bw$postpeak_rate_n) 

#male
m_bw$posttime <- (bw_peak_male$highbound$le - bw_peak_male$all$le) 
m_bw$postpeak_diff <- (bw_peak_male$highbound$fit - bw_peak_male$all$fit) 
m_bw$postpeak_rate <- ((m_bw$postpeak_diff / bw_peak_male$all$fit)/m_bw$posttime)*5*100 
m_bw$postpeak_diff_lcl <- m_bw$postpeak_diff - 1.96*bw_peak_male$highbound$se.fit 
m_bw$postpeak_diff_ucl <- m_bw$postpeak_diff + 1.96*bw_peak_male$highbound$se.fit 
m_bw$postpeak_rate_lcl <- ((m_bw$postpeak_diff_lcl / bw_peak_male$all$fit)/m_bw$posttime)*5*100 
m_bw$postpeak_rate_ucl <- ((m_bw$postpeak_diff_ucl / bw_peak_male$all$fit)/m_bw$posttime)*5*100 
m_bw$postpeak_rate_se <- (m_bw$postpeak_rate_ucl - m_bw$postpeak_rate)/1.96
m_bw$postpeak_rate_n <- bw_peak_male$highbound$n 
m_bw$postpeak_rate_sd <- m_bw$postpeak_rate_se*sqrt(m_bw$postpeak_rate_n) 

bw2 <- as.data.frame(bind_rows(all_bw, f_bw, m_bw)) %>% 
  mutate(group = c("all", "female", "male"), 
         index = c("bw", "bw", "bw")) 

all_fat <- NULL
f_fat <- NULL
m_fat <- NULL

#all
all_fat$pretime <- (fat_peak$all$le - fat_peak$lowbound$le) 
all_fat$prepeak_diff <- (fat_peak$all$fit - fat_peak$lowbound$fit)
all_fat$prepeak_rate <- ((all_fat$prepeak_diff / fat_peak$lowbound$fit)/all_fat$pretime)*5*100
all_fat$prepeak_diff_lcl <- all_fat$prepeak_diff - 1.96*fat_peak$lowbound$se.fit
all_fat$prepeak_diff_ucl <- all_fat$prepeak_diff + 1.96*fat_peak$lowbound$se.fit

all_fat$prepeak_rate_lcl <- ((all_fat$prepeak_diff_lcl / fat_peak$lowbound$fit)/all_fat$pretime)*5*100
all_fat$prepeak_rate_ucl <- ((all_fat$prepeak_diff_ucl / fat_peak$lowbound$fit)/all_fat$pretime)*5*100

all_fat$prepeak_rate_se <- (all_fat$prepeak_rate_ucl - all_fat$prepeak_rate)/1.96
all_fat$prepeak_rate_n <- fat_peak$lowbound$n
all_fat$prepeak_rate_sd <- all_fat$prepeak_rate_se*sqrt(all_fat$prepeak_rate_n) 

#female
f_fat$pretime <- (fat_peak_female$all$le - fat_peak_female$lowbound$le) 
f_fat$prepeak_diff <- (fat_peak_female$all$fit - fat_peak_female$lowbound$fit)
f_fat$prepeak_rate <- ((f_fat$prepeak_diff / fat_peak_female$lowbound$fit)/f_fat$pretime)*5*100
f_fat$prepeak_diff_lcl <- f_fat$prepeak_diff - 1.96*fat_peak_female$lowbound$se.fit
f_fat$prepeak_diff_ucl <- f_fat$prepeak_diff + 1.96*fat_peak_female$lowbound$se.fit
f_fat$prepeak_rate_lcl <- ((f_fat$prepeak_diff_lcl / fat_peak_female$lowbound$fit)/f_fat$pretime)*5*100
f_fat$prepeak_rate_ucl <- ((f_fat$prepeak_diff_ucl / fat_peak_female$lowbound$fit)/f_fat$pretime)*5*100
f_fat$prepeak_rate_se <- (f_fat$prepeak_rate_ucl - f_fat$prepeak_rate)/1.96
f_fat$prepeak_rate_n <- fat_peak_female$lowbound$n
f_fat$prepeak_rate_sd <- f_fat$prepeak_rate_se*sqrt(f_fat$prepeak_rate_n) 

#male
m_fat$pretime <- (fat_peak_male$all$le - fat_peak_male$lowbound$le) 
m_fat$prepeak_diff <- (fat_peak_male$all$fit - fat_peak_male$lowbound$fit)
m_fat$prepeak_rate <- ((m_fat$prepeak_diff / fat_peak_male$lowbound$fit)/m_fat$pretime)*5*100
m_fat$prepeak_diff_lcl <- m_fat$prepeak_diff - 1.96*fat_peak_male$lowbound$se.fit
m_fat$prepeak_diff_ucl <- m_fat$prepeak_diff + 1.96*fat_peak_male$lowbound$se.fit
m_fat$prepeak_rate_lcl <- ((m_fat$prepeak_diff_lcl / fat_peak_male$lowbound$fit)/m_fat$pretime)*5*100
m_fat$prepeak_rate_ucl <- ((m_fat$prepeak_diff_ucl / fat_peak_male$lowbound$fit)/m_fat$pretime)*5*100
m_fat$prepeak_rate_se <- (m_fat$prepeak_rate_ucl - m_fat$prepeak_rate)/1.96
m_fat$prepeak_rate_n <- fat_peak_male$lowbound$n
m_fat$prepeak_rate_sd <- m_fat$prepeak_rate_se*sqrt(m_fat$prepeak_rate_n) 

#all
all_fat$posttime <- (fat_peak$highbound$le - fat_peak$all$le) 
all_fat$postpeak_diff <- (fat_peak$highbound$fit - fat_peak$all$fit)
all_fat$postpeak_rate <- ((all_fat$postpeak_diff / fat_peak$all$fit)/all_fat$posttime)*5*100
all_fat$postpeak_diff_lcl <- all_fat$postpeak_diff - 1.96*fat_peak$highbound$se.fit
all_fat$postpeak_diff_ucl <- all_fat$postpeak_diff + 1.96*fat_peak$highbound$se.fit
all_fat$postpeak_rate_lcl <- ((all_fat$postpeak_diff_lcl / fat_peak$all$fit)/all_fat$posttime)*5*100
all_fat$postpeak_rate_ucl <- ((all_fat$postpeak_diff_ucl / fat_peak$all$fit)/all_fat$posttime)*5*100
all_fat$postpeak_rate_se <- (all_fat$postpeak_rate_ucl - all_fat$postpeak_rate)/1.96
all_fat$postpeak_rate_n <- fat_peak$highbound$n 
all_fat$postpeak_rate_sd <- all_fat$postpeak_rate_se*sqrt(all_fat$postpeak_rate_n) 

#female
f_fat$posttime <- (fat_peak_female$highbound$le - fat_peak_female$all$le) 
f_fat$postpeak_diff <- (fat_peak_female$highbound$fit - fat_peak_female$all$fit) 
f_fat$postpeak_rate <- ((f_fat$postpeak_diff / fat_peak_female$all$fit)/f_fat$posttime)*5*100 
f_fat$postpeak_diff_lcl <- f_fat$postpeak_diff - 1.96*fat_peak_female$highbound$se.fit 
f_fat$postpeak_diff_ucl <- f_fat$postpeak_diff + 1.96*fat_peak_female$highbound$se.fit 
f_fat$postpeak_rate_lcl <- ((f_fat$postpeak_diff_lcl / fat_peak_female$all$fit)/f_fat$posttime)*5*100 
f_fat$postpeak_rate_ucl <- ((f_fat$postpeak_diff_ucl / fat_peak_female$all$fit)/f_fat$posttime)*5*100 
f_fat$postpeak_rate_se <- (f_fat$postpeak_rate_ucl - f_fat$postpeak_rate)/1.96
f_fat$postpeak_rate_n <- fat_peak_female$highbound$n 
f_fat$postpeak_rate_sd <- f_fat$postpeak_rate_se*sqrt(f_fat$postpeak_rate_n) 

#male
m_fat$posttime <- (fat_peak_male$highbound$le - fat_peak_male$all$le) 
m_fat$postpeak_diff <- (fat_peak_male$highbound$fit - fat_peak_male$all$fit) 
m_fat$postpeak_rate <- ((m_fat$postpeak_diff / fat_peak_male$all$fit)/m_fat$posttime)*5*100 
m_fat$postpeak_diff_lcl <- m_fat$postpeak_diff - 1.96*fat_peak_male$highbound$se.fit 
m_fat$postpeak_diff_ucl <- m_fat$postpeak_diff + 1.96*fat_peak_male$highbound$se.fit 
m_fat$postpeak_rate_lcl <- ((m_fat$postpeak_diff_lcl / fat_peak_male$all$fit)/m_fat$posttime)*5*100 
m_fat$postpeak_rate_ucl <- ((m_fat$postpeak_diff_ucl / fat_peak_male$all$fit)/m_fat$posttime)*5*100 
m_fat$postpeak_rate_se <- (m_fat$postpeak_rate_ucl - m_fat$postpeak_rate)/1.96
m_fat$postpeak_rate_n <- fat_peak_male$highbound$n 
m_fat$postpeak_rate_sd <- m_fat$postpeak_rate_se*sqrt(m_fat$postpeak_rate_n) 

fat2 <- as.data.frame(bind_rows(all_fat, f_fat, m_fat)) %>% 
  mutate(group = c("all", "female", "male"), 
         index = c("fat", "fat", "fat")) 


##
phases <- rbind(gluc2, bw2, fat2) %>% 
  mutate(
    
    prepeak_pval = tsum.test(prepeak_rate, prepeak_rate_sd, prepeak_rate_n, mu = 0, conf.level = 0.95)$p.value,
    postpeak_pval = tsum.test(postpeak_rate, postpeak_rate_sd, postpeak_rate_n, mu = 0, conf.level = 0.95)$p.value,
    
    prepeak_mean_95CI =
      
      ifelse(is.na(prepeak_pval),
    paste0(format(round(prepeak_rate, 2), nsmall=2)," ","(",format(round(prepeak_rate_lcl, 2), nsmall=2),", ",format(round(prepeak_rate_ucl, 2), nsmall=2),")"),
      ifelse(prepeak_pval < 0.050,
    paste0(format(round(prepeak_rate, 2), nsmall=2)," ","(",format(round(prepeak_rate_lcl, 2), nsmall=2),", ",format(round(prepeak_rate_ucl, 2), nsmall=2),")*"),
    paste0(format(round(prepeak_rate, 2), nsmall=2)," ","(",format(round(prepeak_rate_lcl, 2), nsmall=2),", ",format(round(prepeak_rate_ucl, 2), nsmall=2),")"))),

    postpeak_mean_95CI = 
      ifelse(is.na(postpeak_pval),
    paste0(format(round(postpeak_rate, 2), nsmall=2)," ","(",format(round(postpeak_rate_lcl, 2), nsmall=2),", ",format(round(postpeak_rate_ucl, 2), nsmall=2),")"),
      ifelse(postpeak_pval < 0.050,
    paste0(format(round(postpeak_rate, 2), nsmall=2)," ","(",format(round(postpeak_rate_lcl, 2), nsmall=2),", ",format(round(postpeak_rate_ucl, 2), nsmall=2),")*"),
    paste0(format(round(postpeak_rate, 2), nsmall=2)," ","(",format(round(postpeak_rate_lcl, 2), nsmall=2),", ",format(round(postpeak_rate_ucl, 2), nsmall=2),")")))
    
  )

gluc_prepeak <- phases %>% 
  filter(index == "gluc") %>% 
  select(index, group, contains("95"), -contains("postpeak")) %>% 
  spread(key = group, value = prepeak_mean_95CI) 
gluc_postpeak <- phases %>% 
  filter(index == "gluc") %>% 
  select(index, group, contains("95"), -contains("prepeak")) %>% 
  spread(key = group, value = postpeak_mean_95CI) 

bw_prepeak <- phases %>% 
  filter(index == "bw") %>% 
  select(index, group, contains("95"), -contains("postpeak")) %>% 
  spread(key = group, value = prepeak_mean_95CI) 
bw_postpeak <- phases %>% 
  filter(index == "bw") %>% 
  select(index, group, contains("95"), -contains("prepeak")) %>% 
  spread(key = group, value = postpeak_mean_95CI) 

fat_prepeak <- phases %>% 
  filter(index == "fat") %>% 
  select(index, group, contains("95"), -contains("postpeak")) %>% 
  spread(key = group, value = prepeak_mean_95CI) 
fat_postpeak <- phases %>% 
  filter(index == "fat") %>% 
  select(index, group, contains("95"), -contains("prepeak")) %>% 
  spread(key = group, value = postpeak_mean_95CI) 

gluc_all <- bind_cols(gluc_prepeak, gluc_postpeak)
bw_all <- bind_cols(bw_prepeak, bw_postpeak)
fat_all <- bind_cols(fat_prepeak, fat_postpeak)

phases2 <- bind_rows(gluc_all, bw_all, fat_all) 
colnames(phases2) <- c("index", "all", "female", "male", "index2", "all2", "female2", "male2")

##peaks

#all
gluc_peak_ci <- paste0(round(gluc_peak$all$le, 2)," (", round(gluc_peak$min$le, 2), ", ", round(gluc_peak$max$le, 2), ")") 
bw_peak_ci <- paste0(round(bw_peak$all$le, 2)," (", round(bw_peak$min$le, 2), ", ", round(bw_peak$max$le, 2), ")") 
fat_peak_ci <- paste0(round(fat_peak$all$le, 2)," (", round(fat_peak$min$le, 2), ", ", round(fat_peak$max$le, 2), ")") 

#male
male_gluc_peak_ci <- paste0(round(gluc_peak_male$all$le, 2)," (", round(gluc_peak_male$min$le, 2), ", ", round(gluc_peak_male$max$le, 2), ")") 
male_bw_peak_ci <- paste0(round(bw_peak_male$all$le, 2)," (", round(bw_peak_male$min$le, 2), ", ", round(bw_peak_male$max$le, 2), ")") 
male_fat_peak_ci <- paste0(round(fat_peak_male$all$le, 2)," (", round(fat_peak_male$min$le, 2), ", ", round(fat_peak_male$max$le, 2), ")") 

#female
female_gluc_peak_ci <- paste0(round(gluc_peak_female$all$le, 2)," (", round(gluc_peak_female$min$le, 2), ", ", round(gluc_peak_female$max$le, 2), ")") 
female_bw_peak_ci <- paste0(round(bw_peak_female$all$le, 2)," (", round(bw_peak_female$min$le, 2), ", ", round(bw_peak_female$max$le, 2), ")") 
female_fat_peak_ci <- paste0(round(fat_peak_female$all$le, 2)," (", round(fat_peak_female$min$le, 2), ", ", round(fat_peak_female$max$le, 2), ")") 

#final datasets
phases2$peak_all <- c(gluc_peak_ci, bw_peak_ci, fat_peak_ci) 
phases2$peak_female <- c(female_gluc_peak_ci, female_bw_peak_ci, female_fat_peak_ci) 
phases2$peak_male <- c(male_gluc_peak_ci, male_bw_peak_ci, male_fat_peak_ci)

phases3 <- phases2 %>% 
  select(index, contains("peak"), everything(), -index2) 

```

```{r nhp_phase_formatting, warning=FALSE, message=FALSE, echo=FALSE}

#html output formatting
phases4 <- phases3 %>% 
  mutate( 
    peak_all = noquote((gsub("\\(", "<br>(", peak_all))), 
    peak_female = noquote((gsub("\\(", "<br>(", peak_female))), 
    peak_male = noquote((gsub("\\(", "<br>(", peak_male))), 
    all = noquote((gsub("\\(", "<br>(", all))), 
    female = noquote((gsub("\\(", "<br>(", female))), 
    male = noquote((gsub("\\(", "<br>(", male))), 
    all2 = noquote((gsub("\\(", "<br>(", all2))), 
    female2 = noquote((gsub("\\(", "<br>(", female2))), 
    male2 = noquote((gsub("\\(", "<br>(", male2)))
    ) 

nhp_phases <- phases4 

```

#### change in metabolic indices pre and post-peak 
- rate of percent change 

```{r nhp_phase_table_prate, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

nhp_phases_out <- kable(nhp_phases, row.names = FALSE, col.names = c("Metabolic <br> Index", "All", "Female", "Male", "All", "Female", "Male", "All", "Female", "Male"), align = c("c","c","c","c","c","c","c"), escape = FALSE, 
      caption = "Rate of Percent Change for Metabolic indices Pre- and Post-peak") %>% 
  kable_styling(bootstrap_options = "striped", full_width = TRUE, position = "left", font_size = 11) %>% 
  column_spec(1, bold = TRUE, border_right = TRUE) %>% 
  column_spec(4, border_right = TRUE) %>% 
  column_spec(7, border_right = TRUE) %>% 
  add_header_above(c(" " = 1, "Age at Peak Value$^$ \n(% life expectancy)" = 3, "Pre-Peak$^$$^$" = 3, "Post-Peak$^$$^$" = 3)) %>% 
  footnote(symbol = c(
    "p value < 0.05",
    "Insufficient sample size to conduct a statistical test of significance",
    "Mean (range of ages at similar values) of age at peak index value", 
    "Mean (95% confidence interval) rate of percent change during phase per five percent change in life expectancy", 
    "Pre-peak is defined as 25% life expectancy to age at peak value; post-peak phase is defined as peak to 125% life expectancy",
    "Male and female estimates are based on sex specific life expectancy")) 

nhp_phases_out
```

*** 

#### quartile creation

```{r nhp_scaling, warning=FALSE, message=FALSE, echo=FALSE}

main_cont_surv <- main_all3 %>% select(idno, sex, strain, age_25, age_50, age_75, age_100, age_125) %>% distinct() %>% inner_join(census_c16_death) 
main_cat_surv_f <- main_cat2_f %>% inner_join(main_cont_surv) %>% filter(sex == "F") 
main_cat_surv_m <- main_cat2_m %>% inner_join(main_cont_surv) %>% filter(sex == "M") 

```
 
```{r nhp_quartiles, warning=FALSE, message=FALSE, echo=FALSE}

quart_breaks = c(0, 0.25, 0.50, 0.75, 1.0) 
quart_labs = c("Low", "Medlow", "Medhigh", "High")

## female
main_cat_surv_f$gluc_50_quart <- cut(main_cat_surv_f$cfgluc_50_m, quantile(main_cat_surv_f$cfgluc_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$gluc_75_quart <- cut(main_cat_surv_f$cfgluc_75_m, quantile(main_cat_surv_f$cfgluc_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$gluc_100_quart <- cut(main_cat_surv_f$cfgluc_100_m, quantile(main_cat_surv_f$cfgluc_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_f$bw_50_quart <- cut(main_cat_surv_f$cfbw_50_m, quantile(main_cat_surv_f$cfbw_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$bw_75_quart <- cut(main_cat_surv_f$cfbw_75_m, quantile(main_cat_surv_f$cfbw_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$bw_100_quart <- cut(main_cat_surv_f$cfbw_100_m, quantile(main_cat_surv_f$cfbw_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_f$fat_50_quart <- cut(main_cat_surv_f$cffat_50_m, quantile(main_cat_surv_f$cffat_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$fat_75_quart <- cut(main_cat_surv_f$cffat_75_m, quantile(main_cat_surv_f$cffat_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$fat_100_quart <- cut(main_cat_surv_f$cffat_100_m, quantile(main_cat_surv_f$cffat_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

## male
main_cat_surv_m$gluc_50_quart <- cut(main_cat_surv_m$cfgluc_50_m, quantile(main_cat_surv_m$cfgluc_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$gluc_75_quart <- cut(main_cat_surv_m$cfgluc_75_m, quantile(main_cat_surv_m$cfgluc_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$gluc_100_quart <- cut(main_cat_surv_m$cfgluc_100_m, quantile(main_cat_surv_m$cfgluc_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_m$bw_50_quart <- cut(main_cat_surv_m$cfbw_50_m, quantile(main_cat_surv_m$cfbw_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$bw_75_quart <- cut(main_cat_surv_m$cfbw_75_m, quantile(main_cat_surv_m$cfbw_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$bw_100_quart <- cut(main_cat_surv_m$cfbw_100_m, quantile(main_cat_surv_m$cfbw_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_m$fat_50_quart <- cut(main_cat_surv_m$cffat_50_m, quantile(main_cat_surv_m$cffat_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$fat_75_quart <- cut(main_cat_surv_m$cffat_75_m, quantile(main_cat_surv_m$cffat_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$fat_100_quart <- cut(main_cat_surv_m$cffat_100_m, quantile(main_cat_surv_m$cffat_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

#create
main_cat_surv <- main_cat_surv_f %>% 
  bind_rows(main_cat_surv_m) 

```

#### kaplan meier survival curves 
- comparing highest to lowest quartile 
- three timepoints: 50, 75, and 100% life expectancy 

```{r nhp_surv_curve, warning=FALSE, message=FALSE, echo=FALSE}

surv_color <- c("orange", "purple")

km_funct <- function(quart, exp_var, start, end, data, title, env = globalenv()){

#subset data
surv_df <- as.data.frame(data) 
surv_df_plot <- subset(data, eval(parse(text = quart)) %in% c("Low", "High"))
surv_df_plot2 <- as.data.frame(surv_df_plot) 

#change level order for legend
surv_df_plot3 <- surv_df_plot2 %>% 
  mutate_at( 
    vars(contains("quart")), 
    funs(fct_rev(.)) 
  ) 

#pull input variable for quartiles
env$quart <- quart 
env$explanatory <- NULL 
env$explanatory$var <- exp_var 

#cph model
formula_cph <- as.formula(paste0("Surv(fu_age_wk, dead_censor) ~", env$quart, "+", paste(env$explanatory$var, collapse = " + "))) 
cph <- coxph(formula_cph, data = surv_df) 
km_fit <- survfit(Surv(fu_age_wk, dead_censor) ~ eval(parse(text = quart)), data = surv_df_plot3) 

cph_summary <- summary(cph) 
value <- format(round(cph_summary$conf.int[,"exp(coef)"][3], 2), nsmall = 2)
le <- format(round(cph_summary$conf.int[,"lower .95"][3], 2), nsmall = 2)
ue <- format(round(cph_summary$conf.int[,"upper .95"][3], 2), nsmall = 2)
pval <- cph_summary$coefficients[,"Pr(>|z|)"][3]

#place asterisk for significant pval
if(pval < 0.050){
  hr <- paste0(value, " (", le, ", ", ue, ")*") 
}else{
  hr <- paste0(value, " (", le, ", ", ue, ")") 
}

km_curve <- ggsurvplot(
    fit = km_fit, 
    data = surv_df_plot3, 
    title = title,
    font.title = c(12), 
    conf.int = FALSE,
    risk.table = TRUE, 
    censor = FALSE,
    risk.table.col = "strata",
    legend.title = "Index Subgroup", 
    ggtheme = theme_bw(), 
    legend.labs = c("Highest Quartile", "Lowest Quartile"),
    palette = surv_color, 
    xlab = "Age (months)", 
    ylab = "Hazard Ratio", 
    xlim = c(start, end)
) 
surv_results <- list(km_curve, hr, pval, cph_summary) 

return(surv_results) 

} 

#Survival
km_start_50 <- nhp_life_exp$life_exp_wk*.05 
km_start_75 <- nhp_life_exp$life_exp_wk*.1
km_start_100 <- nhp_life_exp$life_exp_wk*.1

#function calls 
nhp_surv_gluc_50_all <- km_funct(quart = "gluc_50_quart", exp_var = c("age_50"), start = km_start_50, data = main_cat_surv, title = "Glucose at 50% Life Expectancy", end = 400) 
nhp_surv_gluc_75_all <- km_funct(quart = "gluc_75_quart", exp_var = c("age_75"), start = km_start_75, data = main_cat_surv, title = "Glucose at 75% Life Expectancy", end = 400) 
nhp_surv_gluc_100_all <- km_funct(quart = "gluc_100_quart", exp_var = c("age_100"), start = km_start_100, data = main_cat_surv, title = "Glucose at 100% Life Expectancy", end = 400) 

nhp_surv_bw_50_all <- km_funct(quart = "bw_50_quart", exp_var = c("age_50"), start = km_start_50, data = main_cat_surv, title = "Body Weight at 50% Life Expectancy", end = 400) 
nhp_surv_bw_75_all <- km_funct(quart = "bw_75_quart", exp_var = c("age_75"), start = km_start_75, data = main_cat_surv, title = "Body Weight at 75% Life Expectancy", end = 400) 
nhp_surv_bw_100_all <- km_funct(quart = "bw_100_quart", exp_var = c("age_100"), start = km_start_100, data = main_cat_surv, title = "Body Weight at 100% Life Expectancy", end = 400) 

nhp_surv_fat_50_all <- km_funct(quart = "fat_50_quart", exp_var = c("age_50"), start = km_start_50, data = main_cat_surv, title = "Fat at 50% Life Expectancy", end = 400) 
nhp_surv_fat_75_all <- km_funct(quart = "fat_75_quart", exp_var = c("age_75"), start = km_start_75, data = main_cat_surv, title = "Fat at 75% Life Expectancy", end = 400) 
nhp_surv_fat_100_all <- km_funct(quart = "fat_100_quart", exp_var = c("age_100"), start = km_start_100, data = main_cat_surv, title = "Fat at 100% Life Expectancy", end = 400) 

```

```{r nhp_surv_output, fig.height = 12, fig.width = 12, warning=FALSE, message=FALSE, echo=FALSE}

#add adjustment subscript
nhp_km_gluc_50 <- nhp_surv_gluc_50_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(nhp_surv_gluc_50_all[[2]]))), size = 5, parse =TRUE)
nhp_km_gluc_75 <- nhp_surv_gluc_75_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(nhp_surv_gluc_75_all[[2]]))), size = 5, parse =TRUE)
nhp_km_gluc_100 <- nhp_surv_gluc_100_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(nhp_surv_gluc_100_all[[2]]))), size = 5, parse =TRUE)

nhp_km_bw_50 <- nhp_surv_bw_50_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(nhp_surv_bw_50_all[[2]]))), size = 5, parse =TRUE)
nhp_km_bw_75 <- nhp_surv_bw_75_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(nhp_surv_bw_75_all[[2]]))), size = 5, parse =TRUE)
nhp_km_bw_100 <- nhp_surv_bw_100_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(nhp_surv_bw_100_all[[2]]))), size = 5, parse =TRUE)

nhp_km_fat_50 <- nhp_surv_fat_50_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(nhp_surv_fat_50_all[[2]]))), size = 5, parse =TRUE)
nhp_km_fat_75 <- nhp_surv_fat_75_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(nhp_surv_fat_75_all[[2]]))), size = 5, parse =TRUE)
nhp_km_fat_100 <- nhp_surv_fat_100_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(nhp_surv_fat_100_all[[2]]))), size = 5, parse =TRUE)

km_all <- ggarrange(
  nhp_km_gluc_50, nhp_km_gluc_75, nhp_km_gluc_100, nhp_km_bw_50, nhp_km_bw_75, nhp_km_bw_100, nhp_km_fat_50, nhp_km_fat_75, nhp_km_fat_100,
  common.legend = TRUE, legend = "right") 

```

```{r nhp_surv_kable, fig.height = 12, fig.width = 12, warning=FALSE, message=FALSE, echo=FALSE}

km_all_output <- annotate_figure(km_all,
                                 top = text_grob("Survival Curve and Age Adjusted Mortality Hazard Ratio Comparing Highest and Lowest Metabolic Index Quartiles \n at Varying Timepoints\n", size = 14,  hjust = 0.52) 
                ) 

km_all_output

```

*** 

#### cox proportional mortality hazard ratios 
- sex specific
- covariate adjustment
  - sex, age 
  - additional adjustment for body fat 

```{r nhp_cph_fun, warning=FALSE, message=FALSE, echo=FALSE}

cph_funct <- function(quart, exp_var, data, env = globalenv()){

surv_df <- as.data.frame(data) 

env$quart <- quart 
env$explanatory <- NULL 
env$explanatory$var <- exp_var 

formula_cph <- as.formula(paste0("Surv(fu_age_wk, dead_censor) ~", env$quart, "+", paste(env$explanatory$var, collapse = " + "))) 
cph <- coxph(formula_cph, data = surv_df) 

cph_summary <- summary(cph) 
value <- format(round(cph_summary$conf.int[,"exp(coef)"][3], 2), nsmall = 2) 
le <- format(round(cph_summary$conf.int[,"lower .95"][3], 2), nsmall = 2) 
ue <- format(round(cph_summary$conf.int[,"upper .95"][3], 2), nsmall = 2) 
pval <- cph_summary$coefficients[,"Pr(>|z|)"][3] 

if(pval < 0.050){
  hr <- paste0("HR = ", value, " (", le, ", ", ue, ")*") 
}else{
  hr <- paste0("HR = ", value, " (", le, ", ", ue, ")") 
}
 
surv_results <- list(hr, pval, cph_summary) 

return(surv_results) 

} 

#function calls 
nhp_surv_gluc_50_all2 <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50",  "sex"), data = main_cat_surv)
nhp_surv_gluc_75_all2 <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75", "sex"), data = main_cat_surv)
nhp_surv_gluc_100_all2 <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100", "sex"), data = main_cat_surv)

nhp_surv_bw_50_all2 <- cph_funct(quart = "bw_50_quart", exp_var = c("age_50", "sex"), data = main_cat_surv)
nhp_surv_bw_75_all2 <- cph_funct(quart = "bw_75_quart", exp_var = c("age_75", "sex"), data = main_cat_surv)
nhp_surv_bw_100_all2 <- cph_funct(quart = "bw_100_quart", exp_var = c("age_100", "sex"), data = main_cat_surv)

nhp_surv_fat_50_all2 <- cph_funct(quart = "fat_50_quart", exp_var = c("age_50", "sex"), data = main_cat_surv)
nhp_surv_fat_75_all2 <- cph_funct(quart = "fat_75_quart", exp_var = c("age_75", "sex"), data = main_cat_surv)
nhp_surv_fat_100_all2 <- cph_funct(quart = "fat_100_quart", exp_var = c("age_100", "sex"), data = main_cat_surv)

## by sex 
nhp_surv_gluc_50_f <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50"), data = main_cat_surv_f) 
nhp_surv_gluc_75_f <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75"), data = main_cat_surv_f) 
nhp_surv_gluc_100_f <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100"), data = main_cat_surv_f) 

nhp_surv_bw_50_f <- cph_funct(quart = "bw_50_quart", exp_var = c("age_50"), data = main_cat_surv_f) 
nhp_surv_bw_75_f <- cph_funct(quart = "bw_75_quart", exp_var = c("age_75"), data = main_cat_surv_f) 
nhp_surv_bw_100_f <- cph_funct(quart = "bw_100_quart", exp_var = c("age_100"), data = main_cat_surv_f) 

nhp_surv_fat_50_f <- cph_funct(quart = "fat_50_quart", exp_var = c("age_50"), data = main_cat_surv_f) 
nhp_surv_fat_75_f <- cph_funct(quart = "fat_75_quart", exp_var = c("age_75"), data = main_cat_surv_f) 
nhp_surv_fat_100_f <- cph_funct(quart = "fat_100_quart", exp_var = c("age_100"), data = main_cat_surv_f) 

nhp_surv_gluc_50_m <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50"), data = main_cat_surv_m)
nhp_surv_gluc_75_m <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75"), data = main_cat_surv_m) 
nhp_surv_gluc_100_m <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100"), data = main_cat_surv_m) 

nhp_surv_bw_50_m <- cph_funct(quart = "bw_50_quart", exp_var = c("age_50"), data = main_cat_surv_m)
nhp_surv_bw_75_m <- cph_funct(quart = "bw_75_quart", exp_var = c("age_75"), data = main_cat_surv_m) 
nhp_surv_bw_100_m <- cph_funct(quart = "bw_100_quart", exp_var = c("age_100"), data = main_cat_surv_m) 

nhp_surv_fat_50_m <- cph_funct(quart = "fat_50_quart", exp_var = c("age_50"), data = main_cat_surv_m)
nhp_surv_fat_75_m <- cph_funct(quart = "fat_75_quart", exp_var = c("age_75"), data = main_cat_surv_m) 
nhp_surv_fat_100_m <- cph_funct(quart = "fat_100_quart", exp_var = c("age_100"), data = main_cat_surv_m) 

## adjusted for body fat
nhp_surv_gluc_50_adj <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50", "sex", "fat_50_m"), data = main_cat_surv)
nhp_surv_gluc_75_adj <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75", "sex", "fat_75_m"), data = main_cat_surv)
nhp_surv_gluc_100_adj <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100", "sex", "fat_100_m"), data = main_cat_surv)

## by sex
nhp_surv_gluc_50_adj_f <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50", "fat_50_m"), data = main_cat_surv_f)
nhp_surv_gluc_75_adj_f <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75", "fat_75_m"), data = main_cat_surv_f)
nhp_surv_gluc_100_adj_f <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100", "fat_100_m"), data = main_cat_surv_f)

nhp_surv_gluc_50_adj_m <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50", "fat_50_m"), data = main_cat_surv_m)
nhp_surv_gluc_75_adj_m <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75", "fat_75_m"), data = main_cat_surv_m)
nhp_surv_gluc_100_adj_m <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100", "fat_100_m"), data = main_cat_surv_m)

```

```{r nhp_km_by, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE, include=TRUE}

hr_toc_by <- NULL 
hr_toc_by$indices <- c("Glucose", "Glucose", "Glucose", "Body Weight", "Body Weight", "Body Weight", "Body Composition", "Body Composition", "Body Composition") 
hr_toc_by$timepoint <- c(50, 75, 100, 50, 75, 100, 50, 75, 100) 

##all
hr_toc_by$hr1_all <- c(nhp_surv_gluc_50_all2[[1]], nhp_surv_gluc_75_all2[[1]], nhp_surv_gluc_100_all2[[1]], 
                       nhp_surv_bw_50_all2[[1]], nhp_surv_bw_75_all2[[1]], nhp_surv_bw_100_all2[[1]], 
                       nhp_surv_fat_50_all2[[1]], nhp_surv_fat_75_all2[[1]], nhp_surv_fat_100_all2[[1]]) 
hr_toc_by$hr2_all <- c(nhp_surv_gluc_50_adj[[1]], nhp_surv_gluc_75_adj[[1]], nhp_surv_gluc_100_adj[[1]], 
                       NA, NA, NA, 
                       NA, NA, NA) 
##female
hr_toc_by$hr1_f <- c(nhp_surv_gluc_50_f[[1]], nhp_surv_gluc_75_f[[1]], nhp_surv_gluc_100_f[[1]], 
                     nhp_surv_bw_50_f[[1]], nhp_surv_bw_75_f[[1]], nhp_surv_bw_100_f[[1]], 
                     nhp_surv_fat_50_f[[1]], nhp_surv_fat_75_f[[1]], nhp_surv_fat_100_f[[1]]) 
hr_toc_by$hr2_f <- c(nhp_surv_gluc_50_adj_f[[1]], nhp_surv_gluc_75_adj_f[[1]], nhp_surv_gluc_100_adj_f[[1]], 
                     NA, NA, NA,
                     NA, NA, NA) 
##male
hr_toc_by$hr1_m <- c(nhp_surv_gluc_50_m[[1]], nhp_surv_gluc_75_m[[1]], nhp_surv_gluc_100_m[[1]], 
                     nhp_surv_bw_50_m[[1]], nhp_surv_bw_75_m[[1]], nhp_surv_bw_100_m[[1]], 
                     nhp_surv_fat_50_m[[1]], nhp_surv_fat_75_m[[1]], nhp_surv_fat_100_m[[1]]) 
hr_toc_by$hr2_m <- c(nhp_surv_gluc_50_adj_m[[1]], nhp_surv_gluc_75_adj_m[[1]], nhp_surv_gluc_100_adj_m[[1]], 
                     NA, NA, NA,
                     NA, NA, NA) 

hr_toc_by <- as.data.frame(hr_toc_by) 

##
hr_toc_by2 <- hr_toc_by %>% 
  mutate_all(~gsub("HR = ", "", .)) 

hr_toc_by2 <- hr_toc_by2 %>% 
  mutate( 
    hr1_all = noquote((gsub("\\(", "<br>(", hr1_all))), 
    hr2_all = noquote((gsub("\\(", "<br>(", hr2_all))), 
    hr1_m = noquote((gsub("\\(", "<br>(", hr1_m))), 
    hr2_m = noquote((gsub("\\(", "<br>(", hr2_m))), 
    hr1_f = noquote((gsub("\\(", "<br>(", hr1_f))), 
    hr2_f = noquote((gsub("\\(", "<br>(", hr2_f))),
    ) 
hr_toc_by3 <- hr_toc_by2 %>% 
  mutate(
    hr1_all = ifelse(grepl("Inf", hr1_all), "---", hr1_all),
    hr2_all = ifelse(grepl("Inf", hr2_all), "---", hr2_all), 
    hr1_m = ifelse(grepl("Inf", hr1_m), "---", hr1_m),
    hr2_m = ifelse(grepl("Inf", hr2_m), "---", hr2_m),
    hr1_f = ifelse(grepl("Inf", hr1_f), "---", hr1_f),
    hr2_f = ifelse(grepl("Inf", hr2_f), "---", hr2_f)
    )

nhp_hr_toc_by_3a <- hr_toc_by3 %>% 
  filter(indices == "Glucose") %>% 
  select(-indices)

nhp_hr_toc_by_3b <- hr_toc_by3 %>% 
  filter(indices != "Glucose") %>% 
  select(-contains("hr2")) 

```

```{r nhp_hr_gluc_kable, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE, include=TRUE}

tablehr1 <- kable(nhp_hr_toc_by_3a, row.names = FALSE, col.names = c("Timepoint" , "Model 1", "Model 2", "Model 1", "Model 2", "Model 1", "Model 2"), escape = FALSE, align = "c", caption = "Mortality Hazard Ratios of Highest and Lowest Glucose Quartiles \n at Varying Timepoints Stratified by Sex") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left", font_size = 12) %>% 
  column_spec(1, bold = TRUE, border_right = T) %>% 
  column_spec(3, border_right = T) %>% 
  column_spec(5, border_right = T) %>% 
  column_spec(c(2:7), width = "30em") %>% 
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  add_header_above(c(" " = 1, "All" = 2, "Female" = 2, "Male" = 2)) %>% 
  footnote(symbol = c("p-value < 0.05", "Hazard ratios (95% confidence intervals) are adjusted for: Model 1) age and sex; Model 2) additionally adjusted for body composition")
           ) 

tablehr1 

```

***
```{r nhp_hr_bodycomp_kable, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE, include=TRUE}

tablehr2 <- kable(nhp_hr_toc_by_3b, row.names = FALSE, col.names = c("Metabolic Index", "Timepoint$^$" , "All", "Female", "Male"), escape = FALSE, align = "c", caption = "Mortality Hazard Ratios of Highest and Lowest Body Weight and Composition Quartiles \n at Varying Timepoints Stratified by Sex") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left", font_size = 12) %>% 
  column_spec(1, bold = TRUE, border_right = T) %>% 
  column_spec(2, border_right = T) %>% 
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  footnote(symbol = c("p-value < 0.05", "Hazard ratios are adjusted for age and sex", "Timepoint is percent of mean life expectancy")
           ) 

tablehr2

```

## Author Questions 
#### determine assessments over lifespan 

```{r nhp_imp_summary, warning=FALSE, message=FALSE, echo=FALSE}

entry <- main_all3 %>% 
  group_by(idno) %>% 
  summarise( 
    age_entry = min(age_wk, na.rm=T), 
    
    whenentry =  as.factor( 
                ifelse((age_entry < (life_exp$life_exp_wk*0.125)), 0,
                ifelse((age_entry >= (life_exp$life_exp_wk*0.125)) & (age_entry < (life_exp$life_exp_wk*0.375)), 25,
                ifelse((age_entry >= (life_exp$life_exp_wk*0.375)) & (age_entry < (life_exp$life_exp_wk*0.625)), 50,
                ifelse((age_entry >= (life_exp$life_exp_wk*0.625)) & (age_entry < (life_exp$life_exp_wk*0.875)), 75,
                ifelse((age_entry >= (life_exp$life_exp_wk*0.875)) & (age_entry < (life_exp$life_exp_wk*1.125)), 100,
                ifelse((age_entry >= (life_exp$life_exp_wk*1.125)), 125, NA))))))))

asmt_funct_n <- function(time, gluc, gluc2, bw, bw2, fat, fat2){ 
  
  gluc_a <- paste0("!is.na(", gluc, ")") 
  gluc_b <- paste0("is.na(", gluc2, ")") 
  gluc2_a <- paste0("!is.na(", gluc2, ")") 
  
  bw_a <- paste0("!is.na(", bw, ")") 
  bw_b <- paste0("is.na(", bw2, ")") 
  bw2_a <- paste0("!is.na(", bw2, ")") 
  
  fat_a <- paste0("!is.na(", fat, ")") 
  fat_b <- paste0("is.na(", fat2, ")") 
  fat2_a <- paste0("!is.na(", fat2, ")") 
  

main_all3 %>% 
  left_join(entry) %>% 
  left_join(census_c16_death) %>% 
  mutate(whendead = as.numeric(as.character(whendead))) %>% 
  mutate(whenvisit = as.numeric(as.character(whenvisit))) %>% 
  mutate(whenentry = as.numeric(as.character(whenentry))) %>% 
  ungroup() %>% 
  summarise( 

      timepoint = time, 
      new_participants = n_distinct(idno[whenentry == time], na.rm = T), 

      alive = n_distinct(idno[whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      dead = n_distinct(idno[whendead == time], na.rm = T),
      
      gluc_seen = n_distinct(idno[eval(parse(text = gluc_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      gluc_imputed = n_distinct(idno[eval(parse(text = gluc2_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T) - gluc_seen, 
      gluc_missing = n_distinct(idno[eval(parse(text = gluc_b)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      
      bw_seen = n_distinct(idno[eval(parse(text = bw_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      bw_imputed = n_distinct(idno[eval(parse(text = bw2_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T) - bw_seen, 
      bw_missing = n_distinct(idno[eval(parse(text = bw_b)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      
      fat_seen = n_distinct(idno[eval(parse(text = fat_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      fat_imputed = n_distinct(idno[eval(parse(text = fat2_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T) - fat_seen, 
      fat_missing = n_distinct(idno[eval(parse(text = fat_b)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      ) 
} 

asmt_25 <- asmt_funct_n(25, "gluc_25_m", "cfgluc_25_m", "bw_25_m", "cfbw_25_m", "fat_25_m", "cffat_25_m")
asmt_50 <- asmt_funct_n(50, "gluc_50_m", "cfgluc_50_m", "bw_50_m", "cfbw_50_m", "fat_50_m", "cffat_50_m")
asmt_75 <- asmt_funct_n(75, "gluc_75_m", "cfgluc_75_m", "bw_75_m", "cfbw_75_m", "fat_75_m", "cffat_75_m")
asmt_100 <- asmt_funct_n(100, "gluc_100_m", "cfgluc_100_m", "bw_100_m", "cfbw_100_m", "fat_100_m", "cffat_100_m")
asmt_125 <- asmt_funct_n(125, "gluc_125_m", "cfgluc_125_m", "bw_125_m", "cfbw_125_m", "fat_125_m", "cffat_125_m")

imputation <- rbind(asmt_25, asmt_50, asmt_75, asmt_100, asmt_125)

```

```{r imputation_kable, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}

kable(imputation, row.names = FALSE, align = "c", format = "html", escape = F,
      caption = "Assessments of Metabolic indices Over the Lifespan", col.names = c("Timepoint$^$", "New Participants", "Alive", "Dead", "Measured", "Imputed", "Missing", "Measured", "Imputed", "Missing", "Measured", "Imputed", "Missing")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left") %>% 
  column_spec(1, bold = T, border_right = T) %>% 
  column_spec(4, border_right = T, width = c("250px"), include_thead =F) %>% 
  column_spec(7, border_right = T) %>% 
  column_spec(10, border_right = T) %>% 
  collapse_rows(columns = 1, valign = "middle") %>% 
  add_header_above(c(" " = 1, "Population" = 3, "Glucose" = 3, "Body Weight" = 3, "Body Composition" = 3)) %>%  
  footnote(symbol = c("Imputation using last observed value carried forward", "Timepoint is percent of mean life expectancy"))

```

***

#### trajectories by varying survival groups

```{r nhp_lifespan_dist_fig}

annotate_figure(nhp_fig_d13, 
                top = text_grob("Trends in Metabolic indices Over the Lifespan By Survival Group"), 
                bottom = text_grob("Note: \nHorizontal lines indicate percent of life expectancy: green (25%), orange (50%), red (75%), purple (100%), and black (125%) \nDiamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines")) 
)

```
 
*** 
