---
title: "SLAM Glucose, Body Weight, and Body Comp"
author: "Eric Shiroma, John Nam"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
---

## Study Description 

- This is an investigation into the longitudinal trajectories of glucose, body weight, and body composition in SLAM (a mouse cohort). These results will be used to compare similar trajectories in humans and monkeys. 

***

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}

##loading packages
library (mgcv)
library (BSDA)
library (survival)
library (broom)
library (ggplot2)
library (ggthemes)
library (extrafont)
library (arsenal)
library (psych)
library (survey)
library (lubridate)
library (kableExtra)
library (gridExtra)
library (ggpubr)
library (grid)
library (tinytex)
library (haven)
library (tidyverse)
library (finalfit)
library (survival)
library (survminer)

#setting global knitr options for r markdown output
knitr::opts_chunk$set(autodep = TRUE) 
knitr::opts_chunk$set(warning = FALSE) 
knitr::opts_chunk$set(message = FALSE) 
knitr::opts_chunk$set(echo = FALSE) 
knitr::opts_chunk$set(cache = TRUE) 

```
## Analysis 
#### census data read in
  - input file census.csv (10/30/2019)
  - edit csv column name 'Animal_ID'
  
```{r slam_census, warning=FALSE, message=FALSE, echo=FALSE}

# edit csv column name 'Animal_ID' 
#census <- read.csv("raw files/census.csv") %>% 
census <- read.csv("/Users/shiromaej/Box/ISG Code Check/data/slam/census.csv") %>% 
  select(idno, animal = Animal_ID, tag, sex, strain, dob, cohort, taghistory) 
  census$dob <- as.Date(census$dob, "%m/%d/%Y") 
  census$idno <- as.character(census$idno) 
  census$tag <- as.character(census$tag) 
  census$taghistory <- as.character(census$taghistory) 
  
``` 

#### body weight data read in and clean
  - input file BW_Temp_FC_All_2019-11-16 (11/16/2019)
  - clean body weights (funny values) 
    - text to missing 
    - bw < 10 to missing 
  
```{r slam_bodyweight, warning=FALSE, message=FALSE, echo=FALSE}

##  
bodywt <- read.csv("/Users/shiromaej/Box/ISG Code Check/data/slam/BW_Temp_FC_All_2019-11-16.csv") %>% 
#bodywt <- read.csv("raw files/BW_Temp_FC_All_2019-11-16.csv") %>% 
  select(idno, date, bw) 

invalid_bw <- bodywt$bw[is.na(as.numeric(as.character(bodywt$bw)))] 
bodywt <- bodywt %>% 
  mutate(
    bw = ifelse(bw == "", NA, 
          ifelse(bw == "NA", NA, 
          ifelse(bw == "wholecaedead", NA, 
          ifelse(bw == "MIA", NA, 
          ifelse(bw == "fd5-23", NA, 
          ifelse(bw == "FD", NA, 
          ifelse(bw == "f/d6/7/2018", NA, 
          ifelse(bw == "f/d2/7/18", NA,  
          ifelse(bw == "f/d2/27", NA, 
          ifelse(bw == "DVR", NA, 
          ifelse(bw == "deadd", NA, 
          ifelse(bw == "dead5-9-18", NA, 
          ifelse(bw == "dead5-8-18", NA, 
          ifelse(bw == "dead5-21-18", NA, 
          ifelse(bw == "dead5-14", NA, 
          ifelse(bw == "dead5-10", NA, 
          ifelse(bw == "dead1-30", NA, 
          ifelse(bw == "dead", NA, 
          ifelse(bw == "Dead", NA, 
          ifelse(bw == "DEAD", NA, 
          ifelse(bw == "BW", NA, 
          ifelse(bw == "culled5-18", NA, as.character(bw)))))))))))))))))))))))
    )
  bodywt$date <- as.Date(bodywt$date, "%m/%d/%Y") 
  bodywt$bw <- as.numeric(as.character(bodywt$bw)) 
  bodywt$idno <- as.character(bodywt$idno) 
  bodywt <- bodywt %>% 
    mutate( 
    bw = ifelse(bw < 10, NA, bw) 
    ) %>% 
    filter(!is.na(bw))
  
bodywt <- bodywt[!duplicated(bodywt[c(1,2)]),]
  
```

#### glucose data read in and clean 
  - input file Glucose_Lactate.csv (10/30/2019)
  - edit csv column name 'Stress_Notes'
  - clean glucose (funny values) 
    - text to missing 
    - gluc < 20 or gluc >= 500 to missing 
    
```{r slam_glucose, warning=FALSE, message=FALSE, echo=FALSE}

# edit csv column name 'Stress_Notes'
#glucose <- read.csv("raw files/Glucose_Lactate.csv") %>% 
glucose <- read.csv("/Users/shiromaej/Box/ISG Code Check/data/slam/Glucose_Lactate.csv") %>% 
  select(idno, date, gluc = Glucose) %>% 
  mutate( 
    dead = ifelse(gluc == "DEAD", 1, NA),
    gluc = ifelse(gluc == "DEAD", NA, 
                     ifelse(gluc == "dead", NA, 
                     ifelse(gluc == "missed", NA, 
                     ifelse(gluc == "missing", NA, 
                     ifelse(gluc == "NOT BLED", NA, 
                     ifelse(gluc == "SAC", NA, 
                     ifelse(gluc == "89/148", NA, 
                     ifelse(gluc == "", NA, 
                     ifelse(gluc == "no glucose", NA, 
                     ifelse(gluc == "Not accounted for", NA, 
                     ifelse(gluc == "would not bleed", NA, 
                     ifelse(gluc == "", NA, as.character(gluc)))))))))))))
  )
  glucose$date <- as.Date(glucose$date, "%m/%d/%Y") 
  glucose$gluc <- as.numeric(glucose$gluc) 
  glucose$idno <- as.character(glucose$idno) 
  glucose <- glucose %>% 
    mutate( 
    gluc = ifelse(gluc < 20 | gluc >= 500, NA, gluc) 
    ) 
  
```

#### body comp data read in and clean 
  - input file NMR.csv (10/30/2019)
  - clean bodycomp (funny values) 
    - text to missing 
    - fat <= 0 to missing

```{r slam_bodycomp, warning=FALSE, message=FALSE, echo=FALSE}
##    

#bodycomp <- read.csv("raw files/NMR.csv") %>% 
bodycomp <- read.csv("/Users/shiromaej/Box/ISG Code Check/data/slam/NMR.csv") %>% 
  select (idno, date, fat, lean) 
  
invalid_fat <- bodycomp$fat[is.na(as.numeric(as.character(bodycomp$fat)))] 
bodycomp <- bodycomp %>% 
  mutate(fat = ifelse(fat %in% invalid_fat, NA, fat)) 

  bodycomp$date <- as.Date(bodycomp$date, "%m/%d/%Y") 
  bodycomp$idno <- as.character(bodycomp$idno) 
  bodycomp$fat <- as.numeric(as.character(bodycomp$fat)) 
  bodycomp <- bodycomp %>% 
    mutate( 
      fat = ifelse(fat <= 0, NA, fat) 
    )

```

```{r all_to_new_idno, warning=FALSE, message=FALSE, echo=FALSE}

#changing old idnos to new idnos in body comp file
fixID_history <- function(idno, census){ 
  
 census <- census %>% 
   separate(taghistory, c("th1", "th2", "th3", "th4", "th5", "th6"), sep = ",", extra = "warn") 
  
  idnos <- trimws(idno)
  idnos[idnos %in% census$Animal.ID] <- census$idno[match(idnos[idnos %in% census$Animal.ID], census$Animal.ID)]
  idnos[idnos %in% census$tag] <- census$idno[match(idnos[idnos %in% census$tag], census$tag)]
  idnos[idnos %in% census$th1] <- census$idno[match(idnos[idnos %in% census$th1], census$th1)]
  idnos[idnos %in% census$th2] <- census$idno[match(idnos[idnos %in% census$th2], census$th2)]
  idnos[idnos %in% census$th3] <- census$idno[match(idnos[idnos %in% census$th3], census$th3)]
  idnos[idnos %in% census$th4] <- census$idno[match(idnos[idnos %in% census$th4], census$th4)]
  idnos[idnos %in% census$th5] <- census$idno[match(idnos[idnos %in% census$th5], census$th5)]
  idnos[idnos %in% census$th6] <- census$idno[match(idnos[idnos %in% census$th6], census$th6)]

  return(idnos)
} 

#tag history 
bodycomp$idno <- fixID_history(bodycomp$idno, census) 
 
```

#### survival data read in and clean 
  - input file Survival.csv (10/30/2019)
  - clean survival (funny values) 
    - text to missing 
    - recode old to new idnos 
    
```{r slam_survival, warning=FALSE, message=FALSE, echo=FALSE}

#surv <- read.csv("raw files/Survival.csv") 
surv <- read.csv("/Users/shiromaej/Box/ISG Code Check/data/slam/Survival.csv") 
  surv$tod <- as.Date(surv$Died, "%m/%d/%Y") 
  surv$tag <- as.character(surv$tag) 
  surv$cod <- as.character((surv$CoD)) 
  surv <- surv %>% select(-Died, - CoD) %>% 
  mutate(tag = ifelse(tag == "R", "179060", 
               ifelse(tag == "L", "179217", tag))) 

```

#### merging all datasets into a main file and clean
  - glucose
  - bodywt (body weight)
  - bodycomp (body composition)
  - census
  - removal of timepoints before dob 
  
```{r slam_main_all, warning=FALSE, message=FALSE, echo=FALSE}
##  
main_all <- glucose %>% 
  full_join(bodywt, by = c("idno" , "date")) %>% 
  full_join(bodycomp, by = c("idno" , "date")) %>% 
  left_join(census, by = "idno") %>%
  mutate( 
    age_wk = round(((date - dob) / 7), 2) +1
  ) %>% 
  filter(!is.na(date) & !is.na(idno)) %>% 
  filter(age_wk > 0)
main_all$age_wk <- as.numeric(main_all$age_wk) 
  
#renamed for isg template
slam_main_all <- main_all 
 
```
  
#### restriction to cohorts 1 to 6

```{r slam_ratio, warning=FALSE, message=FALSE, echo=FALSE}

main_all2 <- main_all %>% filter(cohort >= 1 & cohort <= 6)

#getting unique idnos 
main_all2_n <- n_distinct(main_all2$idno)

sexstrain <- main_all2 %>% 
  select(idno, sex, strain) %>% 
  distinct() 

sextable <- table(sexstrain$sex)
straintable <- table(sexstrain$strain)

# calculate % sex and strain
slam_sex_ratio = format(round(((sextable[1] / (sextable[1] + sextable[2]))*100), 2), nsmall = 2)
slam_strain_ratio = format(round(((straintable[1] / (straintable[1] + straintable[2]))*100), 2), nsmall = 2)

```

#### population characteristics
- `r main_all2_n` mice with valid indices
- `r slam_sex_ratio` % female
- `r slam_strain_ratio` % B6

#### age characteristics

```{r slam_age_desc, warning=FALSE, message=FALSE, echo=FALSE}

main_all2_desc <- main_all2 %>%
group_by(idno, sex, strain) %>%
summarise(
  age_wk = min(age_wk, na.rm=T)
) 

main_all2_desc2_all <- format(round(describe(main_all2_desc$age_wk), 2), nsmall=2) %>% 
  mutate_each(funs(as.numeric(.))) %>%
  mutate(group1 = "all") %>% 
  select(-c("vars")) 
main_all2_desc_sex <- as.data.frame(describeBy(main_all2_desc$age_wk, main_all2_desc$sex, mat=TRUE, digits = 2)) %>% 
  select(-c("item", "vars")) %>% 
  mutate(n = as.numeric(n))
main_all2_desc_strain <- as.data.frame(describeBy(main_all2_desc$age_wk, main_all2_desc$strain, mat=TRUE, digits = 2)) %>% 
  select(-c("item", "vars")) %>% 
  mutate(n = as.numeric(n))

main_all2_desc2 <- bind_rows(main_all2_desc2_all, main_all2_desc_sex, main_all2_desc_strain) %>% 
  select(group1, everything())

```

```{r slam_age_desc_kable, warning=FALSE, message=FALSE, echo=FALSE}

kable(main_all2_desc2, caption = "Age Characteristics",row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### frequencies of measures

```{r slam_follow_up, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}

n_asmt <- main_all2 %>% 
  group_by(idno, sex) %>% 
  summarise_all(funs( 
    n = sum(!is.na(.)), 
    sd = format(round(sd(!is.na(.)), 2), nsmall=2) 
  )) 

gluc_asmt <- format(round(as.data.frame(psych::describe(n_asmt$gluc_n)), 2), nsmall=2) %>% 
  mutate(indices = "glucose")
bw_asmt <- format(round(as.data.frame(psych::describe(n_asmt$bw_n)), 2), nsmall=2) %>% 
  mutate(indices = "body weight")
fat_asmt <- format(round(as.data.frame(psych::describe(n_asmt$fat_n)), 2), nsmall=2) %>% 
  mutate(indices = "fat")

all_asmt <- bind_rows(gluc_asmt, bw_asmt, fat_asmt) %>% 
  select(indices, everything(), -vars)

slam_gluc_asmt <- gluc_asmt
slam_bw_asmt <- bw_asmt
slam_fat_asmt <- fat_asmt

```

```{r asmt_kable, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}

kable(all_asmt, caption = "Assessment Frequencies", row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

*** 

#### census_c16_death creation
  - merge 
    - census
    - surv
    - main_all2
  - restrict to cohorts 1 to 6
  - remove mice with missing time of death
  - create all cause death censor category
    - all-cause mortality included: "Found dead", "Per PI", "Culled", "DVR or Pathology", "Culled Per Vet"
    
***

```{r slam_censor_df, warning=FALSE, message=FALSE, echo=FALSE}

census_c16_death <- main_all2 %>% 
  select(idno, tag, dob) %>% 
  distinct() %>% 
  left_join(surv, by = "tag") %>% 
  left_join(main_all2, by = c("idno", "dob")) %>%
  group_by(idno, cod, dob, sex, strain) %>% 
  summarise(maxdate = max(tod), 
            age_first = min(age_wk, na.rm = T), 
            firstdate = min(date, na.rm = T), 
            lastdate = max(date, na.rm = T)) %>% 
  mutate(
    fu_age_wk = floor(((maxdate - firstdate) + 1) / 7),
    le_wk = floor(((maxdate - dob) + 1) / 7),
    dead_censor = ifelse(cod == "Found dead", 1, 
              ifelse(cod == "Per PI", 1, 
              ifelse(cod == "Per Vet", 1,
              ifelse(cod == "Culled", 1, 
              ifelse(cod == "DVR or Pathology", 1,        
              ifelse(cod == "Culled Per Vet", 1,0)))))) 
         ) %>% 
  ungroup() %>% 
  filter(!is.na(maxdate)) %>% 
  select(idno, dead_censor, fu_age_wk, le_wk, cod, age_first, firstdate, lastdate, sex, strain)

codtable <- as.data.frame(table(census_c16_death$cod)) 
dctable1 <- as.data.frame(table(census_c16_death$dead_censor)) 
fu_time_desc <- format(round(as.data.frame(psych::describe(as.numeric(census_c16_death$fu_age_wk))), 2), nsmall=2) 

```

#### cause of death 

```{r cod_cat, warning=FALSE, message=FALSE, echo=FALSE} 
kable(codtable, caption = "Cause of Death", col.names = c("Cause of Death", "Frequency"), row.names = FALSE, align = "c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

***

#### death frequency

```{r death_cat, warning=FALSE, message=FALSE, echo=FALSE} 
kable(dctable1, caption = "Death Frequency", col.names = c("Dead = 1", "Frequency"), row.names = FALSE, align = "c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

***

#### follow up time

```{r followup_cat, warning=FALSE, message=FALSE, echo=FALSE} 
kable(fu_time_desc, caption = "Follow Up Time", row.names = FALSE, align = "c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE) 
```

*** 

#### life expectancy

```{r slam_life_expectancy, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}
#function to create life expectancy for different groups
create_LE <- function(data, criteria, filter){

life_exp <- data %>% 
  ungroup() %>% 
  filter(!!filter) %>% 
  summarise( 
            total = n_distinct(idno), 
            dead = n_distinct(idno[!!criteria]), 
            life_exp_wk = round(mean(le_wk[!!criteria]), 2), 
            sd = round(sd(as.numeric(le_wk[!!criteria])), 2), 
            se = round(sd(as.numeric(le_wk[!!criteria])) / sqrt(dead),2),
            lcl = round(life_exp_wk - 1.96 * se, 2), 
            ucl = round(life_exp_wk + 1.96 * se, 2) 
  ) 
life_exp$life_exp_wk <- as.numeric(life_exp$life_exp_wk) 

return(life_exp)

}

life_exp <- create_LE(data = census_c16_death, criteria = quote(dead_censor == 1), filter = quote(sex %in% c("M", "F"))) %>% 
  mutate(group = "all") 
life_exp_f <- create_LE(data = census_c16_death, criteria = quote(dead_censor == 1), filter = quote(sex %in% c("F"))) %>% 
  mutate(group = "female") 
life_exp_m <- create_LE(data = census_c16_death, criteria = quote(dead_censor == 1), filter = quote(sex %in% c("M"))) %>% 
  mutate(group = "male") 
life_exp_het3 <- create_LE(data = census_c16_death, criteria = quote(dead_censor == 1), filter = quote(strain %in% c("HET3"))) %>% 
  mutate(group = "HET3") 
life_exp_b6 <- create_LE(data = census_c16_death, criteria = quote(dead_censor == 1), filter = quote(strain %in% c("B6"))) %>% 
  mutate(group = "B6") 

life_exp2 <- bind_rows(life_exp, life_exp_f, life_exp_m, life_exp_het3, life_exp_b6) %>% 
  select(group, everything())
  
## for isg template
slam_life_exp <- life_exp 

```

```{r slam_life_exp_kable, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}
kable(life_exp2[,c(1:5)], caption = "Life Expectancy", col.names = c("Sex", "Total", "Dead", "Life Expectancy <br> (weeks)", "sd"), row.names = TRUE, align = "c", escape = FALSE) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left") 
```

***

#### age at peak indices value creation 

```{r slam_life_peaks, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}

#functions to create peaks for different groups
lifespan_peaks <- function(metric, data, le, filter){

#set boundary conditions for peak to lowest and highest life expectancy timepoints
low_bound <- floor(le$life_exp_wk*0.25)
high_bound <- floor(le$life_exp_wk*1.25)

#subset arguments
sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)") 
metric2 <- paste0("!is.na(", metric, ")") 

#subset for non-missing data
df <- base::subset(data, eval(parse(text = metric2))) 

#subset data by xlimit and subgroup restraints
peaks_df <- subset(df, eval(parse(text = sub))) %>% 
  arrange(age_wk) %>% 
  filter(!!filter) 

#gam model fit using geom_smooth parameters
gam_fit <- mgcv::gam(eval(parse(text = metric)) ~ s(age_wk, bs = "cs"), method = "REML", data = peaks_df) 

#create fitted values using gam model from dataframe age values
predict1 <- as.data.frame(predict.gam(gam_fit, peaks_df, se.fit=TRUE)) 

#create n and sd by age group
fit2 <- peaks_df %>% 
  bind_cols(predict1) %>% 
  select(age_wk, fit, se.fit) %>% 
  filter(age_wk >= low_bound, age_wk <= high_bound) %>%
  group_by(age_wk) %>% 
  mutate( 
    n = n(), 
    sd = se.fit*sqrt(n) 
  ) %>% 
  ungroup() 

#create low bound peak object
low_bound_peak <- fit2 %>% 
  # filter(age_wk >= low_bound) %>% 
  filter(age_wk == min(age_wk)) %>% 
  mutate(le = (age_wk/le$life_exp_wk)*100) %>% 
  select(age_wk, le, fit, se.fit, n, sd) %>% 
  distinct()

#create high bound peak object
high_bound_peak <- fit2 %>% 
  # filter(age_wk <= high_bound) %>% 
  filter(age_wk == max(age_wk)) %>% 
  mutate(le = (age_wk/le$life_exp_wk)*100) %>% 
  select(age_wk, le, fit, se.fit, n, sd) %>% 
  distinct()

#pulling peak age and value
infl <- NULL
infl$logic <- c(NA, diff(diff(fit2$fit)) < 0, NA)
infl <- as.data.frame(infl)

#finding age of the first point significantly different from peak for the lower range
#setting to lower bound if no values are present
peak <- fit2 %>% 
  bind_cols(infl) %>% 
  filter(infl$logic == "TRUE") %>% 
  group_by(age_wk, n, sd, se.fit) %>%
  summarise(
    fit = max(fit)
  )  %>% 
  ungroup() %>% 
  filter(fit == max(fit, na.rm=T)) %>% 
  mutate(le = (age_wk/le$life_exp_wk)*100)

#finding age of the first point significantly different from peak for the lower range
#setting to lower bound if no values are present
fit_diff_min <- tryCatch(
fit2 %>% 
  filter(age_wk < peak$age_wk) %>% 
  mutate(fit_diff = fit - peak$fit) %>% 
  mutate(age_diff = peak$age_wk - age_wk) %>% 
  mutate(pval = tsum.test(fit_diff, sd, n, mu = 0, conf.level = 0.95)$p.value) %>% 
  filter(pval <= 0.05) %>% 
  filter( 
    age_diff == min(age_diff) 
    ) %>% 
  distinct() %>% 
  mutate(le = (age_wk/le$life_exp_wk)*100) %>% 
  select(age_wk, le, fit), 
error=function(e) low_bound_peak)

#finding age of the first point significantly different from peak for the higher range
#setting to higher bound if no values are present
fit_diff_max <- tryCatch(
fit2 %>% 
  filter(age_wk > peak$age_wk) %>% 
  mutate(fit_diff = fit - peak$fit) %>% 
  mutate(age_diff = age_wk - peak$age_wk) %>% 
  mutate(pval = tsum.test(fit_diff, sd, n, mu = 0, conf.level = 0.95)$p.value) %>% 
  filter(pval <= 0.05) %>% 
  filter( 
    age_diff == min(age_diff) 
    ) %>% 
  distinct() %>% 
  mutate(le = (age_wk/le$life_exp_wk)*100) %>% 
  select(age_wk, le, fit),
error=function(e) high_bound_peak)

#setting to lower bound if no values are present
if(dim(fit_diff_min)[1] == 0){
  fit_diff_min <- low_bound_peak
}else{
}

#setting to higher bound if no values are present
if(dim(fit_diff_max)[1] == 0){
  fit_diff_max <- high_bound_peak
}else{
}

#putting all outputs into list 
all_peaks <- list(peak, fit_diff_min, fit_diff_max, high_bound_peak, low_bound_peak) 
names(all_peaks) <- c("all", "min", "max", "highbound", "lowbound")

return(all_peaks) 

} 

gluc_peak <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp, filter = quote(sex %in% c("M", "F"))) 
bw_peak <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp, filter = quote(sex %in% c("M", "F"))) 
fat_peak <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp, filter = quote(sex %in% c("M", "F"))) 

gluc_peak_male <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp_m, filter = quote(sex %in% c("M"))) 
bw_peak_male <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp_m, filter = quote(sex %in% c("M"))) 
fat_peak_male <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp_m, filter = quote(sex %in% c("M"))) 

gluc_peak_female <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp_f, filter = quote(sex %in% c("F"))) 
bw_peak_female <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp_f, filter = quote(sex %in% c("F"))) 
fat_peak_female <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp_f, filter = quote(sex %in% c("F"))) 

gluc_peak_HET3 <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp_het3, filter = quote(strain %in% c("HET3"))) 
bw_peak_HET3 <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp_het3, filter = quote(strain %in% c("HET3"))) 
fat_peak_HET3 <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp_het3, filter = quote(strain %in% c("HET3")))  

gluc_peak_B6 <- lifespan_peaks(metric = "gluc", data = main_all2, le = life_exp_b6, filter = quote(strain %in% c("B6")))  
bw_peak_B6 <- lifespan_peaks(metric = "bw", data = main_all2, le = life_exp_b6, filter = quote(strain %in% c("B6"))) 
fat_peak_B6 <- lifespan_peaks(metric = "fat", data = main_all2, le = life_exp_b6, filter = quote(strain %in% c("B6"))) 

```

#### main_cat2 creation 
- merge 
  - main_all2 
  - census_c16_death 
- create categories by survival length 
  - needed for survival bias sensitivity 
  
```{r slam_main_cat, warning=FALSE, message=FALSE, echo=FALSE}
  
create_main_cat <- function(life_exp){
  
main_cat <- main_all2 %>% 
  left_join(census_c16_death) %>% 
  ungroup() %>% 
  mutate( 
    
    whenvisit = 
                ifelse((age_wk >= (life_exp$life_exp_wk*0)) & (age_wk < (life_exp$life_exp_wk*0.125)), 0,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.125)) & (age_wk < (life_exp$life_exp_wk*0.375)), 25,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.375)) & (age_wk < (life_exp$life_exp_wk*0.625)), 50,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.625)) & (age_wk < (life_exp$life_exp_wk*0.875)), 75,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.875)) & (age_wk < (life_exp$life_exp_wk*1.125)), 100,
                ifelse((age_wk >= (life_exp$life_exp_wk*1.125)), 125, NA)))))), 
    whendead =  
                ifelse((le_wk >= (life_exp$life_exp_wk*0)) & (le_wk < (life_exp$life_exp_wk*0.125)), 0,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.125)) & (le_wk < (life_exp$life_exp_wk*0.375)), 25,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.375)) & (le_wk < (life_exp$life_exp_wk*0.625)), 50,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.625)) & (le_wk < (life_exp$life_exp_wk*0.875)), 75,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.875)) & (le_wk < (life_exp$life_exp_wk*1.125)), 100,
                ifelse((le_wk >= (life_exp$life_exp_wk*1.125)), 125, NA)))))), 

    
    gluc_25 = ifelse(whenvisit == 25, gluc, NA),
    gluc_50 = ifelse(whenvisit == 50, gluc, NA),
    gluc_75 = ifelse(whenvisit == 75, gluc, NA),
    gluc_100 = ifelse(whenvisit == 100, gluc, NA),
    gluc_125 = ifelse(whenvisit == 125, gluc, NA), 
    
    bw_25 = ifelse(whenvisit == 25, bw, NA),
    bw_50 = ifelse(whenvisit == 50, bw, NA),
    bw_75 = ifelse(whenvisit == 75, bw, NA),
    bw_100 = ifelse(whenvisit == 100, bw, NA),
    bw_125 = ifelse(whenvisit == 125, bw, NA),
    
    fat_25 = ifelse(whenvisit == 25, fat, NA),
    fat_50 = ifelse(whenvisit == 50, fat, NA),
    fat_75 = ifelse(whenvisit == 75, fat, NA),
    fat_100 = ifelse(whenvisit == 100, fat, NA), 
    fat_125 = ifelse(whenvisit == 125, fat, NA), 
    
    #survival specific categories
    dgluc_25 = ifelse(whendead >= 25 & whenvisit == 25, gluc, NA),
    dgluc_50 = ifelse(whendead >= 50 & whenvisit == 50, gluc, NA),
    dgluc_75 = ifelse(whendead >= 75 & whenvisit == 75, gluc, NA),
    dgluc_100 = ifelse(whendead >= 100 & whenvisit == 100, gluc, NA),
    dgluc_125 = ifelse(whendead >= 125 & whenvisit == 125, gluc, NA),
    
    dbw_25 = ifelse(whendead >= 25 & whenvisit == 25, bw, NA),
    dbw_50 = ifelse(whendead >= 50 & whenvisit == 50, bw, NA),
    dbw_75 = ifelse(whendead >= 75 & whenvisit == 75, bw, NA),
    dbw_100 = ifelse(whendead >= 100 & whenvisit == 100, bw, NA),
    dbw_125 = ifelse(whendead >= 125 & whenvisit == 125, bw, NA),
    
    dfat_25 = ifelse(whendead >= 25 & whenvisit == 25, fat, NA),
    dfat_50 = ifelse(whendead >= 50 & whenvisit == 50, fat, NA),
    dfat_75 = ifelse(whendead >= 75 & whenvisit == 75, fat, NA),
    dfat_100 = ifelse(whendead >= 100 & whenvisit == 100, fat, NA),
    dfat_125 = ifelse(whendead >= 125 & whenvisit == 125, fat, NA)
    
  ) %>% 
  
  group_by(idno, whendead) %>% 
  summarise( 

    gluc_25_m = mean(gluc_25, na.rm = TRUE),
    gluc_50_m = mean(gluc_50, na.rm = TRUE),
    gluc_75_m = mean(gluc_75, na.rm = TRUE),
    gluc_100_m = mean(gluc_100, na.rm = TRUE),
    gluc_125_m = mean(gluc_125, na.rm = TRUE),
    
    bw_25_m = mean(bw_25, na.rm = TRUE),
    bw_50_m = mean(bw_50, na.rm = TRUE),
    bw_75_m = mean(bw_75, na.rm = TRUE),
    bw_100_m = mean(bw_100, na.rm = TRUE),
    bw_125_m = mean(bw_125, na.rm = TRUE),
    
    fat_25_m = mean(fat_25, na.rm = TRUE),
    fat_50_m = mean(fat_50, na.rm = TRUE),
    fat_75_m = mean(fat_75, na.rm = TRUE),
    fat_100_m = mean(fat_100, na.rm = TRUE),
    fat_125_m = mean(fat_125, na.rm = TRUE), 
    
    #survival specific categories
    dgluc_25_m = mean(dgluc_25, na.rm = TRUE),
    dgluc_50_m = mean(dgluc_50, na.rm = TRUE),
    dgluc_75_m = mean(dgluc_75, na.rm = TRUE),
    dgluc_100_m = mean(dgluc_100, na.rm = TRUE),
    dgluc_125_m = mean(dgluc_125, na.rm = TRUE),
    
    dbw_25_m = mean(dbw_25, na.rm = TRUE),
    dbw_50_m = mean(dbw_50, na.rm = TRUE),
    dbw_75_m = mean(dbw_75, na.rm = TRUE),
    dbw_100_m = mean(dbw_100, na.rm = TRUE),
    dbw_125_m = mean(dbw_125, na.rm = TRUE),
    
    dfat_25_m = mean(dfat_25, na.rm = TRUE),
    dfat_50_m = mean(dfat_50, na.rm = TRUE),
    dfat_75_m = mean(dfat_75, na.rm = TRUE),
    dfat_100_m = mean(dfat_100, na.rm = TRUE),
    dfat_125_m = mean(dfat_125, na.rm = TRUE)
  ) 

main_cat2 <- main_cat %>% 
  
  mutate(
    glucd_25_m = gluc_25_m - gluc_50_m, 
    glucd_50_m = gluc_50_m - gluc_50_m, 
    glucd_75_m = gluc_75_m - gluc_50_m, 
    glucd_100_m = gluc_100_m - gluc_50_m,  
    glucd_125_m = gluc_125_m - gluc_50_m, 
    
    bwd_25_m = bw_25_m - bw_50_m, 
    bwd_50_m = bw_50_m - bw_50_m, 
    bwd_75_m = bw_75_m - bw_50_m, 
    bwd_100_m = bw_100_m - bw_50_m,  
    bwd_125_m = bw_125_m - bw_50_m, 
    
    fatd_25_m = fat_25_m - fat_50_m, 
    fatd_50_m = fat_50_m - fat_50_m, 
    fatd_75_m = fat_75_m - fat_50_m, 
    fatd_100_m = fat_100_m - fat_50_m,  
    fatd_125_m = fat_125_m - fat_50_m, 
    
    glucp_25_m = glucd_25_m / gluc_50_m *100, 
    glucp_50_m = glucd_50_m / gluc_50_m *100, 
    glucp_75_m = glucd_75_m / gluc_50_m *100, 
    glucp_100_m = glucd_100_m / gluc_50_m *100, 
    glucp_125_m = glucd_125_m / gluc_50_m *100, 
    
    bwp_25_m = bwd_25_m / bw_50_m *100, 
    bwp_50_m = bwd_50_m / bw_50_m *100, 
    bwp_75_m = bwd_75_m / bw_50_m *100, 
    bwp_100_m = bwd_100_m / bw_50_m *100, 
    bwp_125_m = bwd_125_m / bw_50_m *100, 
    
    fatp_25_m = fatd_25_m / fat_50_m *100, 
    fatp_50_m = fatd_50_m / fat_50_m *100, 
    fatp_75_m = fatd_75_m / fat_50_m *100, 
    fatp_100_m = fatd_100_m / fat_50_m *100, 
    fatp_125_m = fatd_125_m / fat_50_m *100, 
    
##
    cfgluc_25_m = gluc_25_m, 
    cfgluc_50_m = ifelse(whendead >= 50 & is.na(gluc_50_m) & !is.na(gluc_25_m), gluc_25_m, gluc_50_m), 
    cfgluc_75_m = ifelse(whendead >= 75 & is.na(gluc_75_m) & !is.na(gluc_50_m), gluc_50_m, 
                  ifelse(whendead >= 75 & is.na(gluc_75_m) & !is.na(gluc_25_m), gluc_25_m, gluc_75_m)), 
    cfgluc_100_m = ifelse(whendead >= 100 & is.na(gluc_100_m) & !is.na(gluc_75_m), gluc_75_m, 
                   ifelse(whendead >= 100 & is.na(gluc_100_m) & !is.na(gluc_50_m), gluc_50_m, 
                   ifelse(whendead >= 100 & is.na(gluc_100_m) & !is.na(gluc_25_m), gluc_25_m, gluc_100_m))), 
    cfgluc_125_m = ifelse(whendead >= 125 & is.na(gluc_125_m) & !is.na(gluc_100_m), gluc_100_m, 
                   ifelse(whendead >= 125 & is.na(gluc_125_m) & !is.na(gluc_75_m), gluc_75_m,  
                   ifelse(whendead >= 125 & is.na(gluc_125_m) & !is.na(gluc_50_m), gluc_50_m, 
                   ifelse(whendead >= 125 & is.na(gluc_125_m) & !is.na(gluc_25_m), gluc_25_m, gluc_125_m)))), 

    cfbw_25_m = bw_25_m, 
    cfbw_50_m = ifelse(whendead >= 50 & is.na(bw_50_m) & !is.na(bw_25_m), bw_25_m, bw_50_m), 
    cfbw_75_m = ifelse(whendead >= 75 & is.na(bw_75_m) & !is.na(bw_50_m), bw_50_m, 
                  ifelse(whendead >= 75 & is.na(bw_75_m) & !is.na(bw_25_m), bw_25_m, bw_75_m)), 
    cfbw_100_m = ifelse(whendead >= 100 & is.na(bw_100_m) & !is.na(bw_75_m), bw_75_m, 
                   ifelse(whendead >= 100 & is.na(bw_100_m) & !is.na(bw_50_m), bw_50_m, 
                   ifelse(whendead >= 100 & is.na(bw_100_m) & !is.na(bw_25_m), bw_25_m, bw_100_m))), 
    cfbw_125_m = ifelse(whendead >= 125 & is.na(bw_125_m) & !is.na(bw_100_m), bw_100_m, 
                   ifelse(whendead >= 125 & is.na(bw_125_m) & !is.na(bw_75_m), bw_75_m,  
                   ifelse(whendead >= 125 & is.na(bw_125_m) & !is.na(bw_50_m), bw_50_m, 
                   ifelse(whendead >= 125 & is.na(bw_125_m) & !is.na(bw_25_m), bw_25_m, bw_125_m)))), 

    cffat_25_m = fat_25_m, 
    cffat_50_m = ifelse(whendead >= 50 & is.na(fat_50_m) & !is.na(fat_25_m), fat_25_m, fat_50_m), 
    cffat_75_m = ifelse(whendead >= 75 & is.na(fat_75_m) & !is.na(fat_50_m), fat_50_m, 
                  ifelse(whendead >= 75 & is.na(fat_75_m) & !is.na(fat_25_m), fat_25_m, fat_75_m)), 
    cffat_100_m = ifelse(whendead >= 100 & is.na(fat_100_m) & !is.na(fat_75_m), fat_75_m, 
                   ifelse(whendead >= 100 & is.na(fat_100_m) & !is.na(fat_50_m), fat_50_m, 
                   ifelse(whendead >= 100 & is.na(fat_100_m) & !is.na(fat_25_m), fat_25_m, fat_100_m))), 
    cffat_125_m = ifelse(whendead >= 125 & is.na(fat_125_m) & !is.na(fat_100_m), fat_100_m, 
                   ifelse(whendead >= 125 & is.na(fat_125_m) & !is.na(fat_75_m), fat_75_m,  
                   ifelse(whendead >= 125 & is.na(fat_125_m) & !is.na(fat_50_m), fat_50_m, 
                   ifelse(whendead >= 125 & is.na(fat_125_m) & !is.na(fat_25_m), fat_25_m, fat_125_m)))), 

  ) %>% 
  ungroup() 

return(main_cat2)

} 

## 
main_cat2 <- create_main_cat(life_exp) 
main_cat2_f <- create_main_cat(life_exp_f) 
main_cat2_m <- create_main_cat(life_exp_m) 

##for isg template
slam_main_cat2 <- main_cat2 
 
``` 

#### main_all3 creation
- merge
  - main_all2
  - census_c16_death
  - main_cat2

```{r slam_main_all3, warning=FALSE, message=FALSE, echo=FALSE}

main_all3 <- main_all2 %>% 
  left_join(census_c16_death) %>% 
  left_join(main_cat2) %>% 
  
  mutate( 

    whenvisit = as.factor( 
                ifelse((age_wk >= (life_exp$life_exp_wk*0)) & (age_wk < (life_exp$life_exp_wk*0.125)), 0,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.125)) & (age_wk < (life_exp$life_exp_wk*0.375)), 25,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.375)) & (age_wk < (life_exp$life_exp_wk*0.625)), 50,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.625)) & (age_wk < (life_exp$life_exp_wk*0.875)), 75,
                ifelse((age_wk >= (life_exp$life_exp_wk*0.875)) & (age_wk < (life_exp$life_exp_wk*1.125)), 100,
                ifelse((age_wk >= (life_exp$life_exp_wk*1.125)), 125, NA))))))
                ), 

    whendead =  as.factor( 
                ifelse((le_wk >= (life_exp$life_exp_wk*0)) & (le_wk < (life_exp$life_exp_wk*0.125)), 0,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.125)) & (le_wk < (life_exp$life_exp_wk*0.375)), 25,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.375)) & (le_wk < (life_exp$life_exp_wk*0.625)), 50,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.625)) & (le_wk < (life_exp$life_exp_wk*0.875)), 75,
                ifelse((le_wk >= (life_exp$life_exp_wk*0.875)) & (le_wk < (life_exp$life_exp_wk*1.125)), 100,
                ifelse((le_wk >= (life_exp$life_exp_wk*1.125)), 125, NA)))))))
  ) %>% 
  group_by(idno) %>% 
  
  mutate(
    
    age_25 = mean(age_wk[whenvisit == 25], na.rm=T), 
    age_50 = mean(age_wk[whenvisit == 50], na.rm=T), 
    age_75 = mean(age_wk[whenvisit == 75], na.rm=T), 
    age_100 = mean(age_wk[whenvisit == 100], na.rm=T), 
    age_125 = mean(age_wk[whenvisit == 125], na.rm=T) 
    
  ) %>% 
  ungroup() %>% 
  select(-dead_censor, -fu_age_wk, -le_wk, -cod) 

```

## Results 
#### distribution figure with trend line
- generalized additive model for main indices
- linear model with quadratic fit for lifespan groups (survival bias analyses)

```{r slam_ss_dist_funct, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

## creating life expectancy timepoints and labels for function
LE_breaks <- c(life_exp$life_exp_wk*0.25 + 8, life_exp$life_exp_wk*0.50 + 8, life_exp$life_exp_wk*0.75 + 8, life_exp$life_exp_wk*1.00 + 8, life_exp$life_exp_wk*1.25 + 8) 

LE_labels <- c("25%", "50%", "75%", "100%", "125%") 
surv_color <- c("green", "yellow", "red", "purple", "gray")

fcn_plotdist_smooth <- function(metric, x, y, by, xlabel, ylabel, title, xlow, xhigh, peak1, peak2) { 

  #subsetting
  metric2 <- paste0("!is.na(", metric, ")") 
  metric3 <- paste0("df$", metric) 
  df <- base::subset(main_all3, eval(parse(text = metric2))) 
  sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)") 
  z <- Inf

  ggplot(subset(df, eval(parse(text = sub))), aes_string(x = x, y = y, colour = by)) + 
    #background 
    annotate("rect", xmin = life_exp$life_exp_wk*1.25 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*1.25 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.2) + 
    annotate("rect", xmin = life_exp$life_exp_wk*1 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*1 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "purple", alpha = 0.2) + 
    annotate("rect", xmin = life_exp$life_exp_wk*0.75 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.75 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*0.50 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.50 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "orange", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*0.25 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.25 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "green", alpha = 0.2) +
    
    #dots
    geom_point(alpha=0.5, position = "jitter", size = 1, aes_string(colour = by), shape=1) +
    
    #LE lines
    geom_vline(xintercept = life_exp$life_exp_wk *1.25, colour = "black") +
    geom_vline(xintercept = life_exp$life_exp_wk, colour = "purple") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.75, colour = "red") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.5, colour = "orange") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.25, colour = "green") +
    
    #trendlines 
    geom_smooth(aes_string(group = by, linetype = "solid"), linetype = "solid", color = "black", size = 1.8) +
    geom_smooth(aes(linetype = "solid")) + 

    #peaks
    geom_point(aes(x = peak1$all$age_wk, y = peak1$all$fit), size=2, color="black", fill = "white", alpha=0.7, shape=23, stroke=2) +
    geom_point(aes(x = peak2$all$age_wk, y = peak2$all$fit), size=2, color="black", fill = "white", alpha=0.7, shape=23, stroke=2) +
    
    #labels
    xlab(xlabel) + 
    ylab(ylabel) + 
    xlim(xlow, xhigh) +
    
    #second axis
    scale_x_continuous(limits = c(xlow, xhigh), sec.axis = sec_axis(~.*1, breaks = LE_breaks, labels = LE_labels, name="Percent Life Expectancy")) + 
    theme(axis.text.x.top = element_text(angle = 0, margin = margin(t=-2), size = 7, face = "bold", color = "black", vjust = -5), 
        axis.title.x.top = element_text(size = 10, vjust = -1, margin = margin(t=-2)),
        axis.line.x.top = element_line(size = 0, colour = "white"), 
        axis.ticks.x.top = element_blank()) + 
    scale_linetype_manual(labels = c("All"), name = "", values = c("solid", "dashed", "solid"), guide = FALSE)
} 

fcn_plotdist_lm <- function(metric, x, y, by, xlabel, ylabel, title, xlow, xhigh) { 

  #subsetting
  metric2 <- paste0("!is.na(", metric, ")") 
  metric3 <- paste0("df$", metric) 
  main_all3 <- main_all3 %>% filter(as.numeric(whenvisit) <= as.numeric(whendead)) %>% filter(as.numeric(whendead) > 0) 
  df <- base::subset(main_all3, eval(parse(text = metric2))) 
  sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)") 
  z <- Inf

  ggplot(subset(df, eval(parse(text = sub))), aes_string(x = x, y = y, colour = by)) + 
    #background 
    annotate("rect", xmin = life_exp$life_exp_wk*1.25 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*1.25 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*1 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*1 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "purple", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*0.75 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.75 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*0.50 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.50 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "orange", alpha = 0.2) +
    annotate("rect", xmin = life_exp$life_exp_wk*0.25 - 1.96*life_exp$se, xmax = life_exp$life_exp_wk*0.25 + 1.96*life_exp$se, ymin = -Inf, ymax = Inf, fill = "green", alpha = 0.2) +
    
    #LE lines
    geom_vline(xintercept = life_exp$life_exp_wk *1.25, colour = "black") +
    geom_vline(xintercept = life_exp$life_exp_wk, colour = "purple") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.75, colour = "red") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.5, colour = "orange") +
    geom_vline(xintercept = life_exp$life_exp_wk *0.25, colour = "green") +
    
    #trendlines 
    geom_smooth(aes_string(group = by, linetype = "solid"), method = "lm", formula = y ~ poly(x, 3), 
                linetype = "solid", color = "black", size = 1.8, se = FALSE) +
    geom_smooth(aes(linetype = "solid"), method = "lm", formula = y ~ poly(x, 3), size = 1, se = FALSE) + 
    
    #labels
    xlab(xlabel) + 
    ylab(ylabel) +
    xlim(xlow, xhigh) + 
    
    #second axis
    scale_x_continuous(limits = c(xlow, xhigh), sec.axis = sec_axis(~.*1, breaks = LE_breaks, labels = LE_labels, name="Percent Life Expectancy")) + 
    theme(axis.text.x.top = element_text(angle = 0, margin = margin(t=-2), size = 7, face = "bold", color = "black", vjust = -5), 
        axis.title.x.top = element_text(size = 10, vjust = -1, margin = margin(t=-2)),
        axis.line.x.top = element_line(size = 0, colour = "white"), 
        axis.ticks.x.top = element_blank()) + 
    scale_linetype_manual(labels = c("All"), name = "", values = c("solid", "dashed", "solid"), guide = FALSE) + 
    scale_color_manual(values = surv_color, name = "Survival\n(% Life Expectancy)")
} 

``` 

```{r slam_ss_dist_plots, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

cbPalette_dot <- c("#f8766d", "#00b0f6", "black") 


#sex
slam_fig_i1 <- fcn_plotdist_smooth(metric = "gluc", x = "age_wk", xlabel = "Age (weeks)", xlow = 0, xhigh = 150, 
                        y = "gluc", ylabel = "Glucose (mg/dL)", by = "sex", peak1 = gluc_peak_male, peak2 = gluc_peak_female) + 
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)
slam_fig_i2 <- fcn_plotdist_smooth(metric = "bw", x = "age_wk", xlabel = "Age (weeks)", xlow = 0, xhigh = 150, 
                        y = "bw", ylabel = "Body Weight (g)", by = "sex", peak1 = bw_peak_male, peak2 = bw_peak_female) + 
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)
slam_fig_i3 <- fcn_plotdist_smooth(metric = "fat", x = "age_wk", xlabel = "Age (weeks)", xlow = 0, xhigh = 150, 
                        y = "fat", ylabel = "Body Fat (g)", by = "sex", peak1 = fat_peak_male, peak2 = fat_peak_female) + 
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)

#strain
slam_fig_j1 <- fcn_plotdist_smooth(metric = "gluc", x = "age_wk", xlabel = "Age (weeks)", xlow = 0, xhigh = 150, 
                        y = "gluc", ylabel = "Glucose (mg/dL)", by = "strain", peak1 = gluc_peak_HET3, peak2 = gluc_peak_B6) + 
  scale_color_manual(labels = c("B6", "HET3"), name = "Strain", values = cbPalette_dot)
slam_fig_j2 <- fcn_plotdist_smooth(metric = "bw", x = "age_wk", xlabel = "Age (weeks)", xlow = 0, xhigh = 150, 
                        y = "bw", ylabel = "Body Weight (g)", by = "strain", peak1 = bw_peak_HET3, peak2 = bw_peak_B6) + 
  scale_color_manual(labels = c("B6", "HET3"), name = "Strain", values = cbPalette_dot)
slam_fig_j3 <- fcn_plotdist_smooth(metric = "fat", x = "age_wk", xlabel = "Age (weeks)", xlow = 0, xhigh = 150, 
                        y = "fat", ylabel = "Body Fat (g)", by = "strain", peak1 = fat_peak_HET3, peak2 = fat_peak_B6) + 
  scale_color_manual(labels = c("B6", "HET3"), name = "Strain", values = cbPalette_dot)

#death
slam_fig_d1 <- fcn_plotdist_lm(metric = "gluc", x = "age_wk", xlabel = "Age (weeks)", xlow = 0, xhigh = 150, 
                        y = "gluc", ylabel = "Glucose (mg/dL)", by = "whendead")
slam_fig_d2 <- fcn_plotdist_lm(metric = "bw", x = "age_wk", xlabel = "Age (weeks)", xlow = 0, xhigh = 150, 
                        y = "bw", ylabel = "Body Weight (g)", by = "whendead") 
slam_fig_d3 <- fcn_plotdist_lm(metric = "fat", x = "age_wk", xlabel = "Age (weeks)", xlow = 0, xhigh = 150, 
                        y = "fat", ylabel = "Body Fat (g)", by = "whendead") 
 
#
slam_fig_i13 <- ggarrange(slam_fig_i1, slam_fig_i2, slam_fig_i3, 
                      ncol = 2, nrow = 2, 
                     common.legend = TRUE, 
                     legend = "right") 

slam_fig_j13 <- ggarrange(slam_fig_j1, slam_fig_j2, slam_fig_j3, 
                      ncol = 2, nrow = 2, 
                     common.legend = TRUE, 
                     legend = "right") 

slam_fig_d13 <- ggarrange(slam_fig_d1, slam_fig_d2, slam_fig_d3, 
                      ncol = 2, nrow = 2, 
                     common.legend = TRUE, 
                     legend = "right") 
``` 

```{r slam_sex_dist_fig, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

annotate_figure(slam_fig_i13, 
                top = text_grob("Trends in Metabolic indices over the Lifespan by Sex"), 
                bottom = text_grob("Note: \nHorizontal lines indicate percent of life expectancy: green (25%), orange (50%), red (75%), purple (100%), and black (125%) \nDiamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines")) 
                ) 
``` 
 
*** 

```{r slam_strain_dist_fig, warning=FALSE, message=FALSE, echo=FALSE}

annotate_figure(slam_fig_j13, 
                top = text_grob("Trends in Metabolic indices over the Lifespan by Strain"), 
                bottom = text_grob("Note: \nHorizontal lines indicate percent of life expectancy: green (25%), orange (50%), red (75%), purple (100%), and black (125%) \nDiamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines")) 
                ) 
``` 

*** 

#### pre and post-peak change creation 

```{r slam_phase_creation, warning=FALSE, message=FALSE, echo=FALSE}

## assuming peak indices value is fixed (sd = 0) for rate standard deviation calculations

all_gluc <- NULL
f_gluc <- NULL
m_gluc <- NULL

#all
all_gluc$pretime <- (gluc_peak$all$le - gluc_peak$lowbound$le) 
all_gluc$prepeak_diff <- (gluc_peak$all$fit - gluc_peak$lowbound$fit)
all_gluc$prepeak_rate <- ((all_gluc$prepeak_diff / gluc_peak$lowbound$fit)/all_gluc$pretime)*5*100
all_gluc$prepeak_diff_lcl <- all_gluc$prepeak_diff - 1.96*gluc_peak$lowbound$se.fit
all_gluc$prepeak_diff_ucl <- all_gluc$prepeak_diff + 1.96*gluc_peak$lowbound$se.fit
all_gluc$prepeak_rate_lcl <- ((all_gluc$prepeak_diff_lcl / gluc_peak$lowbound$fit)/all_gluc$pretime)*5*100
all_gluc$prepeak_rate_ucl <- ((all_gluc$prepeak_diff_ucl / gluc_peak$lowbound$fit)/all_gluc$pretime)*5*100
all_gluc$prepeak_rate_se <- (all_gluc$prepeak_rate_ucl - all_gluc$prepeak_rate)/1.96
all_gluc$prepeak_rate_n <- gluc_peak$lowbound$n
all_gluc$prepeak_rate_sd <- all_gluc$prepeak_rate_se*sqrt(all_gluc$prepeak_rate_n) 

#female
f_gluc$pretime <- (gluc_peak_female$all$le - gluc_peak_female$lowbound$le) 
f_gluc$prepeak_diff <- (gluc_peak_female$all$fit - gluc_peak_female$lowbound$fit)
f_gluc$prepeak_rate <- ((f_gluc$prepeak_diff / gluc_peak_female$lowbound$fit)/f_gluc$pretime)*5*100
f_gluc$prepeak_diff_lcl <- f_gluc$prepeak_diff - 1.96*gluc_peak_female$lowbound$se.fit
f_gluc$prepeak_diff_ucl <- f_gluc$prepeak_diff + 1.96*gluc_peak_female$lowbound$se.fit
f_gluc$prepeak_rate_lcl <- ((f_gluc$prepeak_diff_lcl / gluc_peak_female$lowbound$fit)/f_gluc$pretime)*5*100
f_gluc$prepeak_rate_ucl <- ((f_gluc$prepeak_diff_ucl / gluc_peak_female$lowbound$fit)/f_gluc$pretime)*5*100
f_gluc$prepeak_rate_se <- (f_gluc$prepeak_rate_ucl - f_gluc$prepeak_rate)/1.96
f_gluc$prepeak_rate_n <- gluc_peak_female$lowbound$n
f_gluc$prepeak_rate_sd <- f_gluc$prepeak_rate_se*sqrt(f_gluc$prepeak_rate_n) 

#male
m_gluc$pretime <- (gluc_peak_male$all$le - gluc_peak_male$lowbound$le) 
m_gluc$prepeak_diff <- (gluc_peak_male$all$fit - gluc_peak_male$lowbound$fit)
m_gluc$prepeak_rate <- ((m_gluc$prepeak_diff / gluc_peak_male$lowbound$fit)/m_gluc$pretime)*5*100
m_gluc$prepeak_diff_lcl <- m_gluc$prepeak_diff - 1.96*gluc_peak_male$lowbound$se.fit
m_gluc$prepeak_diff_ucl <- m_gluc$prepeak_diff + 1.96*gluc_peak_male$lowbound$se.fit
m_gluc$prepeak_rate_lcl <- ((m_gluc$prepeak_diff_lcl / gluc_peak_male$lowbound$fit)/m_gluc$pretime)*5*100
m_gluc$prepeak_rate_ucl <- ((m_gluc$prepeak_diff_ucl / gluc_peak_male$lowbound$fit)/m_gluc$pretime)*5*100
m_gluc$prepeak_rate_se <- (m_gluc$prepeak_rate_ucl - m_gluc$prepeak_rate)/1.96
m_gluc$prepeak_rate_n <- gluc_peak_male$lowbound$n
m_gluc$prepeak_rate_sd <- m_gluc$prepeak_rate_se*sqrt(m_gluc$prepeak_rate_n) 

#all
all_gluc$posttime <- (gluc_peak$highbound$le - gluc_peak$all$le) 
all_gluc$postpeak_diff <- (gluc_peak$highbound$fit - gluc_peak$all$fit)
all_gluc$postpeak_rate <- ((all_gluc$postpeak_diff / gluc_peak$all$fit)/all_gluc$posttime)*5*100
all_gluc$postpeak_diff_lcl <- all_gluc$postpeak_diff - 1.96*gluc_peak$highbound$se.fit
all_gluc$postpeak_diff_ucl <- all_gluc$postpeak_diff + 1.96*gluc_peak$highbound$se.fit
all_gluc$postpeak_rate_lcl <- ((all_gluc$postpeak_diff_lcl / gluc_peak$all$fit)/all_gluc$posttime)*5*100
all_gluc$postpeak_rate_ucl <- ((all_gluc$postpeak_diff_ucl / gluc_peak$all$fit)/all_gluc$posttime)*5*100
all_gluc$postpeak_rate_se <- (all_gluc$postpeak_rate_ucl - all_gluc$postpeak_rate)/1.96
all_gluc$postpeak_rate_n <- gluc_peak$highbound$n 
all_gluc$postpeak_rate_sd <- all_gluc$postpeak_rate_se*sqrt(all_gluc$postpeak_rate_n) 

#female
f_gluc$posttime <- (gluc_peak_female$highbound$le - gluc_peak_female$all$le) 
f_gluc$postpeak_diff <- (gluc_peak_female$highbound$fit - gluc_peak_female$all$fit) 
f_gluc$postpeak_rate <- ((f_gluc$postpeak_diff / gluc_peak_female$all$fit)/f_gluc$posttime)*5*100 
f_gluc$postpeak_diff_lcl <- f_gluc$postpeak_diff - 1.96*gluc_peak_female$highbound$se.fit 
f_gluc$postpeak_diff_ucl <- f_gluc$postpeak_diff + 1.96*gluc_peak_female$highbound$se.fit 
f_gluc$postpeak_rate_lcl <- ((f_gluc$postpeak_diff_lcl / gluc_peak_female$all$fit)/f_gluc$posttime)*5*100 
f_gluc$postpeak_rate_ucl <- ((f_gluc$postpeak_diff_ucl / gluc_peak_female$all$fit)/f_gluc$posttime)*5*100 
f_gluc$postpeak_rate_se <- (f_gluc$postpeak_rate_ucl - f_gluc$postpeak_rate)/1.96
f_gluc$postpeak_rate_n <- gluc_peak_female$highbound$n 
f_gluc$postpeak_rate_sd <- f_gluc$postpeak_rate_se*sqrt(f_gluc$postpeak_rate_n) 

#male
m_gluc$posttime <- (gluc_peak_male$highbound$le - gluc_peak_male$all$le) 
m_gluc$postpeak_diff <- (gluc_peak_male$highbound$fit - gluc_peak_male$all$fit) 
m_gluc$postpeak_rate <- ((m_gluc$postpeak_diff / gluc_peak_male$all$fit)/m_gluc$posttime)*5*100 
m_gluc$postpeak_diff_lcl <- m_gluc$postpeak_diff - 1.96*gluc_peak_male$highbound$se.fit 
m_gluc$postpeak_diff_ucl <- m_gluc$postpeak_diff + 1.96*gluc_peak_male$highbound$se.fit 
m_gluc$postpeak_rate_lcl <- ((m_gluc$postpeak_diff_lcl / gluc_peak_male$all$fit)/m_gluc$posttime)*5*100 
m_gluc$postpeak_rate_ucl <- ((m_gluc$postpeak_diff_ucl / gluc_peak_male$all$fit)/m_gluc$posttime)*5*100 
m_gluc$postpeak_rate_se <- (m_gluc$postpeak_rate_ucl - m_gluc$postpeak_rate)/1.96
m_gluc$postpeak_rate_n <- gluc_peak_male$highbound$n 
m_gluc$postpeak_rate_sd <- m_gluc$postpeak_rate_se*sqrt(m_gluc$postpeak_rate_n) 

gluc2 <- as.data.frame(bind_rows(all_gluc, f_gluc, m_gluc)) %>% 
  mutate(group = c("all", "female", "male"), 
         index = c("gluc", "gluc", "gluc")) 

all_bw <- NULL
f_bw <- NULL
m_bw <- NULL

#all
all_bw$pretime <- (bw_peak$all$le - bw_peak$lowbound$le) 
all_bw$prepeak_diff <- (bw_peak$all$fit - bw_peak$lowbound$fit)
all_bw$prepeak_rate <- ((all_bw$prepeak_diff / bw_peak$lowbound$fit)/all_bw$pretime)*5*100
all_bw$prepeak_diff_lcl <- all_bw$prepeak_diff - 1.96*bw_peak$lowbound$se.fit
all_bw$prepeak_diff_ucl <- all_bw$prepeak_diff + 1.96*bw_peak$lowbound$se.fit
all_bw$prepeak_rate_lcl <- ((all_bw$prepeak_diff_lcl / bw_peak$lowbound$fit)/all_bw$pretime)*5*100
all_bw$prepeak_rate_ucl <- ((all_bw$prepeak_diff_ucl / bw_peak$lowbound$fit)/all_bw$pretime)*5*100
all_bw$prepeak_rate_se <- (all_bw$prepeak_rate_ucl - all_bw$prepeak_rate)/1.96
all_bw$prepeak_rate_n <- bw_peak$lowbound$n
all_bw$prepeak_rate_sd <- all_bw$prepeak_rate_se*sqrt(all_bw$prepeak_rate_n) 

#female
f_bw$pretime <- (bw_peak_female$all$le - bw_peak_female$lowbound$le) 
f_bw$prepeak_diff <- (bw_peak_female$all$fit - bw_peak_female$lowbound$fit)
f_bw$prepeak_rate <- ((f_bw$prepeak_diff / bw_peak_female$lowbound$fit)/f_bw$pretime)*5*100
f_bw$prepeak_diff_lcl <- f_bw$prepeak_diff - 1.96*bw_peak_female$lowbound$se.fit
f_bw$prepeak_diff_ucl <- f_bw$prepeak_diff + 1.96*bw_peak_female$lowbound$se.fit
f_bw$prepeak_rate_lcl <- ((f_bw$prepeak_diff_lcl / bw_peak_female$lowbound$fit)/f_bw$pretime)*5*100
f_bw$prepeak_rate_ucl <- ((f_bw$prepeak_diff_ucl / bw_peak_female$lowbound$fit)/f_bw$pretime)*5*100
f_bw$prepeak_rate_se <- (f_bw$prepeak_rate_ucl - f_bw$prepeak_rate)/1.96
f_bw$prepeak_rate_n <- bw_peak_female$lowbound$n
f_bw$prepeak_rate_sd <- f_bw$prepeak_rate_se*sqrt(f_bw$prepeak_rate_n) 

#male
m_bw$pretime <- (bw_peak_male$all$le - bw_peak_male$lowbound$le) 
m_bw$prepeak_diff <- (bw_peak_male$all$fit - bw_peak_male$lowbound$fit)
m_bw$prepeak_rate <- ((m_bw$prepeak_diff / bw_peak_male$lowbound$fit)/m_bw$pretime)*5*100
m_bw$prepeak_diff_lcl <- m_bw$prepeak_diff - 1.96*bw_peak_male$lowbound$se.fit
m_bw$prepeak_diff_ucl <- m_bw$prepeak_diff + 1.96*bw_peak_male$lowbound$se.fit
m_bw$prepeak_rate_lcl <- ((m_bw$prepeak_diff_lcl / bw_peak_male$lowbound$fit)/m_bw$pretime)*5*100
m_bw$prepeak_rate_ucl <- ((m_bw$prepeak_diff_ucl / bw_peak_male$lowbound$fit)/m_bw$pretime)*5*100
m_bw$prepeak_rate_se <- (m_bw$prepeak_rate_ucl - m_bw$prepeak_rate)/1.96
m_bw$prepeak_rate_n <- bw_peak_male$lowbound$n
m_bw$prepeak_rate_sd <- m_bw$prepeak_rate_se*sqrt(m_bw$prepeak_rate_n) 

#all
all_bw$posttime <- (bw_peak$highbound$le - bw_peak$all$le) 
all_bw$postpeak_diff <- (bw_peak$highbound$fit - bw_peak$all$fit)
all_bw$postpeak_rate <- ((all_bw$postpeak_diff / bw_peak$all$fit)/all_bw$posttime)*5*100
all_bw$postpeak_diff_lcl <- all_bw$postpeak_diff - 1.96*bw_peak$highbound$se.fit
all_bw$postpeak_diff_ucl <- all_bw$postpeak_diff + 1.96*bw_peak$highbound$se.fit
all_bw$postpeak_rate_lcl <- ((all_bw$postpeak_diff_lcl / bw_peak$all$fit)/all_bw$posttime)*5*100
all_bw$postpeak_rate_ucl <- ((all_bw$postpeak_diff_ucl / bw_peak$all$fit)/all_bw$posttime)*5*100
all_bw$postpeak_rate_se <- (all_bw$postpeak_rate_ucl - all_bw$postpeak_rate)/1.96
all_bw$postpeak_rate_n <- bw_peak$highbound$n 
all_bw$postpeak_rate_sd <- all_bw$postpeak_rate_se*sqrt(all_bw$postpeak_rate_n) 

#female
f_bw$posttime <- (bw_peak_female$highbound$le - bw_peak_female$all$le) 
f_bw$postpeak_diff <- (bw_peak_female$highbound$fit - bw_peak_female$all$fit) 
f_bw$postpeak_rate <- ((f_bw$postpeak_diff / bw_peak_female$all$fit)/f_bw$posttime)*5*100 
f_bw$postpeak_diff_lcl <- f_bw$postpeak_diff - 1.96*bw_peak_female$highbound$se.fit 
f_bw$postpeak_diff_ucl <- f_bw$postpeak_diff + 1.96*bw_peak_female$highbound$se.fit 
f_bw$postpeak_rate_lcl <- ((f_bw$postpeak_diff_lcl / bw_peak_female$all$fit)/f_bw$posttime)*5*100 
f_bw$postpeak_rate_ucl <- ((f_bw$postpeak_diff_ucl / bw_peak_female$all$fit)/f_bw$posttime)*5*100 
f_bw$postpeak_rate_se <- (f_bw$postpeak_rate_ucl - f_bw$postpeak_rate)/1.96
f_bw$postpeak_rate_n <- bw_peak_female$highbound$n 
f_bw$postpeak_rate_sd <- f_bw$postpeak_rate_se*sqrt(f_bw$postpeak_rate_n) 

#male
m_bw$posttime <- (bw_peak_male$highbound$le - bw_peak_male$all$le) 
m_bw$postpeak_diff <- (bw_peak_male$highbound$fit - bw_peak_male$all$fit) 
m_bw$postpeak_rate <- ((m_bw$postpeak_diff / bw_peak_male$all$fit)/m_bw$posttime)*5*100 
m_bw$postpeak_diff_lcl <- m_bw$postpeak_diff - 1.96*bw_peak_male$highbound$se.fit 
m_bw$postpeak_diff_ucl <- m_bw$postpeak_diff + 1.96*bw_peak_male$highbound$se.fit 
m_bw$postpeak_rate_lcl <- ((m_bw$postpeak_diff_lcl / bw_peak_male$all$fit)/m_bw$posttime)*5*100 
m_bw$postpeak_rate_ucl <- ((m_bw$postpeak_diff_ucl / bw_peak_male$all$fit)/m_bw$posttime)*5*100 
m_bw$postpeak_rate_se <- (m_bw$postpeak_rate_ucl - m_bw$postpeak_rate)/1.96
m_bw$postpeak_rate_n <- bw_peak_male$highbound$n 
m_bw$postpeak_rate_sd <- m_bw$postpeak_rate_se*sqrt(m_bw$postpeak_rate_n) 

bw2 <- as.data.frame(bind_rows(all_bw, f_bw, m_bw)) %>% 
  mutate(group = c("all", "female", "male"), 
         index = c("bw", "bw", "bw")) 

all_fat <- NULL
f_fat <- NULL
m_fat <- NULL

#all
all_fat$pretime <- (fat_peak$all$le - fat_peak$lowbound$le) 
all_fat$prepeak_diff <- (fat_peak$all$fit - fat_peak$lowbound$fit)
all_fat$prepeak_rate <- ((all_fat$prepeak_diff / fat_peak$lowbound$fit)/all_fat$pretime)*5*100
all_fat$prepeak_diff_lcl <- all_fat$prepeak_diff - 1.96*fat_peak$lowbound$se.fit
all_fat$prepeak_diff_ucl <- all_fat$prepeak_diff + 1.96*fat_peak$lowbound$se.fit

all_fat$prepeak_rate_lcl <- ((all_fat$prepeak_diff_lcl / fat_peak$lowbound$fit)/all_fat$pretime)*5*100
all_fat$prepeak_rate_ucl <- ((all_fat$prepeak_diff_ucl / fat_peak$lowbound$fit)/all_fat$pretime)*5*100

all_fat$prepeak_rate_se <- (all_fat$prepeak_rate_ucl - all_fat$prepeak_rate)/1.96
all_fat$prepeak_rate_n <- fat_peak$lowbound$n
all_fat$prepeak_rate_sd <- all_fat$prepeak_rate_se*sqrt(all_fat$prepeak_rate_n) 

#female
f_fat$pretime <- (fat_peak_female$all$le - fat_peak_female$lowbound$le) 
f_fat$prepeak_diff <- (fat_peak_female$all$fit - fat_peak_female$lowbound$fit)
f_fat$prepeak_rate <- ((f_fat$prepeak_diff / fat_peak_female$lowbound$fit)/f_fat$pretime)*5*100
f_fat$prepeak_diff_lcl <- f_fat$prepeak_diff - 1.96*fat_peak_female$lowbound$se.fit
f_fat$prepeak_diff_ucl <- f_fat$prepeak_diff + 1.96*fat_peak_female$lowbound$se.fit
f_fat$prepeak_rate_lcl <- ((f_fat$prepeak_diff_lcl / fat_peak_female$lowbound$fit)/f_fat$pretime)*5*100
f_fat$prepeak_rate_ucl <- ((f_fat$prepeak_diff_ucl / fat_peak_female$lowbound$fit)/f_fat$pretime)*5*100
f_fat$prepeak_rate_se <- (f_fat$prepeak_rate_ucl - f_fat$prepeak_rate)/1.96
f_fat$prepeak_rate_n <- fat_peak_female$lowbound$n
f_fat$prepeak_rate_sd <- f_fat$prepeak_rate_se*sqrt(f_fat$prepeak_rate_n) 

#male
m_fat$pretime <- (fat_peak_male$all$le - fat_peak_male$lowbound$le) 
m_fat$prepeak_diff <- (fat_peak_male$all$fit - fat_peak_male$lowbound$fit)
m_fat$prepeak_rate <- ((m_fat$prepeak_diff / fat_peak_male$lowbound$fit)/m_fat$pretime)*5*100
m_fat$prepeak_diff_lcl <- m_fat$prepeak_diff - 1.96*fat_peak_male$lowbound$se.fit
m_fat$prepeak_diff_ucl <- m_fat$prepeak_diff + 1.96*fat_peak_male$lowbound$se.fit
m_fat$prepeak_rate_lcl <- ((m_fat$prepeak_diff_lcl / fat_peak_male$lowbound$fit)/m_fat$pretime)*5*100
m_fat$prepeak_rate_ucl <- ((m_fat$prepeak_diff_ucl / fat_peak_male$lowbound$fit)/m_fat$pretime)*5*100
m_fat$prepeak_rate_se <- (m_fat$prepeak_rate_ucl - m_fat$prepeak_rate)/1.96
m_fat$prepeak_rate_n <- fat_peak_male$lowbound$n
m_fat$prepeak_rate_sd <- m_fat$prepeak_rate_se*sqrt(m_fat$prepeak_rate_n) 

#all
all_fat$posttime <- (fat_peak$highbound$le - fat_peak$all$le) 
all_fat$postpeak_diff <- (fat_peak$highbound$fit - fat_peak$all$fit)
all_fat$postpeak_rate <- ((all_fat$postpeak_diff / fat_peak$all$fit)/all_fat$posttime)*5*100
all_fat$postpeak_diff_lcl <- all_fat$postpeak_diff - 1.96*fat_peak$all$se.fit
all_fat$postpeak_diff_ucl <- all_fat$postpeak_diff + 1.96*fat_peak$all$se.fit
all_fat$postpeak_rate_lcl <- ((all_fat$postpeak_diff_lcl / fat_peak$all$fit)/all_fat$posttime)*5*100
all_fat$postpeak_rate_ucl <- ((all_fat$postpeak_diff_ucl / fat_peak$all$fit)/all_fat$posttime)*5*100
all_fat$postpeak_rate_se <- (all_fat$postpeak_rate_ucl - all_fat$postpeak_rate)/1.96
all_fat$postpeak_rate_n <- fat_peak$highbound$n 
all_fat$postpeak_rate_sd <- all_fat$postpeak_rate_se*sqrt(all_fat$postpeak_rate_n) 

#female
f_fat$posttime <- (fat_peak_female$highbound$le - fat_peak_female$all$le) 
f_fat$postpeak_diff <- (fat_peak_female$highbound$fit - fat_peak_female$all$fit) 
f_fat$postpeak_rate <- ((f_fat$postpeak_diff / fat_peak_female$all$fit)/f_fat$posttime)*5*100 
f_fat$postpeak_diff_lcl <- f_fat$postpeak_diff - 1.96*fat_peak_female$highbound$se.fit 
f_fat$postpeak_diff_ucl <- f_fat$postpeak_diff + 1.96*fat_peak_female$highbound$se.fit 
f_fat$postpeak_rate_lcl <- ((f_fat$postpeak_diff_lcl / fat_peak_female$all$fit)/f_fat$posttime)*5*100 
f_fat$postpeak_rate_ucl <- ((f_fat$postpeak_diff_ucl / fat_peak_female$all$fit)/f_fat$posttime)*5*100 
f_fat$postpeak_rate_se <- (f_fat$postpeak_rate_ucl - f_fat$postpeak_rate)/1.96
f_fat$postpeak_rate_n <- fat_peak_female$highbound$n 
f_fat$postpeak_rate_sd <- f_fat$postpeak_rate_se*sqrt(f_fat$postpeak_rate_n) 

#male
m_fat$posttime <- (fat_peak_male$highbound$le - fat_peak_male$all$le) 
m_fat$postpeak_diff <- (fat_peak_male$highbound$fit - fat_peak_male$all$fit) 
m_fat$postpeak_rate <- ((m_fat$postpeak_diff / fat_peak_male$all$fit)/m_fat$posttime)*5*100 
m_fat$postpeak_diff_lcl <- m_fat$postpeak_diff - 1.96*fat_peak_male$highbound$se.fit 
m_fat$postpeak_diff_ucl <- m_fat$postpeak_diff + 1.96*fat_peak_male$highbound$se.fit 
m_fat$postpeak_rate_lcl <- ((m_fat$postpeak_diff_lcl / fat_peak_male$all$fit)/m_fat$posttime)*5*100 
m_fat$postpeak_rate_ucl <- ((m_fat$postpeak_diff_ucl / fat_peak_male$all$fit)/m_fat$posttime)*5*100 
m_fat$postpeak_rate_se <- (m_fat$postpeak_rate_ucl - m_fat$postpeak_rate)/1.96
m_fat$postpeak_rate_n <- fat_peak_male$highbound$n 
m_fat$postpeak_rate_sd <- m_fat$postpeak_rate_se*sqrt(m_fat$postpeak_rate_n) 

fat2 <- as.data.frame(bind_rows(all_fat, f_fat, m_fat)) %>% 
  mutate(group = c("all", "female", "male"), 
         index = c("fat", "fat", "fat")) 


##
phases <- rbind(gluc2, bw2, fat2) %>% 
  mutate(
    
    prepeak_pval = tsum.test(prepeak_rate, prepeak_rate_sd, prepeak_rate_n, mu = 0, conf.level = 0.95)$p.value,
    postpeak_pval = tsum.test(postpeak_rate, postpeak_rate_sd, postpeak_rate_n, mu = 0, conf.level = 0.95)$p.value,

    prepeak_mean_95CI = 
      ifelse(prepeak_pval < 0.050,
    paste0(format(round(prepeak_rate, 2), nsmall=2)," ","(",format(round(prepeak_rate_lcl, 2), nsmall=2),", ",format(round(prepeak_rate_ucl, 2), nsmall=2),")*"),
    paste0(format(round(prepeak_rate, 2), nsmall=2)," ","(",format(round(prepeak_rate_lcl, 2), nsmall=2),", ",format(round(prepeak_rate_ucl, 2), nsmall=2),")")), 
    
    postpeak_mean_95CI = 
      ifelse(postpeak_pval < 0.050,
    paste0(format(round(postpeak_rate, 2), nsmall=2)," ","(",format(round(postpeak_rate_lcl, 2), nsmall=2),", ",format(round(postpeak_rate_ucl, 2), nsmall=2),")*"),
    paste0(format(round(postpeak_rate, 2), nsmall=2)," ","(",format(round(postpeak_rate_lcl, 2), nsmall=2),", ",format(round(postpeak_rate_ucl, 2), nsmall=2),")"))
    
  )

gluc_prepeak <- phases %>% 
  filter(index == "gluc") %>% 
  select(index, group, contains("95"), -contains("postpeak")) %>% 
  spread(key = group, value = prepeak_mean_95CI) 
gluc_postpeak <- phases %>% 
  filter(index == "gluc") %>% 
  select(index, group, contains("95"), -contains("prepeak")) %>% 
  spread(key = group, value = postpeak_mean_95CI) 

bw_prepeak <- phases %>% 
  filter(index == "bw") %>% 
  select(index, group, contains("95"), -contains("postpeak")) %>% 
  spread(key = group, value = prepeak_mean_95CI) 
bw_postpeak <- phases %>% 
  filter(index == "bw") %>% 
  select(index, group, contains("95"), -contains("prepeak")) %>% 
  spread(key = group, value = postpeak_mean_95CI) 

fat_prepeak <- phases %>% 
  filter(index == "fat") %>% 
  select(index, group, contains("95"), -contains("postpeak")) %>% 
  spread(key = group, value = prepeak_mean_95CI) 
fat_postpeak <- phases %>% 
  filter(index == "fat") %>% 
  select(index, group, contains("95"), -contains("prepeak")) %>% 
  spread(key = group, value = postpeak_mean_95CI) 

gluc_all <- bind_cols(gluc_prepeak, gluc_postpeak)
bw_all <- bind_cols(bw_prepeak, bw_postpeak)
fat_all <- bind_cols(fat_prepeak, fat_postpeak)

phases2 <- bind_rows(gluc_all, bw_all, fat_all) 
colnames(phases2) <- c("index", "all", "female", "male", "index2", "all2", "female2", "male2")

##peaks

#all
gluc_peak_ci <- paste0(round(gluc_peak$all$le, 2)," (", round(gluc_peak$min$le, 2), ", ", round(gluc_peak$max$le, 2), ")") 
bw_peak_ci <- paste0(round(bw_peak$all$le, 2)," (", round(bw_peak$min$le, 2), ", ", round(bw_peak$max$le, 2), ")") 
fat_peak_ci <- paste0(round(fat_peak$all$le, 2)," (", round(fat_peak$min$le, 2), ", ", round(fat_peak$max$le, 2), ")") 

#male
male_gluc_peak_ci <- paste0(round(gluc_peak_male$all$le, 2)," (", round(gluc_peak_male$min$le, 2), ", ", round(gluc_peak_male$max$le, 2), ")") 
male_bw_peak_ci <- paste0(round(bw_peak_male$all$le, 2)," (", round(bw_peak_male$min$le, 2), ", ", round(bw_peak_male$max$le, 2), ")") 
male_fat_peak_ci <- paste0(round(fat_peak_male$all$le, 2)," (", round(fat_peak_male$min$le, 2), ", ", round(fat_peak_male$max$le, 2), ")") 

#female
female_gluc_peak_ci <- paste0(round(gluc_peak_female$all$le, 2)," (", round(gluc_peak_female$min$le, 2), ", ", round(gluc_peak_female$max$le, 2), ")") 
female_bw_peak_ci <- paste0(round(bw_peak_female$all$le, 2)," (", round(bw_peak_female$min$le, 2), ", ", round(bw_peak_female$max$le, 2), ")") 
female_fat_peak_ci <- paste0(round(fat_peak_female$all$le, 2)," (", round(fat_peak_female$min$le, 2), ", ", round(fat_peak_female$max$le, 2), ")") 

#final datasets
phases2$peak_all <- c(gluc_peak_ci, bw_peak_ci, fat_peak_ci) 
phases2$peak_female <- c(female_gluc_peak_ci, female_bw_peak_ci, female_fat_peak_ci) 
phases2$peak_male <- c(male_gluc_peak_ci, male_bw_peak_ci, male_fat_peak_ci)

phases3 <- phases2 %>% 
  select(index, contains("peak"), everything(), -index2) 

```

```{r slam_phase_formatting, warning=FALSE, message=FALSE, echo=FALSE}

#html output formatting
phases4 <- phases3 %>% 
  mutate( 
    peak_all = noquote((gsub("\\(", "<br>(", peak_all))), 
    peak_female = noquote((gsub("\\(", "<br>(", peak_female))), 
    peak_male = noquote((gsub("\\(", "<br>(", peak_male))), 
    all = noquote((gsub("\\(", "<br>(", all))), 
    female = noquote((gsub("\\(", "<br>(", female))), 
    male = noquote((gsub("\\(", "<br>(", male))), 
    all2 = noquote((gsub("\\(", "<br>(", all2))), 
    female2 = noquote((gsub("\\(", "<br>(", female2))), 
    male2 = noquote((gsub("\\(", "<br>(", male2)))
    ) 

slam_phases <- phases4 

```

#### change in metabolic indices pre and post-peak 
- rate of percent change 

```{r slam_phase_table_prate, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

slam_phases_out <- kable(slam_phases, row.names = FALSE, col.names = c("Metabolic <br> Index", "All", "Female", "Male", "All", "Female", "Male", "All", "Female", "Male"), align = c("c","c","c","c","c","c","c"), escape = FALSE, 
      caption = "Rate of Percent Change for Metabolic indices Pre- and Post-peak") %>% 
  kable_styling(bootstrap_options = "striped", full_width = TRUE, position = "left", font_size = 12) %>% 
  column_spec(1, bold = TRUE, border_right = TRUE) %>% 
  column_spec(4, border_right = TRUE) %>% 
  column_spec(7, border_right = TRUE) %>% 
  add_header_above(c(" " = 1, "Age at Peak Value$^$ \n(% life expectancy)" = 3, "Pre-Peak$^$$^$" = 3, "Post-Peak$^$$^$" = 3)) %>% 
  footnote(symbol = c(
    "p value < 0.05",
    "Mean (range of ages at similar values) of age at peak index value", 
    "Mean (95% confidence interval) rate of percent change during phase per five percent change in life expectancy", 
    "Pre-peak is defined as 25% life expectancy to age at peak value; post-peak phase is defined as peak to 125% life expectancy", 
    "Male and female estimates are based on sex specific life expectancy")) 

slam_phases_out 

```

*** 

#### quartile creation

```{r slam_scaling, warning=FALSE, message=FALSE, echo=FALSE}

main_cont_surv <- main_all3 %>% select(idno, sex, strain, age_25, age_50, age_75, age_100, age_125) %>% distinct() %>% inner_join(census_c16_death) 

main_cat_surv_f <- main_cat2_f %>% inner_join(main_cont_surv) %>% filter(sex == "F") 
main_cat_surv_m <- main_cat2_m %>% inner_join(main_cont_surv) %>% filter(sex == "M") 

```
 
```{r slam_quartiles, warning=FALSE, message=FALSE, echo=FALSE}

quart_breaks = c(0, 0.25, 0.50, 0.75, 1.0) 
quart_labs = c("Low", "Medlow", "Medhigh", "High")

## female
main_cat_surv_f$gluc_50_quart <- cut(main_cat_surv_f$cfgluc_50_m, quantile(main_cat_surv_f$cfgluc_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$gluc_75_quart <- cut(main_cat_surv_f$cfgluc_75_m, quantile(main_cat_surv_f$cfgluc_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$gluc_100_quart <- cut(main_cat_surv_f$cfgluc_100_m, quantile(main_cat_surv_f$cfgluc_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_f$bw_50_quart <- cut(main_cat_surv_f$cfbw_50_m, quantile(main_cat_surv_f$cfbw_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$bw_75_quart <- cut(main_cat_surv_f$cfbw_75_m, quantile(main_cat_surv_f$cfbw_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$bw_100_quart <- cut(main_cat_surv_f$cfbw_100_m, quantile(main_cat_surv_f$cfbw_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_f$fat_50_quart <- cut(main_cat_surv_f$cffat_50_m, quantile(main_cat_surv_f$cffat_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$fat_75_quart <- cut(main_cat_surv_f$cffat_75_m, quantile(main_cat_surv_f$cffat_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$fat_100_quart <- cut(main_cat_surv_f$cffat_100_m, quantile(main_cat_surv_f$cffat_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

## male
main_cat_surv_m$gluc_50_quart <- cut(main_cat_surv_m$cfgluc_50_m, quantile(main_cat_surv_m$cfgluc_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$gluc_75_quart <- cut(main_cat_surv_m$cfgluc_75_m, quantile(main_cat_surv_m$cfgluc_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$gluc_100_quart <- cut(main_cat_surv_m$cfgluc_100_m, quantile(main_cat_surv_m$cfgluc_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_m$bw_50_quart <- cut(main_cat_surv_m$cfbw_50_m, quantile(main_cat_surv_m$cfbw_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$bw_75_quart <- cut(main_cat_surv_m$cfbw_75_m, quantile(main_cat_surv_m$cfbw_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$bw_100_quart <- cut(main_cat_surv_m$cfbw_100_m, quantile(main_cat_surv_m$cfbw_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_m$fat_50_quart <- cut(main_cat_surv_m$cffat_50_m, quantile(main_cat_surv_m$cffat_50_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$fat_75_quart <- cut(main_cat_surv_m$cffat_75_m, quantile(main_cat_surv_m$cffat_75_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$fat_100_quart <- cut(main_cat_surv_m$cffat_100_m, quantile(main_cat_surv_m$cffat_100_m, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

#create
main_cat_surv <- main_cat_surv_f %>% 
  bind_rows(main_cat_surv_m) 

```

#### kaplan meier survival curves 
- comparing highest to lowest quartile 
- three timepoints: 50, 75, and 100% life expectancy 

```{r slam_surv_curve, warning=FALSE, message=FALSE, echo=FALSE}

surv_color <- c("orange", "purple")

km_funct <- function(quart, exp_var, start, end, data, title, env = globalenv()){

#subset data
surv_df <- as.data.frame(data) 
surv_df_plot <- subset(data, eval(parse(text = quart)) %in% c("Low", "High"))
surv_df_plot2 <- as.data.frame(surv_df_plot) 

#change level order for legend
surv_df_plot3 <- surv_df_plot2 %>% 
  mutate_at( 
    vars(contains("quart")), 
    funs(fct_rev(.)) 
  ) 

#pull input variable for quartiles
env$quart <- quart 
env$explanatory <- NULL 
env$explanatory$var <- exp_var 

#cph model
formula_cph <- as.formula(paste0("Surv(fu_age_wk, dead_censor) ~", env$quart, "+", paste(env$explanatory$var, collapse = " + "))) 
cph <- coxph(formula_cph, data = surv_df) 
km_fit <- survfit(Surv(fu_age_wk, dead_censor) ~ eval(parse(text = quart)), data = surv_df_plot3) 

cph_summary <- summary(cph) 
value <- format(round(cph_summary$conf.int[,"exp(coef)"][3], 2), nsmall = 2)
le <- format(round(cph_summary$conf.int[,"lower .95"][3], 2), nsmall = 2)
ue <- format(round(cph_summary$conf.int[,"upper .95"][3], 2), nsmall = 2)
pval <- cph_summary$coefficients[,"Pr(>|z|)"][3]

#place asterisk for significant pval
if(pval < 0.050){
  hr <- paste0(value, " (", le, ", ", ue, ")*") 
}else{
  hr <- paste0(value, " (", le, ", ", ue, ")") 
}

km_curve <- ggsurvplot(
    fit = km_fit, 
    data = surv_df_plot3, 
    title = title,
    font.title = c(12), 
    conf.int = FALSE,
    risk.table = TRUE, 
    censor = FALSE,
    risk.table.col = "strata",
    legend.title = "Index Subgroup", 
    ggtheme = theme_bw(), 
    legend.labs = c("Highest Quartile", "Lowest Quartile"),
    palette = surv_color, 
    xlab = "Age (months)", 
    ylab = "Hazard Ratio", 
    xlim = c(start, end)
) 
surv_results <- list(km_curve, hr, pval, cph_summary) 

return(surv_results) 

} 

#Survival
km_start_50 <- slam_life_exp$life_exp_wk*.3 
km_start_75 <- slam_life_exp$life_exp_wk*.45 
km_start_100 <- slam_life_exp$life_exp_wk*.65 

#function calls 
slam_surv_gluc_50_all <- km_funct(quart = "gluc_50_quart", exp_var = c("age_50"), start = km_start_50, data = main_cat_surv, title = "Glucose at 50% Life Expectancy", end = 180) 
slam_surv_gluc_75_all <- km_funct(quart = "gluc_75_quart", exp_var = c("age_75"), start = km_start_75, data = main_cat_surv, title = "Glucose at 75% Life Expectancy", end = 180) 
slam_surv_gluc_100_all <- km_funct(quart = "gluc_100_quart", exp_var = c("age_100"), start = km_start_100, data = main_cat_surv, title = "Glucose at 100% Life Expectancy", end = 180) 

slam_surv_bw_50_all <- km_funct(quart = "bw_50_quart", exp_var = c("age_50"), start = km_start_50, data = main_cat_surv, title = "Body Weight at 50% Life Expectancy", end = 180) 
slam_surv_bw_75_all <- km_funct(quart = "bw_75_quart", exp_var = c("age_75"), start = km_start_75, data = main_cat_surv, title = "Body Weight at 75% Life Expectancy", end = 180) 
slam_surv_bw_100_all <- km_funct(quart = "bw_100_quart", exp_var = c("age_100"), start = km_start_100, data = main_cat_surv, title = "Body Weight at 100% Life Expectancy", end = 180) 

slam_surv_fat_50_all <- km_funct(quart = "fat_50_quart", exp_var = c("age_50"), start = km_start_50, data = main_cat_surv, title = "Fat at 50% Life Expectancy", end = 180) 
slam_surv_fat_75_all <- km_funct(quart = "fat_75_quart", exp_var = c("age_75"), start = km_start_75, data = main_cat_surv, title = "Fat at 75% Life Expectancy", end = 180) 
slam_surv_fat_100_all <- km_funct(quart = "fat_100_quart", exp_var = c("age_100"), start = km_start_100, data = main_cat_surv, title = "Fat at 100% Life Expectancy", end = 180) 

```

```{r slam_surv_output, warning=FALSE, message=FALSE, echo=FALSE}

#add adjustment subscript
slam_km_gluc_50 <- slam_surv_gluc_50_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(slam_surv_gluc_50_all[[2]]))), size = 5, parse =TRUE)
slam_km_gluc_75 <- slam_surv_gluc_75_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(slam_surv_gluc_75_all[[2]]))), size = 5, parse =TRUE)
slam_km_gluc_100 <- slam_surv_gluc_100_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(slam_surv_gluc_100_all[[2]]))), size = 5, parse =TRUE)

slam_km_bw_50 <- slam_surv_bw_50_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(slam_surv_bw_50_all[[2]]))), size = 5, parse =TRUE)
slam_km_bw_75 <- slam_surv_bw_75_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(slam_surv_bw_75_all[[2]]))), size = 5, parse =TRUE)
slam_km_bw_100 <- slam_surv_bw_100_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(slam_surv_bw_100_all[[2]]))), size = 5, parse =TRUE)

slam_km_fat_50 <- slam_surv_fat_50_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(slam_surv_fat_50_all[[2]]))), size = 5, parse =TRUE)
slam_km_fat_75 <- slam_surv_fat_75_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(slam_surv_fat_75_all[[2]]))), size = 5, parse =TRUE)
slam_km_fat_100 <- slam_surv_fat_100_all[[1]]$plot +
  annotate("text", x = -Inf, y = .02, hjust = -.09, label = deparse(bquote(HR[adj]== .(slam_surv_fat_100_all[[2]]))), size = 5, parse =TRUE)

km_all <- ggarrange(
  slam_km_gluc_50, slam_km_gluc_75, slam_km_gluc_100, slam_km_bw_50, slam_km_bw_75, slam_km_bw_100, slam_km_fat_50, slam_km_fat_75, slam_km_fat_100,
  common.legend = TRUE, legend = "right") 

```

```{r slam_surv_kable, fig.height = 12, fig.width = 12, warning=FALSE, message=FALSE, echo=FALSE}

km_all_output <- annotate_figure(km_all,
                                 top = text_grob("Survival Curve and Age Adjusted Mortality Hazard Ratio Comparing Highest and Lowest Metabolic Index Quartiles \n at Varying Timepoints\n", size = 14,  hjust = 0.52) 
                ) 

km_all_output

```

*** 

#### cox proportional mortality hazard ratios 
- sex specific
- covariate adjustment
  - sex, age 
  - additional adjustment for body fat 

```{r slam_cph_fun, warning=FALSE, message=FALSE, echo=FALSE}

cph_funct <- function(quart, exp_var, data, env = globalenv()){

surv_df <- as.data.frame(data) 

env$quart <- quart 
env$explanatory <- NULL 
env$explanatory$var <- exp_var 

formula_cph <- as.formula(paste0("Surv(fu_age_wk, dead_censor) ~", env$quart, "+", paste(env$explanatory$var, collapse = " + "))) 
cph <- coxph(formula_cph, data = surv_df) 

cph_summary <- summary(cph) 
value <- format(round(cph_summary$conf.int[,"exp(coef)"][3], 2), nsmall = 2) 
le <- format(round(cph_summary$conf.int[,"lower .95"][3], 2), nsmall = 2) 
ue <- format(round(cph_summary$conf.int[,"upper .95"][3], 2), nsmall = 2) 
pval <- cph_summary$coefficients[,"Pr(>|z|)"][3] 

if(pval < 0.050){
  hr <- paste0("HR = ", value, " (", le, ", ", ue, ")*") 
}else{
  hr <- paste0("HR = ", value, " (", le, ", ", ue, ")") 
}
 
surv_results <- list(hr, pval, cph_summary) 

return(surv_results) 

} 

#function calls 
slam_surv_gluc_50_all2 <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50",  "sex"), data = main_cat_surv)
slam_surv_gluc_75_all2 <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75", "sex"), data = main_cat_surv)
slam_surv_gluc_100_all2 <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100", "sex"), data = main_cat_surv)

slam_surv_bw_50_all2 <- cph_funct(quart = "bw_50_quart", exp_var = c("age_50", "sex"), data = main_cat_surv)
slam_surv_bw_75_all2 <- cph_funct(quart = "bw_75_quart", exp_var = c("age_75", "sex"), data = main_cat_surv)
slam_surv_bw_100_all2 <- cph_funct(quart = "bw_100_quart", exp_var = c("age_100", "sex"), data = main_cat_surv)

slam_surv_fat_50_all2 <- cph_funct(quart = "fat_50_quart", exp_var = c("age_50", "sex"), data = main_cat_surv)
slam_surv_fat_75_all2 <- cph_funct(quart = "fat_75_quart", exp_var = c("age_75", "sex"), data = main_cat_surv)
slam_surv_fat_100_all2 <- cph_funct(quart = "fat_100_quart", exp_var = c("age_100", "sex"), data = main_cat_surv)

## by sex 
slam_surv_gluc_50_f <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50"), data = main_cat_surv_f) 
slam_surv_gluc_75_f <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75"), data = main_cat_surv_f) 
slam_surv_gluc_100_f <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100"), data = main_cat_surv_f) 

slam_surv_bw_50_f <- cph_funct(quart = "bw_50_quart", exp_var = c("age_50"), data = main_cat_surv_f) 
slam_surv_bw_75_f <- cph_funct(quart = "bw_75_quart", exp_var = c("age_75"), data = main_cat_surv_f) 
slam_surv_bw_100_f <- cph_funct(quart = "bw_100_quart", exp_var = c("age_100"), data = main_cat_surv_f) 

slam_surv_fat_50_f <- cph_funct(quart = "fat_50_quart", exp_var = c("age_50"), data = main_cat_surv_f) 
slam_surv_fat_75_f <- cph_funct(quart = "fat_75_quart", exp_var = c("age_75"), data = main_cat_surv_f) 
slam_surv_fat_100_f <- cph_funct(quart = "fat_100_quart", exp_var = c("age_100"), data = main_cat_surv_f) 

slam_surv_gluc_50_m <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50"), data = main_cat_surv_m)
slam_surv_gluc_75_m <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75"), data = main_cat_surv_m) 
slam_surv_gluc_100_m <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100"), data = main_cat_surv_m) 

slam_surv_bw_50_m <- cph_funct(quart = "bw_50_quart", exp_var = c("age_50"), data = main_cat_surv_m)
slam_surv_bw_75_m <- cph_funct(quart = "bw_75_quart", exp_var = c("age_75"), data = main_cat_surv_m) 
slam_surv_bw_100_m <- cph_funct(quart = "bw_100_quart", exp_var = c("age_100"), data = main_cat_surv_m) 

slam_surv_fat_50_m <- cph_funct(quart = "fat_50_quart", exp_var = c("age_50"), data = main_cat_surv_m)
slam_surv_fat_75_m <- cph_funct(quart = "fat_75_quart", exp_var = c("age_75"), data = main_cat_surv_m) 
slam_surv_fat_100_m <- cph_funct(quart = "fat_100_quart", exp_var = c("age_100"), data = main_cat_surv_m) 

## adjusted for body fat
slam_surv_gluc_50_adj <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50", "sex", "fat_50_m"), data = main_cat_surv)
slam_surv_gluc_75_adj <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75", "sex", "fat_75_m"), data = main_cat_surv)
slam_surv_gluc_100_adj <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100", "sex", "fat_100_m"), data = main_cat_surv)

## by sex
slam_surv_gluc_50_adj_f <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50", "fat_50_m"), data = main_cat_surv_f)
slam_surv_gluc_75_adj_f <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75", "fat_75_m"), data = main_cat_surv_f)
slam_surv_gluc_100_adj_f <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100", "fat_100_m"), data = main_cat_surv_f)

slam_surv_gluc_50_adj_m <- cph_funct(quart = "gluc_50_quart", exp_var = c("age_50", "fat_50_m"), data = main_cat_surv_m)
slam_surv_gluc_75_adj_m <- cph_funct(quart = "gluc_75_quart", exp_var = c("age_75", "fat_75_m"), data = main_cat_surv_m)
slam_surv_gluc_100_adj_m <- cph_funct(quart = "gluc_100_quart", exp_var = c("age_100", "fat_100_m"), data = main_cat_surv_m)

```

```{r slam_km_by, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE, include=TRUE}

hr_toc_by <- NULL 
hr_toc_by$indices <- c("Glucose", "Glucose", "Glucose", "Body Weight", "Body Weight", "Body Weight", "Body Composition", "Body Composition", "Body Composition") 
hr_toc_by$timepoint <- c(50, 75, 100, 50, 75, 100, 50, 75, 100) 

##all
hr_toc_by$hr1_all <- c(slam_surv_gluc_50_all2[[1]], slam_surv_gluc_75_all2[[1]], slam_surv_gluc_100_all2[[1]], 
                       slam_surv_bw_50_all2[[1]], slam_surv_bw_75_all2[[1]], slam_surv_bw_100_all2[[1]], 
                       slam_surv_fat_50_all2[[1]], slam_surv_fat_75_all2[[1]], slam_surv_fat_100_all2[[1]]) 
hr_toc_by$hr2_all <- c(slam_surv_gluc_50_adj[[1]], slam_surv_gluc_75_adj[[1]], slam_surv_gluc_100_adj[[1]], 
                       NA, NA, NA, 
                       NA, NA, NA) 
##female
hr_toc_by$hr1_f <- c(slam_surv_gluc_50_f[[1]], slam_surv_gluc_75_f[[1]], slam_surv_gluc_100_f[[1]], 
                     slam_surv_bw_50_f[[1]], slam_surv_bw_75_f[[1]], slam_surv_bw_100_f[[1]], 
                     slam_surv_fat_50_f[[1]], slam_surv_fat_75_f[[1]], slam_surv_fat_100_f[[1]]) 
hr_toc_by$hr2_f <- c(slam_surv_gluc_50_adj_f[[1]], slam_surv_gluc_75_adj_f[[1]], slam_surv_gluc_100_adj_f[[1]], 
                     NA, NA, NA,
                     NA, NA, NA) 
##male
hr_toc_by$hr1_m <- c(slam_surv_gluc_50_m[[1]], slam_surv_gluc_75_m[[1]], slam_surv_gluc_100_m[[1]], 
                     slam_surv_bw_50_m[[1]], slam_surv_bw_75_m[[1]], slam_surv_bw_100_m[[1]], 
                     slam_surv_fat_50_m[[1]], slam_surv_fat_75_m[[1]], slam_surv_fat_100_m[[1]]) 
hr_toc_by$hr2_m <- c(slam_surv_gluc_50_adj_m[[1]], slam_surv_gluc_75_adj_m[[1]], slam_surv_gluc_100_adj_m[[1]], 
                     NA, NA, NA,
                     NA, NA, NA) 

hr_toc_by <- as.data.frame(hr_toc_by) 

##
hr_toc_by2 <- hr_toc_by %>% 
  mutate_all(~gsub("HR = ", "", .)) 

hr_toc_by2 <- hr_toc_by2 %>% 
  mutate( 
    hr1_all = noquote((gsub("\\(", "<br>(", hr1_all))), 
    hr2_all = noquote((gsub("\\(", "<br>(", hr2_all))), 
    hr1_m = noquote((gsub("\\(", "<br>(", hr1_m))), 
    hr2_m = noquote((gsub("\\(", "<br>(", hr2_m))), 
    hr1_f = noquote((gsub("\\(", "<br>(", hr1_f))), 
    hr2_f = noquote((gsub("\\(", "<br>(", hr2_f))),
    ) 
hr_toc_by3 <- hr_toc_by2 %>% 
  mutate(
    hr1_all = ifelse(grepl("Inf", hr1_all), "---", hr1_all),
    hr2_all = ifelse(grepl("Inf", hr2_all), "---", hr2_all), 
    hr1_m = ifelse(grepl("Inf", hr1_m), "---", hr1_m),
    hr2_m = ifelse(grepl("Inf", hr2_m), "---", hr2_m),
    hr1_f = ifelse(grepl("Inf", hr1_f), "---", hr1_f),
    hr2_f = ifelse(grepl("Inf", hr2_f), "---", hr2_f)
    )

slam_hr_toc_by_3a <- hr_toc_by3 %>% 
  filter(indices == "Glucose") %>% 
  select(-indices)

slam_hr_toc_by_3b <- hr_toc_by3 %>% 
  filter(indices != "Glucose") %>% 
  select(-contains("hr2")) 

```

```{r slam_hr_gluc_kable, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE, include=TRUE}

tablehr1 <- kable(slam_hr_toc_by_3a, row.names = FALSE, col.names = c("Timepoint" , "Model 1", "Model 2", "Model 1", "Model 2", "Model 1", "Model 2"), escape = FALSE, align = "c", caption = "Mortality Hazard Ratios of Highest and Lowest Glucose Quartiles \n at Varying Timepoints Stratified by Sex") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left", font_size = 12) %>% 
  column_spec(1, bold = TRUE, border_right = T) %>% 
  column_spec(3, border_right = T) %>% 
  column_spec(5, border_right = T) %>% 
  column_spec(c(2:7), width = "30em") %>% 
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  add_header_above(c(" " = 1, "All" = 2, "Female" = 2, "Male" = 2)) %>% 
  footnote(symbol = c("p-value < 0.05", "Hazard ratios (95% confidence intervals) are adjusted for: Model 1) age and sex; Model 2) additionally adjusted for body composition")
           ) 

tablehr1 

```

***
```{r slam_hr_bodycomp_kable, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE, include=TRUE}

tablehr2 <- kable(slam_hr_toc_by_3b, row.names = FALSE, col.names = c("Metabolic Index", "Timepoint$^$" , "All", "Female", "Male"), escape = FALSE, align = "c", caption = "Mortality Hazard Ratios of Highest and Lowest Body Weight and Composition Quartiles \n at Varying Timepoints Stratified by Sex") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left", font_size = 12) %>% 
  column_spec(1, bold = TRUE, border_right = T) %>% 
  column_spec(2, border_right = T) %>% 
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  footnote(symbol = c("p-value < 0.05", "Hazard ratios are adjusted for age and sex", "Timepoint is percent of mean life expectancy")
           ) 

tablehr2

```

## Author Questions 

#### determine assessments over lifespan 

```{r slam_imp_summary, warning=FALSE, message=FALSE, echo=FALSE}
#create when entry dataset
entry <- main_all3 %>% 
  group_by(idno) %>% 
  summarise( 
    age_entry = min(age_wk, na.rm=T), 
    
    whenentry =  as.factor( 
                ifelse((age_entry < (life_exp$life_exp_wk*0.125)), 0,
                ifelse((age_entry >= (life_exp$life_exp_wk*0.125)) & (age_entry < (life_exp$life_exp_wk*0.375)), 25,
                ifelse((age_entry >= (life_exp$life_exp_wk*0.375)) & (age_entry < (life_exp$life_exp_wk*0.625)), 50,
                ifelse((age_entry >= (life_exp$life_exp_wk*0.625)) & (age_entry < (life_exp$life_exp_wk*0.875)), 75,
                ifelse((age_entry >= (life_exp$life_exp_wk*0.875)) & (age_entry < (life_exp$life_exp_wk*1.125)), 100,
                ifelse((age_entry >= (life_exp$life_exp_wk*1.125)), 125, NA))))))))

asmt_funct_n <- function(time, gluc, gluc2, bw, bw2, fat, fat2){ 
  
  gluc_a <- paste0("!is.na(", gluc, ")") 
  gluc_b <- paste0("is.na(", gluc2, ")") 
  gluc2_a <- paste0("!is.na(", gluc2, ")") 
  
  bw_a <- paste0("!is.na(", bw, ")") 
  bw_b <- paste0("is.na(", bw2, ")") 
  bw2_a <- paste0("!is.na(", bw2, ")") 
  
  fat_a <- paste0("!is.na(", fat, ")") 
  fat_b <- paste0("is.na(", fat2, ")") 
  fat2_a <- paste0("!is.na(", fat2, ")") 
  

main_all3 %>% 
  left_join(entry) %>% 
  left_join(census_c16_death) %>% 
  mutate(whendead = as.numeric(as.character(whendead))) %>% 
  mutate(whenvisit = as.numeric(as.character(whenvisit))) %>% 
  mutate(whenentry = as.numeric(as.character(whenentry))) %>% 
  ungroup() %>% 
  summarise( 

      timepoint = time, 
      new_participants = n_distinct(idno[whenentry == time], na.rm = T), 

      alive = n_distinct(idno[whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      dead = n_distinct(idno[whendead == time], na.rm = T),
      
      gluc_seen = n_distinct(idno[eval(parse(text = gluc_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      gluc_imputed = n_distinct(idno[eval(parse(text = gluc2_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T) - gluc_seen, 
      gluc_missing = n_distinct(idno[eval(parse(text = gluc_b)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      
      bw_seen = n_distinct(idno[eval(parse(text = bw_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      bw_imputed = n_distinct(idno[eval(parse(text = bw2_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T) - bw_seen, 
      bw_missing = n_distinct(idno[eval(parse(text = bw_b)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      
      fat_seen = n_distinct(idno[eval(parse(text = fat_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      fat_imputed = n_distinct(idno[eval(parse(text = fat2_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T) - fat_seen, 
      fat_missing = n_distinct(idno[eval(parse(text = fat_b)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      ) 
} 

asmt_25 <- asmt_funct_n(25, "gluc_25_m", "cfgluc_25_m", "bw_25_m", "cfbw_25_m", "fat_25_m", "cffat_25_m")
asmt_50 <- asmt_funct_n(50, "gluc_50_m", "cfgluc_50_m", "bw_50_m", "cfbw_50_m", "fat_50_m", "cffat_50_m")
asmt_75 <- asmt_funct_n(75, "gluc_75_m", "cfgluc_75_m", "bw_75_m", "cfbw_75_m", "fat_75_m", "cffat_75_m")
asmt_100 <- asmt_funct_n(100, "gluc_100_m", "cfgluc_100_m", "bw_100_m", "cfbw_100_m", "fat_100_m", "cffat_100_m")
asmt_125 <- asmt_funct_n(125, "gluc_125_m", "cfgluc_125_m", "bw_125_m", "cfbw_125_m", "fat_125_m", "cffat_125_m")

imputation <- rbind(asmt_25, asmt_50, asmt_75, asmt_100, asmt_125)

```

```{r imputation_kable, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}

kable(imputation, row.names = FALSE, align = "c", format = "html", escape = F,
      caption = "Assessments of Metabolic indices Over the Lifespan", col.names = c("Timepoint$^$", "New Participants", "Alive", "Dead", "Measured", "Imputed", "Missing", "Measured", "Imputed", "Missing", "Measured", "Imputed", "Missing")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left") %>% 
  column_spec(1, bold = T, border_right = T) %>% 
  column_spec(4, border_right = T, width = c("250px"), include_thead =F) %>% 
  column_spec(7, border_right = T) %>% 
  column_spec(10, border_right = T) %>%
  collapse_rows(columns = 1, valign = "middle") %>% 
  add_header_above(c(" " = 1, "Population" = 3, "Glucose" = 3, "Body Weight" = 3, "Body Composition" = 3)) %>%  
  footnote(symbol = c("Imputation using last observed value carried forward", "Timepoint is percent of mean life expectancy"))

```

***

#### trajectories by varying survival groups

```{r slam_lifespan_dist_fig, cache=TRUE}

annotate_figure(slam_fig_d13, 
                top = text_grob("Trends in Metabolic indices Over the Lifespan By Survival Group"), 
                bottom = text_grob("Note: \nHorizontal lines indicate percent of life expectancy: green (25%), orange (50%), red (75%), purple (100%), and black (125%) \nDiamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines")) 
)

```
 
*** 
