---
title: "BLSA Glucose, Body Weight, and Body Comp"
author: "Eric Shiroma, John Nam"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    df_print: paged
---

## Study Description 

- This is an investigation into the longitudinal trajectories of glucose, body weight, and body composition in BLSA (a human cohort). These results will be used to compare similar trajectories in mice and monkeys

***

```{r packages, warning=FALSE, message=FALSE, echo=FALSE}

library (survival)
library (broom)
library (ggplot2)
library (ggthemes)
library (extrafont)
library (arsenal)
library (psych)
library (survey)
library (lubridate)
library (kableExtra)
library (gridExtra)
library (ggpubr)
library (grid)
library (tinytex)
library (haven)
library (tidyverse)
library (finalfit)
library (survival)
library (survminer)
library (BSDA)
library (readxl)
library (mgcv)

knitr::opts_chunk$set(autodep = TRUE) 
knitr::opts_chunk$set(warning = FALSE) 
knitr::opts_chunk$set(message = FALSE) 
knitr::opts_chunk$set(echo = FALSE) 
knitr::opts_chunk$set(cache = TRUE) 

```

## Analysis 
#### women blsa data read in
  - input file slam_blsa_w.sas7bdat (11/26/2019)

```{r women_blsa_data, warning=FALSE, message=FALSE, echo=FALSE}

w_blsa <- read_sas("data/blsa/slam_blsa_w.sas7bdat") %>% 
# w_blsa <- read_sas("/Users/shiromaej/Box/ISG Code Check/data/blsa/slam_blsa_w.sas7bdat") %>% 
  select(idno, visit, dov, Age, WtKg, Waist, male, RACECD, glucose000_adj, HtCm, diabetes) %>% 
  rename(age_wk = Age, bw = WtKg, fat = Waist, sex = male, race = RACECD, date = dov, gluc = glucose000_adj)
  
```

#### men blsa data read in
  - input file slam_blsa_m.sas7bdat (1/3/2020)
  
```{r men_blsa_data, warning=FALSE, message=FALSE, echo=FALSE}

m_blsa <- read_sas("data/blsa/slam_blsa_m.sas7bdat") %>% 
# m_blsa <- read_sas("/Users/shiromaej/Box/ISG Code Check/data/blsa/slam_blsa_m.sas7bdat") %>% 
  select(idno, visit, dov, Age, WtKg, Waist, male, RACECD, glucose000_adj, HtCm, diabetes) %>% 
  rename(age_wk = Age, bw = WtKg, fat = Waist, sex = male, race = RACECD, date = dov, gluc = glucose000_adj)

```

#### census data read in
  - input file blsa_death.sas7bdat (11/26/2019)

```{r census_blsa_data, warning=FALSE, message=FALSE, echo=FALSE}
 
census <- read_sas("data/blsa/blsa_death.sas7bdat") %>% 
# census <- read_sas("/Users/shiromaej/Box/ISG Code Check/data/blsa/blsa_death.sas7bdat") %>% 
  select(idno, visit, date = dov, dod)  
census$idno = as.character(census$idno) 
 
```

#### merging female and male datasets into a main file
  - w_blsa  (women blsa data) 
  - m_blsa (men blsa data) 
  - recode sex values 
    - 1 to "M"
    - 2 to "F"
  - recode race values
    - 1 to "white"
    - All else to "non-white

```{r blsa_filter_data, warning=FALSE, message=FALSE, echo=FALSE}

main_all <- rbind(w_blsa, m_blsa)

main_all$idno = as.character(main_all$idno) 
main_all$sex = as.character(main_all$sex) 
main_all$race = as.character(main_all$race) 

main_all <- main_all %>% 
  mutate(sex = as.factor(ifelse(sex %in% c(1, 2), "M", "F"))) %>% 
  mutate(race = as.factor(ifelse(is.na(race), "non-white", 
                  ifelse(race == 1, "white", "non-white" 
                         ))))
```

#### removal of duplicated rows of information
- keeping latest date for follow up age creation

```{r merge_blsa_data, warning=FALSE, message=FALSE, echo=FALSE}

main_all <- main_all %>% 
  arrange(desc(visit)) 

main_all2 <- main_all[!duplicated(main_all[c(1,4:10)]),] 

```

#### diabetic status creation 
- coded people with diabetes anytime during life as "diabetic" 

```{r blsa_diab_status_var, warning=FALSE, message=FALSE, echo=FALSE}

diab_status <- main_all %>% 
    group_by(idno) %>% 
    summarise( 
      diab_types = n_distinct(diabetes, na.rm =T),
      diab_count = sum(diabetes, na.rm=T), 
    )  %>% 
  #sets nonchanging diabetes status to 0
  mutate(diab_types2 = ifelse(diab_types == 1, 0, 1)) %>% 
  #sets no diabetes to 0
  mutate(diab_count2 = ifelse(diab_count == 0, 0, 1)) %>% 
  mutate(diabetes = ifelse(diab_count2 + diab_types2 == 0, 0, 1)) %>% 
  mutate(diabetes2 = ifelse(diabetes == 1, "diabetic", 
                     ifelse(diabetes == 0, "non-diabetic", NA))) %>%
  select(idno, diabetes2) 

main_all2 <- main_all %>% 
  left_join(diab_status, by = "idno") 

```

```{r blsa_ratios, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}

main_all2_n <- n_distinct(main_all2$idno)

## sexrace tables 
subgroup <- main_all2 %>% 
  select(idno, sex, race, diabetes2) %>% 
  distinct() 

sextable <- table(subgroup$sex)
racetable <- table(subgroup$race)
diabetestable <- table(subgroup$diabetes2)

blsa_sex_ratio = format(round(((sextable[1] / (sextable[1] + sextable[2]))*100), 2), nsmall = 2)
blsa_race_ratio = format(round(((racetable[1] / (racetable[1] + racetable[2]))*100), 2), nsmall = 2)
blsa_diabetes_ratio = format(round(((diabetestable[1] / (diabetestable[1] + diabetestable[2]))*100), 2), nsmall = 2)

```

#### population characteristics 
- `r main_all2_n` participants with valid indices 
- `r blsa_sex_ratio` % female 
- `r blsa_race_ratio` % non-white 
- `r blsa_diabetes_ratio` % diabetic 
 
#### age characteristics

```{r blsa_age_desc, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

main_all2_desc <- main_all2 %>%
group_by(idno, sex, race) %>%
summarise(
  age_wk = min(age_wk, na.rm=T)
) 

main_all2_desc2_all <- format(round(psych::describe(main_all2_desc$age_wk), 2), nsmall=2) %>% 
  mutate_each(funs(as.numeric(.))) %>%
  mutate(group1 = "all") %>% 
  select(-c("vars")) 
main_all2_desc_sex <- as.data.frame(describeBy(main_all2_desc$age_wk, main_all2_desc$sex, mat=TRUE, digits = 2)) %>% 
  select(-c("item", "vars")) %>% 
  mutate(n = as.numeric(n))
main_all2_desc_race <- as.data.frame(describeBy(main_all2_desc$age_wk, main_all2_desc$race, mat=TRUE, digits = 2)) %>% 
  select(-c("item", "vars")) %>% 
  mutate(n = as.numeric(n))

main_all2_desc2 <- bind_rows(main_all2_desc2_all, main_all2_desc_sex, main_all2_desc_race) %>% 
  select(group = group1, everything())

```

```{r blsa_age_desc_kable, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

kable(main_all2_desc2, caption = "Age Characteristics",row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

#### census_c16_death creation
  - merge 
    - census
    - main_all2
  - create all cause death censor category
  
***

```{r blsa_censor_df, warning=FALSE, message=FALSE, echo=FALSE}

census_c16_death <- main_all2 %>% 
  select(idno, date, age_wk, sex, race) %>% 
  distinct() %>% 
  group_by(idno, sex, race) %>% 
  summarise( 
    lastvisit = max(date), 
    age_first = min(age_wk), 
    age_last = max(age_wk), 
    ) %>% 
  left_join(census, by = c("idno", "lastvisit" = "date")) %>% 
  mutate(dead_censor = ifelse(is.na(dod), 0, 1)) %>% 
  #time between last visit and death
  mutate(time2dod = as.numeric(floor(((dod - lastvisit) + 1) / 365))) %>% 
  #age at death
  mutate(le_wk = ifelse(!is.na(time2dod), age_last + time2dod, age_last)) %>% 
  mutate(fu_age_wk = ifelse(dead_censor == 1, le_wk - age_first, age_last - age_first)) %>% 
  left_join(diab_status, by = "idno") %>% 
  select(idno, dod, fu_age_wk, le_wk, dead_censor, sex, race, diabetes2) %>% 
  distinct() 

dctable <- as.data.frame(table(census_c16_death$dead_censor)) 
fu_time_desc <- format(round(as.data.frame(psych::describe(as.numeric(census_c16_death$fu_age_wk))), 2), nsmall=2) 

``` 

#### death frequency

```{r death_cat, warning=FALSE, message=FALSE, echo=FALSE} 
kable(dctable, caption = "Death Frequency", col.names = c("Dead = 1", "Frequency"), row.names = FALSE, align = "c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

***

#### follow up time

```{r followup_cat, warning=FALSE, message=FALSE, echo=FALSE} 
kable(fu_time_desc, caption = "Follow Up Time", row.names = FALSE, align = "c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE) 
```

*** 

#### life expectancy

```{r blsa_life_expectancy, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}

create_LE <- function(data, criteria, filter){

life_exp <- data %>% 
  ungroup() %>% 
  filter(!!filter) %>% 
  summarise( 
            total = n_distinct(idno), 
            dead = n_distinct(idno[!!criteria]), 
            life_exp_wk = round(mean(le_wk[!!criteria]), 2), 
            sd = round(sd(as.numeric(le_wk[!!criteria])), 2), 
            se = round(sd(as.numeric(le_wk[!!criteria])) / sqrt(dead),2),
            lcl = round(life_exp_wk - 1.96 * se, 2), 
            ucl = round(life_exp_wk + 1.96 * se, 2) 
  ) 
life_exp$life_exp_wk <- as.numeric(life_exp$life_exp_wk) 

return(life_exp)

}

life_exp <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(sex %in% c("M", "F"))) %>% 
  mutate(group = "all") 
life_exp_f <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(sex %in% c("F"))) %>% 
  mutate(group = "female") 
life_exp_m <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(sex %in% c("M"))) %>% 
  mutate(group = "male") 
life_exp_nw <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(race %in% c("non-white"))) %>% 
  mutate(group = "non-white") 
life_exp_w <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(race %in% c("white"))) %>% 
  mutate(group = "white") 
life_exp_diab <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(diabetes2 %in% c("diabetic"))) %>% 
  mutate(group = "diabetic") 
life_exp_nondiab <- create_LE(data = census_c16_death, criteria = quote(dead_censor %in% c(1)), filter = quote(diabetes2 %in% c("non-diabetic"))) %>% 
  mutate(group = "non-diabetic") 

life_exp2 <- bind_rows(life_exp, life_exp_f, life_exp_m, life_exp_nw, life_exp_w) %>% 
  select(group, everything())
  
## for isg template
blsa_life_exp <- life_exp 
blsa_life_exp_f <- life_exp_f
blsa_life_exp_m <- life_exp_m 
blsa_life_exp_nw <- life_exp_nw 
blsa_life_exp_w <- life_exp_w 
blsa_life_exp_diab <- life_exp_diab 
blsa_life_exp_nondiab <- life_exp_nondiab

```

```{r blsa_life_exp_kable, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}
kable(life_exp2[,c(1:5)], caption = "Life Expectancy", col.names = c("Group", "Total", "Dead", "Life Expectancy <br> (weeks)", "sd"), row.names = TRUE, align = "c", escape = FALSE) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left") 
```

***

#### create timepoints as percent life expectancy
- "per_age_wk" in main_all2
- "per_fu_wk" and "per_le_wk" in census_c16_death

```{r percent_le, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}

main_all2 <- main_all2 %>% 
  mutate(per_age_wk = 
           ifelse(sex == "M", (age_wk / life_exp_m$life_exp_wk)*100, 
           ifelse(sex == "F", (age_wk / life_exp_f$life_exp_wk)*100, age_wk / life_exp$life_exp_wk))
  )

#renamed for isg template
blsa_main_all <- main_all2 

census_c16_death <- census_c16_death %>% 
  mutate(per_fu_wk = 
           ifelse(sex == "M", (fu_age_wk / life_exp_m$life_exp_wk)*100, 
           ifelse(sex == "F", (fu_age_wk / life_exp_f$life_exp_wk)*100, fu_age_wk / life_exp$life_exp_wk)), 
         per_le_wk = 
           ifelse(sex == "M", (le_wk / life_exp_m$life_exp_wk)*100, 
           ifelse(sex == "F", (le_wk / life_exp_f$life_exp_wk)*100, le_wk / life_exp$life_exp_wk)), 
  ) 

```

#### frequencies of measures

```{r blsa_follow_up, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}

n_asmt <- main_all2 %>% 
  group_by(idno, sex) %>% 
  summarise_all(funs( 
    n = sum(!is.na(.)), 
    sd = format(round(sd(!is.na(.)), 2), nsmall=2) 
  )) 

gluc_asmt <- format(round(as.data.frame(psych::describe(n_asmt$gluc_n)), 2), nsmall=2) %>% 
  mutate(group1 = "all") %>% 
  select(group1, everything(), -vars)%>% 
  mutate(n = as.numeric(n))
bw_asmt <- format(round(as.data.frame(psych::describe(n_asmt$bw_n)), 2), nsmall=2) %>% 
  mutate(group1 = "all") %>% 
  select(group1, everything(), -vars)%>% 
  mutate(n = as.numeric(n))
fat_asmt <- format(round(as.data.frame(psych::describe(n_asmt$fat_n)), 2), nsmall=2) %>% 
  mutate(group1 = "all") %>% 
  select(group1, everything(), -vars)%>% 
  mutate(n = as.numeric(n))

##by sex
gluc_asmt_sex <- as.data.frame(describeBy(n_asmt$gluc_n, n_asmt$sex, mat=TRUE, digits = 2)) %>% 
  select(-c("item", "vars")) %>% 
  mutate(n = as.numeric(n))
bw_asmt_sex <- as.data.frame(describeBy(n_asmt$bw_n, n_asmt$sex, mat=TRUE, digits = 2)) %>% 
  select(-c("item", "vars")) %>% 
  mutate(n = as.numeric(n))
fat_asmt_sex <- as.data.frame(describeBy(n_asmt$fat_n, n_asmt$sex, mat=TRUE, digits = 2)) %>% 
  select(-c("item", "vars")) %>% 
  mutate(n = as.numeric(n))

## age desc by life expectancy
main_all2_desc_gluc <- main_all2 %>%
filter(!is.na(gluc)) %>% 
group_by(idno, sex) 
main_all2_desc_bw <- main_all2 %>%
filter(!is.na(bw)) %>% 
group_by(idno, sex) 
main_all2_desc_fat <- main_all2 %>%
filter(!is.na(fat)) %>% 
group_by(idno, sex) 

#by age weeks
allgluc_age_range <- format(round(as.data.frame(psych::describe(main_all2_desc_gluc$age_wk)), 2), nsmall=2) %>% 
  mutate(group1 = "all") %>% 
  select(group1, age_min = min, age_max = max)
allbw_age_range <- format(round(as.data.frame(psych::describe(main_all2_desc_bw$age_wk)), 2), nsmall=2) %>% 
  mutate(group1 = "all") %>% 
  select(group1, age_min = min, age_max = max)
allfat_age_range <- format(round(as.data.frame(psych::describe(main_all2_desc_fat$age_wk)), 2), nsmall=2) %>% 
  mutate(group1 = "all") %>% 
  select(group1, age_min = min, age_max = max)
## by sex
gluc_age_range <- as.data.frame(describeBy(main_all2_desc_gluc$age_wk, main_all2_desc_gluc$sex, mat=TRUE, digits = 2)) %>% 
  select(group1, age_min = min, age_max = max) 
bw_age_range <- as.data.frame(describeBy(main_all2_desc_bw$age_wk, main_all2_desc_bw$sex, mat=TRUE, digits = 2)) %>% 
  select(group1, age_min = min, age_max = max) 
fat_age_range <- as.data.frame(describeBy(main_all2_desc_fat$age_wk, main_all2_desc_fat$sex, mat=TRUE, digits = 2)) %>% 
  select(group1, age_min = min, age_max = max) 

## by percent age 
allgluc_perage_range <- format(round(as.data.frame(psych::describe(main_all2_desc_gluc$per_age_wk)), 2), nsmall=2) %>% 
  mutate(group1 = "all") %>% 
  select(group1, le_min = min, le_max = max)
allbw_perage_range <- format(round(as.data.frame(psych::describe(main_all2_desc_bw$per_age_wk)), 2), nsmall=2) %>% 
  mutate(group1 = "all") %>% 
  select(group1, le_min = min, le_max = max)
allfat_perage_range <- format(round(as.data.frame(psych::describe(main_all2_desc_fat$per_age_wk)), 2), nsmall=2) %>% 
  mutate(group1 = "all") %>% 
  select(group1, le_min = min, le_max = max)
#by sex
gluc_perage_range <- as.data.frame(describeBy(main_all2_desc_gluc$per_age_wk, main_all2_desc_gluc$sex, mat=TRUE, digits = 2)) %>% 
  select(group1, le_min = min, le_max = max) 
bw_perage_range <- as.data.frame(describeBy(main_all2_desc_bw$per_age_wk, main_all2_desc_bw$sex, mat=TRUE, digits = 2)) %>% 
  select(group1, le_min = min, le_max = max) 
fat_perage_range <- as.data.frame(describeBy(main_all2_desc_fat$per_age_wk, main_all2_desc_fat$sex, mat=TRUE, digits = 2)) %>% 
  select(group1, le_min = min, le_max = max) 

##

gluc_asmt2 <- gluc_asmt %>% 
  left_join(allgluc_age_range) %>% 
  left_join(allgluc_perage_range) 
gluc_asmt_sex2 <- gluc_asmt_sex %>% 
  left_join(gluc_age_range) %>% 
  left_join(gluc_perage_range) 
bw_asmt2 <- bw_asmt %>% 
  left_join(allbw_age_range) %>% 
  left_join(allbw_perage_range) 
bw_asmt_sex2 <- bw_asmt_sex %>% 
  left_join(bw_age_range) %>% 
  left_join(bw_perage_range) 
fat_asmt2 <- fat_asmt %>% 
  left_join(allfat_age_range) %>% 
  left_join(allfat_perage_range) 
fat_asmt_sex2 <- fat_asmt_sex %>% 
  left_join(fat_age_range) %>% 
  left_join(fat_perage_range) 

gluc_amst3 <- rbind(gluc_asmt2, gluc_asmt_sex2)
bw_amst3 <- rbind(bw_asmt2, bw_asmt_sex2)
fat_amst3 <- rbind(fat_asmt2, fat_asmt_sex2)

all_asmt <- rbind(gluc_amst3, bw_amst3, fat_amst3) 
all_asmt$indices <- c("glucose", "glucose", "glucose", "body weight", "body weight", "body weight", "body fat", "body fat", "body fat")
all_asmt <- all_asmt %>% 
  select(indices, group = group1, everything())

## for template
blsa_gluc_asmt <- gluc_asmt
blsa_bw_asmt <- bw_asmt
blsa_fat_asmt <- fat_asmt

```

```{r asmt_kable, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}

kable(all_asmt, caption = "Assessment Frequencies", row.names = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE) %>% 
  collapse_rows(columns = 1, valign = "middle") %>% 
  column_spec(c(1), border_right = TRUE) 

```

*** 

#### age at peak indices value creation 

```{r blsa_life_peaks, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}

#functions to create peaks for different groups
lifespan_peaks <- function(metric, data, filter){

#set boundary conditions for peak to lowest and highest life expectancy timepoints
low_bound <- 25
high_bound <- 125

#subset arguments
sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)") 
metric2 <- paste0("!is.na(", metric, ")") 

#subset for non-missing data
df <- base::subset(data, eval(parse(text = metric2))) 

#subset data by xlimit and subgroup restraints
peaks_df <- subset(df, eval(parse(text = sub))) %>% 
  arrange(per_age_wk) %>% 
  filter(!!filter) 

#gam model fit using geom_smooth parameters
gam_fit <- mgcv::gam(eval(parse(text = metric)) ~ s(per_age_wk, bs = "cs"), method = "REML", data = peaks_df) 

#create fitted values using gam model from dataframe age values
predict1 <- as.data.frame(predict.gam(gam_fit, peaks_df, se.fit=TRUE)) 

#create n and sd by age group
fit2 <- peaks_df %>% 
  bind_cols(predict1) %>% 
  select(per_age_wk, fit, se.fit) %>% 
  filter(per_age_wk >= low_bound, per_age_wk <= high_bound) %>%
  group_by(per_age_wk) %>% 
  mutate( 
    n = n(), 
    sd = se.fit*sqrt(n) 
  ) %>% 
  ungroup() 

#create low bound peak object
low_bound_peak <- fit2 %>% 
  filter(per_age_wk == min(per_age_wk)) %>% 
  select(per_age_wk, fit, se.fit, n, sd) %>% 
  distinct()

#create high bound peak object
high_bound_peak <- fit2 %>% 
  filter(per_age_wk == max(per_age_wk)) %>% 
  select(per_age_wk, fit, se.fit, n, sd) %>% 
  distinct()

#pulling peak age and value
infl <- NULL
infl$logic <- c(NA, diff(diff(fit2$fit)) < 0, NA)
infl <- as.data.frame(infl)

#finding age of the first point significantly different from peak for the lower range
#setting to lower bound if no values are present
peak <- fit2 %>% 
  bind_cols(infl) %>% 
  filter(infl$logic == "TRUE") %>% 
  group_by(per_age_wk, n, sd, se.fit) %>%
  summarise(
    fit = max(fit)
  )  %>% 
  ungroup() %>% 
  filter(fit == max(fit, na.rm=T))

#finding age of the first point significantly different from peak for the lower range
#setting to lower bound if no values are present
fit_diff_min <- tryCatch(
fit2 %>% 
  filter(per_age_wk < peak$per_age_wk) %>% 
  mutate(fit_diff = fit - peak$fit) %>% 
  mutate(age_diff = peak$per_age_wk - per_age_wk) %>% 
  mutate(pval = tsum.test(fit_diff, sd, n, mu = 0, conf.level = 0.95)$p.value) %>% 
  filter(pval <= 0.05) %>% 
  filter( 
    age_diff == min(age_diff) 
    ) %>% 
  distinct() %>% 
  select(per_age_wk, fit), 
error=function(e) low_bound_peak)

#finding age of the first point significantly different from peak for the higher range
#setting to higher bound if no values are present
fit_diff_max <- tryCatch(
fit2 %>% 
  filter(per_age_wk > peak$per_age_wk) %>% 
  mutate(fit_diff = fit - peak$fit) %>% 
  mutate(age_diff = per_age_wk - peak$per_age_wk) %>% 
  mutate(pval = tsum.test(fit_diff, sd, n, mu = 0, conf.level = 0.95)$p.value) %>% 
  filter(pval <= 0.05) %>% 
  filter( 
    age_diff == min(age_diff) 
    ) %>% 
  distinct() %>% 
  select(per_age_wk, fit),
error=function(e) high_bound_peak)

#setting to lower bound if no values are present
if(dim(fit_diff_min)[1] == 0){
  fit_diff_min <- low_bound_peak
}else{
}

#setting to higher bound if no values are present
if(dim(fit_diff_max)[1] == 0){
  fit_diff_max <- high_bound_peak
}else{
}

#putting all outputs into list 
all_peaks <- list(peak, fit_diff_min, fit_diff_max, high_bound_peak, low_bound_peak) 
names(all_peaks) <- c("all", "min", "max", "highbound", "lowbound")

return(all_peaks) 

} 

gluc_peak <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(sex %in% c("M", "F"))) 
bw_peak <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(sex %in% c("M", "F"))) 
fat_peak <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(sex %in% c("M", "F"))) 

gluc_peak_male <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(sex %in% c("M"))) 
bw_peak_male <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(sex %in% c("M"))) 
fat_peak_male <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(sex %in% c("M"))) 

gluc_peak_female <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(sex %in% c("F"))) 
bw_peak_female <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(sex %in% c("F"))) 
fat_peak_female <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(sex %in% c("F"))) 

gluc_peak_w <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(race %in% c("white"))) 
bw_peak_w <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(race %in% c("white"))) 
fat_peak_w <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(race %in% c("white"))) 

gluc_peak_nw <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(race %in% c("non-white"))) 
bw_peak_nw <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(race %in% c("non-white"))) 
fat_peak_nw <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(race %in% c("non-white"))) 

gluc_peak_diab <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(diabetes2 %in% c("diabetic")))
bw_peak_diab <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(diabetes2 %in% c("diabetic")))
fat_peak_diab <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(diabetes2 %in% c("diabetic")))

gluc_peak_nondiab <- lifespan_peaks(metric = "gluc", data = main_all2, filter = quote(diabetes2 %in% c("non-diabetic"))) 
bw_peak_nondiab <- lifespan_peaks(metric = "bw", data = main_all2, filter = quote(diabetes2 %in% c("non-diabetic"))) 
fat_peak_nondiab <- lifespan_peaks(metric = "fat", data = main_all2, filter = quote(diabetes2 %in% c("non-diabetic"))) 

```

#### main_cat2 creation 
- merge 
  - main_all2 
  - census_c16_death 

#### imputation 
- carry forward previous observed non-missing value 
  - filter out all missing values 
  - filter for rows less than than 50 + 1, 75 + 1, and 100 + 1 (previous observed values)
  - identify last non-missing value for imputation if needed 
  - leave missing if no last non-missing value 
  - merge into main_cat2 dataset
  
```{r blsa_main_cat, warning=FALSE, message=FALSE, echo=FALSE}

main_cat2 <- main_all2 %>% 
  left_join(census_c16_death) %>% 
  select(idno) %>% 
  distinct() 

carry_forward <- function(index, cat){
  
index_cat <- paste0(index, "_", cat)

 output <- main_all2 %>% 
   left_join(census_c16_death) %>% 
   group_by(idno) %>% 
   filter(!is.na(!!index)) %>% 
   #get all previous values to carry forward if needed
   filter(per_age_wk < (!!cat + 1)) %>%
   mutate(!!index_cat := !!index) %>% 
   #choose closst value
   filter(per_age_wk == max(per_age_wk)) %>% 
   #filter for those alive at timepoint 
   filter(per_le_wk >= !!cat) %>% 
   #mean of data at same timepoint 
   summarise(
     !!index_cat := mean(!!index, na.rm=T)
   ) %>% 
   select(idno, !!index_cat) %>% 
   distinct() %>% 
   right_join(main_cat2, by = "idno")
 
 return(output)
 
} 

main_cat2 <- carry_forward(quote(gluc), quote(25))
main_cat2 <- carry_forward(quote(gluc), quote(50))
main_cat2 <- carry_forward(quote(gluc), quote(75))
main_cat2 <- carry_forward(quote(gluc), quote(100))
main_cat2 <- carry_forward(quote(gluc), quote(125))

main_cat2 <- carry_forward(quote(bw), quote(25))
main_cat2 <- carry_forward(quote(bw), quote(50))
main_cat2 <- carry_forward(quote(bw), quote(75))
main_cat2 <- carry_forward(quote(bw), quote(100))
main_cat2 <- carry_forward(quote(bw), quote(125))

main_cat2 <- carry_forward(quote(fat), quote(25))
main_cat2 <- carry_forward(quote(fat), quote(50))
main_cat2 <- carry_forward(quote(fat), quote(75))
main_cat2 <- carry_forward(quote(fat), quote(100))
main_cat2 <- carry_forward(quote(fat), quote(125))

#age adjustment 
main_cat2 <- carry_forward(quote(per_age_wk), quote(25))
main_cat2 <- carry_forward(quote(per_age_wk), quote(50))
main_cat2 <- carry_forward(quote(per_age_wk), quote(75))
main_cat2 <- carry_forward(quote(per_age_wk), quote(100))
main_cat2 <- carry_forward(quote(per_age_wk), quote(125))

```

#### main_all3 creation
- merge 
  - main_all2 
  - census_c16_death 
  - main_cat2
 
```{r blsa_main_all3, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

main_all3 <- main_all2 %>% 
  left_join(census_c16_death) %>% 
  left_join(main_cat2) %>% 

  mutate( 

    whendead = as.factor( 
                ifelse(per_le_wk >= 0 & per_le_wk < 37.5, 25,
                ifelse(per_le_wk >= 37.5 & per_le_wk < 62.5, 50,
                ifelse(per_le_wk >= 62.5 & per_le_wk < 87.5, 75,
                ifelse(per_le_wk >= 87.5 & per_le_wk < 112.5, 100,
                ifelse(per_le_wk >= 112.5 & per_le_wk < 150.0, 125, NA)))))), 
    
    whenvisit = as.factor( 
                ifelse(per_age_wk >= 0 & per_age_wk < 37.5, 25,
                ifelse(per_age_wk >= 37.5 & per_age_wk < 62.5, 50,
                ifelse(per_age_wk >= 62.5 & per_age_wk < 87.5, 75,
                ifelse(per_age_wk >= 87.5 & per_age_wk < 112.5, 100,
                ifelse(per_age_wk >= 112.5 & per_age_wk < 150.0, 125, NA))))))
  ) 

```
 
## Results 
#### distribution figure with trend line
- generalized additive model for main indices
- linear model with quadratic fit for lifespan groups (survival bias analyses)
 
```{r blsa_ss_dist_funct, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

surv_color <- c("green", "yellow", "red", "purple", "gray")

fcn_plotdist_smooth <- function(metric, x, y, by, xlabel, ylabel, title, xlow, xhigh, peak1, peak2) { 

  #subsetting
  metric2 <- paste0("!is.na(", metric, ")") 
  metric3 <- paste0("df$", metric) 
  main_all3 <- main_all3 %>% 
    filter(per_age_wk < as.numeric(all_asmt$le_max[1]))
  
  df <- base::subset(main_all3, eval(parse(text = metric2))) 
  sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)") 
    
  z <- Inf

  ggplot(subset(df, eval(parse(text = sub))), aes_string(x = x, y = y, colour = by)) + 
    
    #dots
    geom_point(alpha=0.5, position = "jitter", size = 1, aes_string(colour = by), shape=1) +
    
    #trendlines 
    geom_smooth(aes_string(group = by, linetype = "solid"), linetype = "solid", color = "black", size = 1.8) +
    geom_smooth(aes(linetype = "solid")) + 
    
    #peaks
    geom_point(aes(x = peak1$all$per_age_wk, y = peak1$all$fit), size=2, color="black", fill = "white", alpha=0.7, shape=23, stroke=2) +
    geom_point(aes(x = peak2$all$per_age_wk, y = peak2$all$fit), size=2, color="black", fill = "white", alpha=0.7, shape=23, stroke=2) +
    
    #labels
    xlab(xlabel) + 
    ylab(ylabel) + 
    xlim(xlow, xhigh) +
    
    scale_linetype_manual(labels = c("All"), name = "", values = c("solid", "dashed", "solid"), guide = FALSE)
} 

fcn_plotdist_lm <- function(metric, x, y, by, xlabel, ylabel, title, xlow, xhigh) { 

  #subsetting
  metric2 <- paste0("!is.na(", metric, ")") 
  metric3 <- paste0("df$", metric) 
  main_all3 <- main_all3 %>% filter(as.numeric(whenvisit) <= as.numeric(whendead)) %>% filter(as.numeric(whendead) > 0) 
  df <- base::subset(main_all3, eval(parse(text = metric2))) 
  sub <- paste0(metric, " > quantile(", metric, ", 0.025) & ", metric, " < quantile(", metric, ", 0.975)") 
  z <- Inf

  ggplot(subset(df, eval(parse(text = sub))), aes_string(x = x, y = y, colour = by)) + 
    #trendlines 
    geom_smooth(aes_string(group = by, linetype = "solid"), method = "lm", formula = y ~ poly(x, 3), 
                linetype = "solid", color = "black", size = 1.8, se = FALSE) +
    geom_smooth(aes(linetype = "solid"), method = "lm", formula = y ~ poly(x, 3), size = 1, se = FALSE) + 
    
    #labels
    xlab(xlabel) + 
    ylab(ylabel) +
    xlim(xlow, xhigh) + 
    
    scale_linetype_manual(labels = c("All"), name = "", values = c("solid", "dashed", "solid"), guide = FALSE) +
    scale_color_manual(values = surv_color, name = "Survival\n(% Life Expectancy)")
} 

``` 
 
```{r blsa_ss_dist_plots, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

cbPalette_dot <- c("#f8766d", "#00b0f6", "black") 

#sex
blsa_fig_i1 <- fcn_plotdist_smooth(metric = "gluc", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 170, 
                        y = "gluc", ylabel = "Glucose (mg/dL)\n", by = "sex", peak1 = gluc_peak_male, peak2 = gluc_peak_female) + 
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)
blsa_fig_i2 <- fcn_plotdist_smooth(metric = "bw", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 150, 
                        y = "bw", ylabel = "Body Weight (kg)\n", by = "sex", peak1 = bw_peak_male, peak2 = bw_peak_female) + 
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)
blsa_fig_i3 <- fcn_plotdist_smooth(metric = "fat", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 150, 
                        y = "fat", ylabel = "Body Fat (kg)\n", by = "sex", peak1 = fat_peak_male, peak2 = fat_peak_female) + 
  scale_color_manual(labels = c("Female", "Male"), name = "Sex", values = cbPalette_dot)
#race
blsa_fig_j1 <- fcn_plotdist_smooth(metric = "gluc", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 170, 
                        y = "gluc", ylabel = "Glucose (mg/dL)", by = "race", peak1 = gluc_peak_w, peak2 = gluc_peak_nw) +
  scale_color_manual(labels = c("Non-White", "White"), name = "Colony", values = cbPalette_dot)
blsa_fig_j2 <- fcn_plotdist_smooth(metric = "bw", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 150, 
                        y = "bw", ylabel = "Body Weight (kg)", by = "race", peak1 = bw_peak_w, peak2 = bw_peak_nw) + 
  scale_color_manual(labels = c("Non-White", "White"), name = "Colony", values = cbPalette_dot)
blsa_fig_j3 <- fcn_plotdist_smooth(metric = "fat", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 150, 
                        y = "fat", ylabel = "Body Fat (kg)", by = "race", peak1 = fat_peak_w, peak2 = fat_peak_nw) + 
  scale_color_manual(labels = c("Non-White", "White"), name = "Colony", values = cbPalette_dot)
#diabetes
blsa_fig_diab1 <- fcn_plotdist_smooth(metric = "gluc", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 170, 
                        y = "gluc", ylabel = "Glucose (mg/dL)", by = "diabetes2", peak1 = gluc_peak_diab, peak2 = gluc_peak_nondiab) +
  scale_color_manual(labels = c("Diabetic", "Non-Diabetic"), name = "Health Status", values = cbPalette_dot)
blsa_fig_diab2 <- fcn_plotdist_smooth(metric = "bw", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 150, 
                        y = "bw", ylabel = "Body Weight (kg)", by = "diabetes2", peak1 = bw_peak_diab, peak2 = bw_peak_nondiab) + 
  scale_color_manual(labels = c("Diabetic", "Non-Diabetic"), name = "Health Status", values = cbPalette_dot)
blsa_fig_diab3 <- fcn_plotdist_smooth(metric = "fat", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 150, 
                        y = "fat", ylabel = "Body Fat (kg)", by = "diabetes2", peak1 = fat_peak_diab, peak2 = fat_peak_nondiab) + 
  scale_color_manual(labels = c("Diabetic", "Non-Diabetic"), name = "Health Status", values = cbPalette_dot)
#death
blsa_fig_d1 <- fcn_plotdist_lm(metric = "gluc", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 170, 
                        y = "gluc", ylabel = "Glucose (mg/dL)", by = "whendead")
blsa_fig_d2 <- fcn_plotdist_lm(metric = "bw", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 150, 
                        y = "bw", ylabel = "Body Weight (kg)", by = "whendead") 
blsa_fig_d3 <- fcn_plotdist_lm(metric = "fat", x = "per_age_wk", xlabel = "Percent Life Expectancy", xlow = 0, xhigh = 150, 
                        y = "fat", ylabel = "Body Fat (kg)", by = "whendead") 

## 
blsa_fig_i13 <- ggarrange(blsa_fig_i1, blsa_fig_i2, blsa_fig_i3,
                      ncol = 2, nrow = 2,
                     common.legend = TRUE,
                     legend = "right")

blsa_fig_j13 <- ggarrange(blsa_fig_j1, blsa_fig_j2, blsa_fig_j3,
                      ncol = 2, nrow = 2,
                     common.legend = TRUE,
                     legend = "right")

blsa_fig_diab13 <- ggarrange(blsa_fig_diab1, blsa_fig_diab2, blsa_fig_diab3,
                      ncol = 2, nrow = 2,
                     common.legend = TRUE,
                     legend = "right")

blsa_fig_d13 <- ggarrange(blsa_fig_d1, blsa_fig_d2, blsa_fig_d3,
                      ncol = 2, nrow = 2,
                     common.legend = TRUE,
                     legend = "right")

``` 

```{r blsa_sex_dist_fig, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

annotate_figure(blsa_fig_i13, 
                top = text_grob("Trends in Metabolic indices over the Lifespan by Sex", size = 12, hjust = 0, x = unit(0, "lines")), 
                bottom = text_grob("Note: Diamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines")) 
                ) 
``` 

*** 

```{r blsa_race_dist_fig, warning=FALSE, message=FALSE, echo=FALSE}

annotate_figure(blsa_fig_j13, 
                top = text_grob("Trends in Metabolic indices over the Lifespan by Race", size = 12, hjust = 0, x = unit(0, "lines")), 
                bottom = text_grob("Note: Diamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines")) 
                ) 
``` 

*** 

```{r blsa_diab_dist_fig, warning=FALSE, message=FALSE, echo=FALSE}

annotate_figure(blsa_fig_diab13, 
                top = text_grob("Trends in Metabolic indices over the Lifespan by Health Status", size = 12, hjust = 0, x = unit(0, "lines")), 
                bottom = text_grob("Note: Diamonds represent age at peak index value", size = 9, hjust = 0, x = unit(0, "lines")) 
                )
``` 

*** 

#### pre and post-peak change creation 

```{r blsa_phase_creation, warning=FALSE, message=FALSE, echo=FALSE}

## assuming peak indices value is fixed (sd = 0) for rate standard deviation calculations

all_gluc <- NULL
f_gluc <- NULL
m_gluc <- NULL

#all
all_gluc$pretime <- (gluc_peak$all$per_age_wk - gluc_peak$lowbound$per_age_wk) 
all_gluc$prepeak_diff <- (gluc_peak$all$fit - gluc_peak$lowbound$fit)
all_gluc$prepeak_rate <- ((all_gluc$prepeak_diff / gluc_peak$lowbound$fit)/all_gluc$pretime)*5*100
all_gluc$prepeak_diff_lcl <- all_gluc$prepeak_diff - 1.96*gluc_peak$lowbound$se.fit
all_gluc$prepeak_diff_ucl <- all_gluc$prepeak_diff + 1.96*gluc_peak$lowbound$se.fit
all_gluc$prepeak_rate_lcl <- ((all_gluc$prepeak_diff_lcl / gluc_peak$lowbound$fit)/all_gluc$pretime)*5*100
all_gluc$prepeak_rate_ucl <- ((all_gluc$prepeak_diff_ucl / gluc_peak$lowbound$fit)/all_gluc$pretime)*5*100
all_gluc$prepeak_rate_se <- (all_gluc$prepeak_rate_ucl - all_gluc$prepeak_rate)/1.96
all_gluc$prepeak_rate_n <- gluc_peak$lowbound$n
all_gluc$prepeak_rate_sd <- all_gluc$prepeak_rate_se*sqrt(all_gluc$prepeak_rate_n) 

#female
f_gluc$pretime <- (gluc_peak_female$all$per_age_wk - gluc_peak_female$lowbound$per_age_wk) 
f_gluc$prepeak_diff <- (gluc_peak_female$all$fit - gluc_peak_female$lowbound$fit)
f_gluc$prepeak_rate <- ((f_gluc$prepeak_diff / gluc_peak_female$lowbound$fit)/f_gluc$pretime)*5*100
f_gluc$prepeak_diff_lcl <- f_gluc$prepeak_diff - 1.96*gluc_peak_female$lowbound$se.fit
f_gluc$prepeak_diff_ucl <- f_gluc$prepeak_diff + 1.96*gluc_peak_female$lowbound$se.fit
f_gluc$prepeak_rate_lcl <- ((f_gluc$prepeak_diff_lcl / gluc_peak_female$lowbound$fit)/f_gluc$pretime)*5*100
f_gluc$prepeak_rate_ucl <- ((f_gluc$prepeak_diff_ucl / gluc_peak_female$lowbound$fit)/f_gluc$pretime)*5*100
f_gluc$prepeak_rate_se <- (f_gluc$prepeak_rate_ucl - f_gluc$prepeak_rate)/1.96
f_gluc$prepeak_rate_n <- gluc_peak_female$lowbound$n
f_gluc$prepeak_rate_sd <- f_gluc$prepeak_rate_se*sqrt(f_gluc$prepeak_rate_n) 

#male
m_gluc$pretime <- (gluc_peak_male$all$per_age_wk - gluc_peak_male$lowbound$per_age_wk) 
m_gluc$prepeak_diff <- (gluc_peak_male$all$fit - gluc_peak_male$lowbound$fit)
m_gluc$prepeak_rate <- ((m_gluc$prepeak_diff / gluc_peak_male$lowbound$fit)/m_gluc$pretime)*5*100
m_gluc$prepeak_diff_lcl <- m_gluc$prepeak_diff - 1.96*gluc_peak_male$lowbound$se.fit
m_gluc$prepeak_diff_ucl <- m_gluc$prepeak_diff + 1.96*gluc_peak_male$lowbound$se.fit
m_gluc$prepeak_rate_lcl <- ((m_gluc$prepeak_diff_lcl / gluc_peak_male$lowbound$fit)/m_gluc$pretime)*5*100
m_gluc$prepeak_rate_ucl <- ((m_gluc$prepeak_diff_ucl / gluc_peak_male$lowbound$fit)/m_gluc$pretime)*5*100
m_gluc$prepeak_rate_se <- (m_gluc$prepeak_rate_ucl - m_gluc$prepeak_rate)/1.96
m_gluc$prepeak_rate_n <- gluc_peak_male$lowbound$n
m_gluc$prepeak_rate_sd <- m_gluc$prepeak_rate_se*sqrt(m_gluc$prepeak_rate_n) 

#all
all_gluc$posttime <- (gluc_peak$highbound$per_age_wk - gluc_peak$all$per_age_wk) 
all_gluc$postpeak_diff <- (gluc_peak$highbound$fit - gluc_peak$all$fit)
all_gluc$postpeak_rate <- ((all_gluc$postpeak_diff / gluc_peak$all$fit)/all_gluc$posttime)*5*100
all_gluc$postpeak_diff_lcl <- all_gluc$postpeak_diff - 1.96*gluc_peak$highbound$se.fit
all_gluc$postpeak_diff_ucl <- all_gluc$postpeak_diff + 1.96*gluc_peak$highbound$se.fit
all_gluc$postpeak_rate_lcl <- ((all_gluc$postpeak_diff_lcl / gluc_peak$all$fit)/all_gluc$posttime)*5*100
all_gluc$postpeak_rate_ucl <- ((all_gluc$postpeak_diff_ucl / gluc_peak$all$fit)/all_gluc$posttime)*5*100
all_gluc$postpeak_rate_se <- (all_gluc$postpeak_rate_ucl - all_gluc$postpeak_rate)/1.96
all_gluc$postpeak_rate_n <- gluc_peak$highbound$n 
all_gluc$postpeak_rate_sd <- all_gluc$postpeak_rate_se*sqrt(all_gluc$postpeak_rate_n) 

#female
f_gluc$posttime <- (gluc_peak_female$highbound$per_age_wk - gluc_peak_female$all$per_age_wk) 
f_gluc$postpeak_diff <- (gluc_peak_female$highbound$fit - gluc_peak_female$all$fit) 
f_gluc$postpeak_rate <- ((f_gluc$postpeak_diff / gluc_peak_female$all$fit)/f_gluc$posttime)*5*100 
f_gluc$postpeak_diff_lcl <- f_gluc$postpeak_diff - 1.96*gluc_peak_female$highbound$se.fit 
f_gluc$postpeak_diff_ucl <- f_gluc$postpeak_diff + 1.96*gluc_peak_female$highbound$se.fit 
f_gluc$postpeak_rate_lcl <- ((f_gluc$postpeak_diff_lcl / gluc_peak_female$all$fit)/f_gluc$posttime)*5*100 
f_gluc$postpeak_rate_ucl <- ((f_gluc$postpeak_diff_ucl / gluc_peak_female$all$fit)/f_gluc$posttime)*5*100 
f_gluc$postpeak_rate_se <- (f_gluc$postpeak_rate_ucl - f_gluc$postpeak_rate)/1.96
f_gluc$postpeak_rate_n <- gluc_peak_female$highbound$n 
f_gluc$postpeak_rate_sd <- f_gluc$postpeak_rate_se*sqrt(f_gluc$postpeak_rate_n) 

#male
m_gluc$posttime <- (gluc_peak_male$highbound$per_age_wk - gluc_peak_male$all$per_age_wk) 
m_gluc$postpeak_diff <- (gluc_peak_male$highbound$fit - gluc_peak_male$all$fit) 
m_gluc$postpeak_rate <- ((m_gluc$postpeak_diff / gluc_peak_male$all$fit)/m_gluc$posttime)*5*100 
m_gluc$postpeak_diff_lcl <- m_gluc$postpeak_diff - 1.96*gluc_peak_male$highbound$se.fit 
m_gluc$postpeak_diff_ucl <- m_gluc$postpeak_diff + 1.96*gluc_peak_male$highbound$se.fit 
m_gluc$postpeak_rate_lcl <- ((m_gluc$postpeak_diff_lcl / gluc_peak_male$all$fit)/m_gluc$posttime)*5*100 
m_gluc$postpeak_rate_ucl <- ((m_gluc$postpeak_diff_ucl / gluc_peak_male$all$fit)/m_gluc$posttime)*5*100 
m_gluc$postpeak_rate_se <- (m_gluc$postpeak_rate_ucl - m_gluc$postpeak_rate)/1.96
m_gluc$postpeak_rate_n <- gluc_peak_male$highbound$n 
m_gluc$postpeak_rate_sd <- m_gluc$postpeak_rate_se*sqrt(m_gluc$postpeak_rate_n) 

gluc2 <- as.data.frame(bind_rows(all_gluc, f_gluc, m_gluc)) %>% 
  mutate(group = c("all", "female", "male"), 
         index = c("gluc", "gluc", "gluc")) 

all_bw <- NULL
f_bw <- NULL
m_bw <- NULL

#all
all_bw$pretime <- (bw_peak$all$per_age_wk - bw_peak$lowbound$per_age_wk) 
all_bw$prepeak_diff <- (bw_peak$all$fit - bw_peak$lowbound$fit)
all_bw$prepeak_rate <- ((all_bw$prepeak_diff / bw_peak$lowbound$fit)/all_bw$pretime)*5*100
all_bw$prepeak_diff_lcl <- all_bw$prepeak_diff - 1.96*bw_peak$lowbound$se.fit
all_bw$prepeak_diff_ucl <- all_bw$prepeak_diff + 1.96*bw_peak$lowbound$se.fit
all_bw$prepeak_rate_lcl <- ((all_bw$prepeak_diff_lcl / bw_peak$lowbound$fit)/all_bw$pretime)*5*100
all_bw$prepeak_rate_ucl <- ((all_bw$prepeak_diff_ucl / bw_peak$lowbound$fit)/all_bw$pretime)*5*100
all_bw$prepeak_rate_se <- (all_bw$prepeak_rate_ucl - all_bw$prepeak_rate)/1.96
all_bw$prepeak_rate_n <- bw_peak$lowbound$n
all_bw$prepeak_rate_sd <- all_bw$prepeak_rate_se*sqrt(all_bw$prepeak_rate_n) 

#female
f_bw$pretime <- (bw_peak_female$all$per_age_wk - bw_peak_female$lowbound$per_age_wk) 
f_bw$prepeak_diff <- (bw_peak_female$all$fit - bw_peak_female$lowbound$fit)
f_bw$prepeak_rate <- ((f_bw$prepeak_diff / bw_peak_female$lowbound$fit)/f_bw$pretime)*5*100
f_bw$prepeak_diff_lcl <- f_bw$prepeak_diff - 1.96*bw_peak_female$lowbound$se.fit
f_bw$prepeak_diff_ucl <- f_bw$prepeak_diff + 1.96*bw_peak_female$lowbound$se.fit
f_bw$prepeak_rate_lcl <- ((f_bw$prepeak_diff_lcl / bw_peak_female$lowbound$fit)/f_bw$pretime)*5*100
f_bw$prepeak_rate_ucl <- ((f_bw$prepeak_diff_ucl / bw_peak_female$lowbound$fit)/f_bw$pretime)*5*100
f_bw$prepeak_rate_se <- (f_bw$prepeak_rate_ucl - f_bw$prepeak_rate)/1.96
f_bw$prepeak_rate_n <- bw_peak_female$lowbound$n
f_bw$prepeak_rate_sd <- f_bw$prepeak_rate_se*sqrt(f_bw$prepeak_rate_n) 

#male
m_bw$pretime <- (bw_peak_male$all$per_age_wk - bw_peak_male$lowbound$per_age_wk) 
m_bw$prepeak_diff <- (bw_peak_male$all$fit - bw_peak_male$lowbound$fit)
m_bw$prepeak_rate <- ((m_bw$prepeak_diff / bw_peak_male$lowbound$fit)/m_bw$pretime)*5*100
m_bw$prepeak_diff_lcl <- m_bw$prepeak_diff - 1.96*bw_peak_male$lowbound$se.fit
m_bw$prepeak_diff_ucl <- m_bw$prepeak_diff + 1.96*bw_peak_male$lowbound$se.fit
m_bw$prepeak_rate_lcl <- ((m_bw$prepeak_diff_lcl / bw_peak_male$lowbound$fit)/m_bw$pretime)*5*100
m_bw$prepeak_rate_ucl <- ((m_bw$prepeak_diff_ucl / bw_peak_male$lowbound$fit)/m_bw$pretime)*5*100
m_bw$prepeak_rate_se <- (m_bw$prepeak_rate_ucl - m_bw$prepeak_rate)/1.96
m_bw$prepeak_rate_n <- bw_peak_male$lowbound$n
m_bw$prepeak_rate_sd <- m_bw$prepeak_rate_se*sqrt(m_bw$prepeak_rate_n) 

#all
all_bw$posttime <- (bw_peak$highbound$per_age_wk - bw_peak$all$per_age_wk) 
all_bw$postpeak_diff <- (bw_peak$highbound$fit - bw_peak$all$fit)
all_bw$postpeak_rate <- ((all_bw$postpeak_diff / bw_peak$all$fit)/all_bw$posttime)*5*100
all_bw$postpeak_diff_lcl <- all_bw$postpeak_diff - 1.96*bw_peak$highbound$se.fit
all_bw$postpeak_diff_ucl <- all_bw$postpeak_diff + 1.96*bw_peak$highbound$se.fit
all_bw$postpeak_rate_lcl <- ((all_bw$postpeak_diff_lcl / bw_peak$all$fit)/all_bw$posttime)*5*100
all_bw$postpeak_rate_ucl <- ((all_bw$postpeak_diff_ucl / bw_peak$all$fit)/all_bw$posttime)*5*100
all_bw$postpeak_rate_se <- (all_bw$postpeak_rate_ucl - all_bw$postpeak_rate)/1.96
all_bw$postpeak_rate_n <- bw_peak$highbound$n 
all_bw$postpeak_rate_sd <- all_bw$postpeak_rate_se*sqrt(all_bw$postpeak_rate_n) 

#female
f_bw$posttime <- (bw_peak_female$highbound$per_age_wk - bw_peak_female$all$per_age_wk) 
f_bw$postpeak_diff <- (bw_peak_female$highbound$fit - bw_peak_female$all$fit) 
f_bw$postpeak_rate <- ((f_bw$postpeak_diff / bw_peak_female$all$fit)/f_bw$posttime)*5*100 
f_bw$postpeak_diff_lcl <- f_bw$postpeak_diff - 1.96*bw_peak_female$highbound$se.fit 
f_bw$postpeak_diff_ucl <- f_bw$postpeak_diff + 1.96*bw_peak_female$highbound$se.fit 
f_bw$postpeak_rate_lcl <- ((f_bw$postpeak_diff_lcl / bw_peak_female$all$fit)/f_bw$posttime)*5*100 
f_bw$postpeak_rate_ucl <- ((f_bw$postpeak_diff_ucl / bw_peak_female$all$fit)/f_bw$posttime)*5*100 
f_bw$postpeak_rate_se <- (f_bw$postpeak_rate_ucl - f_bw$postpeak_rate)/1.96
f_bw$postpeak_rate_n <- bw_peak_female$highbound$n 
f_bw$postpeak_rate_sd <- f_bw$postpeak_rate_se*sqrt(f_bw$postpeak_rate_n) 

#male
m_bw$posttime <- (bw_peak_male$highbound$per_age_wk - bw_peak_male$all$per_age_wk) 
m_bw$postpeak_diff <- (bw_peak_male$highbound$fit - bw_peak_male$all$fit) 
m_bw$postpeak_rate <- ((m_bw$postpeak_diff / bw_peak_male$all$fit)/m_bw$posttime)*5*100 
m_bw$postpeak_diff_lcl <- m_bw$postpeak_diff - 1.96*bw_peak_male$highbound$se.fit 
m_bw$postpeak_diff_ucl <- m_bw$postpeak_diff + 1.96*bw_peak_male$highbound$se.fit 
m_bw$postpeak_rate_lcl <- ((m_bw$postpeak_diff_lcl / bw_peak_male$all$fit)/m_bw$posttime)*5*100 
m_bw$postpeak_rate_ucl <- ((m_bw$postpeak_diff_ucl / bw_peak_male$all$fit)/m_bw$posttime)*5*100 
m_bw$postpeak_rate_se <- (m_bw$postpeak_rate_ucl - m_bw$postpeak_rate)/1.96
m_bw$postpeak_rate_n <- bw_peak_male$highbound$n 
m_bw$postpeak_rate_sd <- m_bw$postpeak_rate_se*sqrt(m_bw$postpeak_rate_n) 

bw2 <- as.data.frame(bind_rows(all_bw, f_bw, m_bw)) %>% 
  mutate(group = c("all", "female", "male"), 
         index = c("bw", "bw", "bw")) 

all_fat <- NULL
f_fat <- NULL
m_fat <- NULL

#all
all_fat$pretime <- (fat_peak$all$per_age_wk - fat_peak$lowbound$per_age_wk) 
all_fat$prepeak_diff <- (fat_peak$all$fit - fat_peak$lowbound$fit)
all_fat$prepeak_rate <- ((all_fat$prepeak_diff / fat_peak$lowbound$fit)/all_fat$pretime)*5*100
all_fat$prepeak_diff_lcl <- all_fat$prepeak_diff - 1.96*fat_peak$lowbound$se.fit
all_fat$prepeak_diff_ucl <- all_fat$prepeak_diff + 1.96*fat_peak$lowbound$se.fit
all_fat$prepeak_rate_lcl <- ((all_fat$prepeak_diff_lcl / fat_peak$lowbound$fit)/all_fat$pretime)*5*100
all_fat$prepeak_rate_ucl <- ((all_fat$prepeak_diff_ucl / fat_peak$lowbound$fit)/all_fat$pretime)*5*100
all_fat$prepeak_rate_se <- (all_fat$prepeak_rate_ucl - all_fat$prepeak_rate)/1.96
all_fat$prepeak_rate_n <- fat_peak$lowbound$n
all_fat$prepeak_rate_sd <- all_fat$prepeak_rate_se*sqrt(all_fat$prepeak_rate_n) 

#female
f_fat$pretime <- (fat_peak_female$all$per_age_wk - fat_peak_female$lowbound$per_age_wk) 
f_fat$prepeak_diff <- (fat_peak_female$all$fit - fat_peak_female$lowbound$fit)
f_fat$prepeak_rate <- ((f_fat$prepeak_diff / fat_peak_female$lowbound$fit)/f_fat$pretime)*5*100
f_fat$prepeak_diff_lcl <- f_fat$prepeak_diff - 1.96*fat_peak_female$lowbound$se.fit
f_fat$prepeak_diff_ucl <- f_fat$prepeak_diff + 1.96*fat_peak_female$lowbound$se.fit
f_fat$prepeak_rate_lcl <- ((f_fat$prepeak_diff_lcl / fat_peak_female$lowbound$fit)/f_fat$pretime)*5*100
f_fat$prepeak_rate_ucl <- ((f_fat$prepeak_diff_ucl / fat_peak_female$lowbound$fit)/f_fat$pretime)*5*100
f_fat$prepeak_rate_se <- (f_fat$prepeak_rate_ucl - f_fat$prepeak_rate)/1.96
f_fat$prepeak_rate_n <- fat_peak_female$lowbound$n
f_fat$prepeak_rate_sd <- f_fat$prepeak_rate_se*sqrt(f_fat$prepeak_rate_n) 

#male
m_fat$pretime <- (fat_peak_male$all$per_age_wk - fat_peak_male$lowbound$per_age_wk) 
m_fat$prepeak_diff <- (fat_peak_male$all$fit - fat_peak_male$lowbound$fit)
m_fat$prepeak_rate <- ((m_fat$prepeak_diff / fat_peak_male$lowbound$fit)/m_fat$pretime)*5*100
m_fat$prepeak_diff_lcl <- m_fat$prepeak_diff - 1.96*fat_peak_male$lowbound$se.fit
m_fat$prepeak_diff_ucl <- m_fat$prepeak_diff + 1.96*fat_peak_male$lowbound$se.fit
m_fat$prepeak_rate_lcl <- ((m_fat$prepeak_diff_lcl / fat_peak_male$lowbound$fit)/m_fat$pretime)*5*100
m_fat$prepeak_rate_ucl <- ((m_fat$prepeak_diff_ucl / fat_peak_male$lowbound$fit)/m_fat$pretime)*5*100
m_fat$prepeak_rate_se <- (m_fat$prepeak_rate_ucl - m_fat$prepeak_rate)/1.96
m_fat$prepeak_rate_n <- fat_peak_male$lowbound$n
m_fat$prepeak_rate_sd <- m_fat$prepeak_rate_se*sqrt(m_fat$prepeak_rate_n) 

#all
all_fat$posttime <- (fat_peak$highbound$per_age_wk - fat_peak$all$per_age_wk) 
all_fat$postpeak_diff <- (fat_peak$highbound$fit - fat_peak$all$fit)
all_fat$postpeak_rate <- ((all_fat$postpeak_diff / fat_peak$all$fit)/all_fat$posttime)*5*100
all_fat$postpeak_diff_lcl <- all_fat$postpeak_diff - 1.96*fat_peak$all$se.fit
all_fat$postpeak_diff_ucl <- all_fat$postpeak_diff + 1.96*fat_peak$all$se.fit
all_fat$postpeak_rate_lcl <- ((all_fat$postpeak_diff_lcl / fat_peak$all$fit)/all_fat$posttime)*5*100
all_fat$postpeak_rate_ucl <- ((all_fat$postpeak_diff_ucl / fat_peak$all$fit)/all_fat$posttime)*5*100
all_fat$postpeak_rate_se <- (all_fat$postpeak_rate_ucl - all_fat$postpeak_rate)/1.96
all_fat$postpeak_rate_n <- fat_peak$highbound$n 
all_fat$postpeak_rate_sd <- all_fat$postpeak_rate_se*sqrt(all_fat$postpeak_rate_n) 

#female
f_fat$posttime <- (fat_peak_female$highbound$per_age_wk - fat_peak_female$all$per_age_wk) 
f_fat$postpeak_diff <- (fat_peak_female$highbound$fit - fat_peak_female$all$fit) 
f_fat$postpeak_rate <- ((f_fat$postpeak_diff / fat_peak_female$all$fit)/f_fat$posttime)*5*100 
f_fat$postpeak_diff_lcl <- f_fat$postpeak_diff - 1.96*fat_peak_female$highbound$se.fit 
f_fat$postpeak_diff_ucl <- f_fat$postpeak_diff + 1.96*fat_peak_female$highbound$se.fit 
f_fat$postpeak_rate_lcl <- ((f_fat$postpeak_diff_lcl / fat_peak_female$all$fit)/f_fat$posttime)*5*100 
f_fat$postpeak_rate_ucl <- ((f_fat$postpeak_diff_ucl / fat_peak_female$all$fit)/f_fat$posttime)*5*100 
f_fat$postpeak_rate_se <- (f_fat$postpeak_rate_ucl - f_fat$postpeak_rate)/1.96
f_fat$postpeak_rate_n <- fat_peak_female$highbound$n 
f_fat$postpeak_rate_sd <- f_fat$postpeak_rate_se*sqrt(f_fat$postpeak_rate_n) 

#male
m_fat$posttime <- (fat_peak_male$highbound$per_age_wk - fat_peak_male$all$per_age_wk) 
m_fat$postpeak_diff <- (fat_peak_male$highbound$fit - fat_peak_male$all$fit) 
m_fat$postpeak_rate <- ((m_fat$postpeak_diff / fat_peak_male$all$fit)/m_fat$posttime)*5*100 
m_fat$postpeak_diff_lcl <- m_fat$postpeak_diff - 1.96*fat_peak_male$highbound$se.fit 
m_fat$postpeak_diff_ucl <- m_fat$postpeak_diff + 1.96*fat_peak_male$highbound$se.fit 
m_fat$postpeak_rate_lcl <- ((m_fat$postpeak_diff_lcl / fat_peak_male$all$fit)/m_fat$posttime)*5*100 
m_fat$postpeak_rate_ucl <- ((m_fat$postpeak_diff_ucl / fat_peak_male$all$fit)/m_fat$posttime)*5*100 
m_fat$postpeak_rate_se <- (m_fat$postpeak_rate_ucl - m_fat$postpeak_rate)/1.96
m_fat$postpeak_rate_n <- fat_peak_male$highbound$n 
m_fat$postpeak_rate_sd <- m_fat$postpeak_rate_se*sqrt(m_fat$postpeak_rate_n) 

fat2 <- as.data.frame(bind_rows(all_fat, f_fat, m_fat)) %>% 
  mutate(group = c("all", "female", "male"), 
         index = c("fat", "fat", "fat")) 

##
phases <- rbind(gluc2, bw2, fat2) %>% 
  mutate(
    
    prepeak_pval = tsum.test(prepeak_rate, prepeak_rate_sd, prepeak_rate_n, mu = 0, conf.level = 0.95)$p.value,
    postpeak_pval = tsum.test(postpeak_rate, postpeak_rate_sd, postpeak_rate_n, mu = 0, conf.level = 0.95)$p.value,
    
    prepeak_mean_95CI =
      
      ifelse(is.na(prepeak_pval),
    paste0(format(round(prepeak_rate, 2), nsmall=2)," ","(",format(round(prepeak_rate_lcl, 2), nsmall=2),", ",format(round(prepeak_rate_ucl, 2), nsmall=2),")"),
      ifelse(prepeak_pval < 0.050,
    paste0(format(round(prepeak_rate, 2), nsmall=2)," ","(",format(round(prepeak_rate_lcl, 2), nsmall=2),", ",format(round(prepeak_rate_ucl, 2), nsmall=2),")*"),
    paste0(format(round(prepeak_rate, 2), nsmall=2)," ","(",format(round(prepeak_rate_lcl, 2), nsmall=2),", ",format(round(prepeak_rate_ucl, 2), nsmall=2),")"))),

    postpeak_mean_95CI = 
      ifelse(is.na(postpeak_pval),
    paste0(format(round(postpeak_rate, 2), nsmall=2)," ","(",format(round(postpeak_rate_lcl, 2), nsmall=2),", ",format(round(postpeak_rate_ucl, 2), nsmall=2),")"),
      ifelse(postpeak_pval < 0.050,
    paste0(format(round(postpeak_rate, 2), nsmall=2)," ","(",format(round(postpeak_rate_lcl, 2), nsmall=2),", ",format(round(postpeak_rate_ucl, 2), nsmall=2),")*"),
    paste0(format(round(postpeak_rate, 2), nsmall=2)," ","(",format(round(postpeak_rate_lcl, 2), nsmall=2),", ",format(round(postpeak_rate_ucl, 2), nsmall=2),")")))
    
  )

gluc_prepeak <- phases %>% 
  filter(index == "gluc") %>% 
  select(index, group, contains("95"), -contains("postpeak")) %>% 
  spread(key = group, value = prepeak_mean_95CI) 
gluc_postpeak <- phases %>% 
  filter(index == "gluc") %>% 
  select(index, group, contains("95"), -contains("prepeak")) %>% 
  spread(key = group, value = postpeak_mean_95CI) 

bw_prepeak <- phases %>% 
  filter(index == "bw") %>% 
  select(index, group, contains("95"), -contains("postpeak")) %>% 
  spread(key = group, value = prepeak_mean_95CI) 
bw_postpeak <- phases %>% 
  filter(index == "bw") %>% 
  select(index, group, contains("95"), -contains("prepeak")) %>% 
  spread(key = group, value = postpeak_mean_95CI) 

fat_prepeak <- phases %>% 
  filter(index == "fat") %>% 
  select(index, group, contains("95"), -contains("postpeak")) %>% 
  spread(key = group, value = prepeak_mean_95CI) 
fat_postpeak <- phases %>% 
  filter(index == "fat") %>% 
  select(index, group, contains("95"), -contains("prepeak")) %>% 
  spread(key = group, value = postpeak_mean_95CI) 

gluc_all <- bind_cols(gluc_prepeak, gluc_postpeak)
bw_all <- bind_cols(bw_prepeak, bw_postpeak)
fat_all <- bind_cols(fat_prepeak, fat_postpeak)

phases2 <- bind_rows(gluc_all, bw_all, fat_all) 
colnames(phases2) <- c("index", "all", "female", "male", "index2", "all2", "female2", "male2")

##peaks

#all
gluc_peak_ci <- paste0(round(gluc_peak$all$per_age_wk, 2)," (", round(gluc_peak$min$per_age_wk, 2), ", ", round(gluc_peak$max$per_age_wk, 2), ")") 
bw_peak_ci <- paste0(round(bw_peak$all$per_age_wk, 2)," (", round(bw_peak$min$per_age_wk, 2), ", ", round(bw_peak$max$per_age_wk, 2), ")") 
fat_peak_ci <- paste0(round(fat_peak$all$per_age_wk, 2)," (", round(fat_peak$min$per_age_wk, 2), ", ", round(fat_peak$max$per_age_wk, 2), ")") 

#male
male_gluc_peak_ci <- paste0(round(gluc_peak_male$all$per_age_wk, 2)," (", round(gluc_peak_male$min$per_age_wk, 2), ", ", round(gluc_peak_male$max$per_age_wk, 2), ")") 
male_bw_peak_ci <- paste0(round(bw_peak_male$all$per_age_wk, 2)," (", round(bw_peak_male$min$per_age_wk, 2), ", ", round(bw_peak_male$max$per_age_wk, 2), ")") 
male_fat_peak_ci <- paste0(round(fat_peak_male$all$per_age_wk, 2)," (", round(fat_peak_male$min$per_age_wk, 2), ", ", round(fat_peak_male$max$per_age_wk, 2), ")") 

#female
female_gluc_peak_ci <- paste0(round(gluc_peak_female$all$per_age_wk, 2)," (", round(gluc_peak_female$min$per_age_wk, 2), ", ", round(gluc_peak_female$max$per_age_wk, 2), ")") 
female_bw_peak_ci <- paste0(round(bw_peak_female$all$per_age_wk, 2)," (", round(bw_peak_female$min$per_age_wk, 2), ", ", round(bw_peak_female$max$per_age_wk, 2), ")") 
female_fat_peak_ci <- paste0(round(fat_peak_female$all$per_age_wk, 2)," (", round(fat_peak_female$min$per_age_wk, 2), ", ", round(fat_peak_female$max$per_age_wk, 2), ")") 

#final datasets
phases2$peak_all <- c(gluc_peak_ci, bw_peak_ci, fat_peak_ci) 
phases2$peak_female <- c(female_gluc_peak_ci, female_bw_peak_ci, female_fat_peak_ci) 
phases2$peak_male <- c(male_gluc_peak_ci, male_bw_peak_ci, male_fat_peak_ci)

phases3 <- phases2 %>% 
  select(index, contains("peak"), everything(), -index2) 

```

```{r blsa_phase_formatting, warning=FALSE, message=FALSE, echo=FALSE}

#html output formatting
phases4 <- phases3 %>% 
  mutate( 
    peak_all = noquote((gsub("\\(", "<br>(", peak_all))), 
    peak_female = noquote((gsub("\\(", "<br>(", peak_female))), 
    peak_male = noquote((gsub("\\(", "<br>(", peak_male))), 
    all = noquote((gsub("\\(", "<br>(", all))), 
    female = noquote((gsub("\\(", "<br>(", female))), 
    male = noquote((gsub("\\(", "<br>(", male))), 
    all2 = noquote((gsub("\\(", "<br>(", all2))), 
    female2 = noquote((gsub("\\(", "<br>(", female2))), 
    male2 = noquote((gsub("\\(", "<br>(", male2)))
    ) 

blsa_phases <- phases4 

```

#### change in metabolic indices pre and post-peak 
- rate of percent change 

```{r blsa_phase_table_prate, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE}

blsa_phases_out <- kable(blsa_phases, row.names = FALSE, col.names = c("Metabolic <br> Index", "All", "Female", "Male", "All", "Female", "Male", "All", "Female", "Male"), align = c("c","c","c","c","c","c","c"), escape = FALSE, 
      caption = "Rate of Percent Change for Metabolic indices Pre- and Post-peak") %>% 
  kable_styling(bootstrap_options = "striped", full_width = TRUE, position = "left", font_size = 11) %>% 
  column_spec(1, bold = TRUE, border_right = TRUE) %>% 
  column_spec(4, border_right = TRUE) %>% 
  column_spec(7, border_right = TRUE) %>% 
  add_header_above(c(" " = 1, "Age at Peak Value$^$ \n(% life expectancy)" = 3, "Pre-Peak$^$$^$" = 3, "Post-Peak$^$$^$" = 3)) %>% 
  footnote(symbol = c(
    "p value < 0.05",
    "Insufficient sample size to conduct a statistical test of significance",
    "Mean (range of ages at similar values) of age at peak index value", 
    "Mean (95% confidence interval) rate of percent change during phase per five percent change in life expectancy", 
    "Pre-peak is defined as 25% life expectancy to age at peak value; post-peak phase is defined as peak to 125% life expectancy",
    "Male and female estimates are based on sex specific life expectancy")) 

blsa_phases_out
```

*** 

#### quartile creation

```{r blsa_scaling, warning=FALSE, message=FALSE, echo=FALSE}

main_cont_surv <- main_all3 %>% select(idno, sex, race, per_age_wk_25, per_age_wk_50, per_age_wk_75, per_age_wk_100, per_age_wk_125) %>% distinct() %>% inner_join(census_c16_death) 

main_cat_surv_f <- main_cat2 %>% inner_join(main_cont_surv) %>% filter(sex == "F") 
main_cat_surv_m <- main_cat2 %>% inner_join(main_cont_surv) %>% filter(sex == "M") 

```
 
```{r blsa_quartiles, warning=FALSE, message=FALSE, echo=FALSE}

quart_breaks = c(0, 0.25, 0.50, 0.75, 1.0) 
quart_labs = c("Low", "Medlow", "Medhigh", "High")

## female
main_cat_surv_f$gluc_50_quart <- cut(main_cat_surv_f$gluc_50, quantile(main_cat_surv_f$gluc_50, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$gluc_75_quart <- cut(main_cat_surv_f$gluc_75, quantile(main_cat_surv_f$gluc_75, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$gluc_100_quart <- cut(main_cat_surv_f$gluc_100, quantile(main_cat_surv_f$gluc_100, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_f$bw_50_quart <- cut(main_cat_surv_f$bw_50, quantile(main_cat_surv_f$bw_50, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$bw_75_quart <- cut(main_cat_surv_f$bw_75, quantile(main_cat_surv_f$bw_75, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$bw_100_quart <- cut(main_cat_surv_f$bw_100, quantile(main_cat_surv_f$bw_100, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_f$fat_50_quart <- cut(main_cat_surv_f$fat_50, quantile(main_cat_surv_f$fat_50, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$fat_75_quart <- cut(main_cat_surv_f$fat_75, quantile(main_cat_surv_f$fat_75, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_f$fat_100_quart <- cut(main_cat_surv_f$fat_100, quantile(main_cat_surv_f$fat_100, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

## male
main_cat_surv_m$gluc_50_quart <- cut(main_cat_surv_m$gluc_50, quantile(main_cat_surv_m$gluc_50, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$gluc_75_quart <- cut(main_cat_surv_m$gluc_75, quantile(main_cat_surv_m$gluc_75, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$gluc_100_quart <- cut(main_cat_surv_m$gluc_100, quantile(main_cat_surv_m$gluc_100, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_m$bw_50_quart <- cut(main_cat_surv_m$bw_50, quantile(main_cat_surv_m$bw_50, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$bw_75_quart <- cut(main_cat_surv_m$bw_75, quantile(main_cat_surv_m$bw_75, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$bw_100_quart <- cut(main_cat_surv_m$bw_100, quantile(main_cat_surv_m$bw_100, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

main_cat_surv_m$fat_50_quart <- cut(main_cat_surv_m$fat_50, quantile(main_cat_surv_m$fat_50, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$fat_75_quart <- cut(main_cat_surv_m$fat_75, quantile(main_cat_surv_m$fat_75, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 
main_cat_surv_m$fat_100_quart <- cut(main_cat_surv_m$fat_100, quantile(main_cat_surv_m$fat_100, breaks = quart_breaks, na.rm=T, include.lowest=T), labels = quart_labs) 

#merge into all dataset 
main_cat_surv <- main_cat_surv_f %>% 
  rbind(main_cat_surv_m) 

```
 
#### kaplan meier survival curves 
- comparing highest to lowest quartile 
- three timepoints: 50, 75, and 100% life expectancy 

```{r blsa_surv_curve, warning=FALSE, message=FALSE, echo=FALSE}

surv_color <- c("orange", "purple")

km_funct <- function(quart, exp_var, start, end, data, title, env = globalenv()){

#subset data
surv_df <- as.data.frame(data) 
surv_df_plot <- subset(data, eval(parse(text = quart)) %in% c("Low", "High"))
surv_df_plot2 <- as.data.frame(surv_df_plot) 

#change level order for legend
surv_df_plot3 <- surv_df_plot2 %>% 
  mutate_at( 
    vars(contains("quart")), 
    funs(fct_rev(.)) 
  )  

#pull input variable for quartiles
env$quart <- quart 
env$explanatory <- NULL 
env$explanatory$var <- exp_var 

#cph model
formula_cph <- as.formula(paste0("Surv(per_le_wk, dead_censor) ~", env$quart, "+", paste(env$explanatory$var, collapse = " + "))) 
cph <- coxph(formula_cph, data = surv_df) 
km_fit <- survfit(Surv(per_le_wk, dead_censor) ~ eval(parse(text = quart)), data = surv_df_plot3) 

cph_summary <- summary(cph) 
value <- format(round(cph_summary$conf.int[,"exp(coef)"][3], 2), nsmall = 2)
le <- format(round(cph_summary$conf.int[,"lower .95"][3], 2), nsmall = 2)
ue <- format(round(cph_summary$conf.int[,"upper .95"][3], 2), nsmall = 2)
pval <- cph_summary$coefficients[,"Pr(>|z|)"][3]

#place asterisk for significant pval
if(pval < 0.050){
  hr <- paste0(value, " (", le, ", ", ue, ")*") 
}else{
  hr <- paste0(value, " (", le, ", ", ue, ")") 
}

km_curve <- ggsurvplot(
    fit = km_fit, 
    data = surv_df_plot3, 
    title = title,
    font.title = c(12), 
    conf.int = FALSE,
    risk.table = TRUE, 
    censor = FALSE,
    risk.table.col = "strata",
    legend.title = "Index Subgroup", 
    ggtheme = theme_bw(), 
    legend.labs = c("Highest Quartile", "Lowest Quartile"),
    palette = surv_color, 
    xlab = "Percent Life Expectancy", 
    ylab = "Cumulative Survival\n", 
    xlim = c(start, end)

) 
surv_results <- list(km_curve, hr, pval, cph_summary) 

return(surv_results) 

} 

#Survival
km_start_50 <- 47.5
km_start_75 <- 72.5
km_start_100 <- 97.5

#function calls 
blsa_surv_gluc_50_all <- km_funct(quart = "gluc_50_quart", exp_var = c("per_age_wk_50"), start = km_start_50, data = main_cat_surv, title = "Glucose at 50% Life Expectancy", end = 180) 
blsa_surv_gluc_75_all <- km_funct(quart = "gluc_75_quart", exp_var = c("per_age_wk_75"), start = km_start_75, data = main_cat_surv, title = "Glucose at 75% Life Expectancy", end = 180) 
blsa_surv_gluc_100_all <- km_funct(quart = "gluc_100_quart", exp_var = c("per_age_wk_100"), start = km_start_100, data = main_cat_surv, title = "Glucose at 100% Life Expectancy", end = 180) 

blsa_surv_bw_50_all <- km_funct(quart = "bw_50_quart", exp_var = c("per_age_wk_50"), start = km_start_50, data = main_cat_surv, title = "Body Weight at 50% Life Expectancy", end = 180) 
blsa_surv_bw_75_all <- km_funct(quart = "bw_75_quart", exp_var = c("per_age_wk_75"), start = km_start_75, data = main_cat_surv, title = "Body Weight at 75% Life Expectancy", end = 180) 
blsa_surv_bw_100_all <- km_funct(quart = "bw_100_quart", exp_var = c("per_age_wk_100"), start = km_start_100, data = main_cat_surv, title = "Body Weight at 100% Life Expectancy", end = 180) 

blsa_surv_fat_50_all <- km_funct(quart = "fat_50_quart", exp_var = c("per_age_wk_50"), start = km_start_50, data = main_cat_surv, title = "Fat at 50% Life Expectancy", end = 180) 
blsa_surv_fat_75_all <- km_funct(quart = "fat_75_quart", exp_var = c("per_age_wk_75"), start = km_start_75, data = main_cat_surv, title = "Fat at 75% Life Expectancy", end = 180) 
blsa_surv_fat_100_all <- km_funct(quart = "fat_100_quart", exp_var = c("per_age_wk_100"), start = km_start_100, data = main_cat_surv, title = "Fat at 100% Life Expectancy", end = 180) 

```

```{r blsa_surv_output, warning=FALSE, message=FALSE, echo=FALSE}

#change font
nytimes <- "Verdana"

#add adjustment subscript
blsa_km_gluc_50 <- blsa_surv_gluc_50_all[[1]]$plot + ylim(-0.07, 1) + 
  annotate("text", x = -Inf, y = -0.07, hjust = -.09, label = deparse(bquote(HR[adj]== .(blsa_surv_gluc_50_all[[2]]))), size = 4, parse =TRUE, family = nytimes)
blsa_km_gluc_75 <- blsa_surv_gluc_75_all[[1]]$plot + ylim(-0.07, 1) + 
  annotate("text", x = -Inf, y = -0.07, hjust = -.09, label = deparse(bquote(HR[adj]== .(blsa_surv_gluc_75_all[[2]]))), size = 4, parse =TRUE, family = nytimes)
blsa_km_gluc_100 <- blsa_surv_gluc_100_all[[1]]$plot + ylim(-0.07, 1) + 
  annotate("text", x = -Inf, y = -0.07, hjust = -.09, label = deparse(bquote(HR[adj]== .(blsa_surv_gluc_100_all[[2]]))), size = 4, parse =TRUE, family = nytimes)

blsa_km_bw_50 <- blsa_surv_bw_50_all[[1]]$plot + ylim(-0.07, 1) + 
  annotate("text", x = -Inf, y = -0.07, hjust = -.09, label = deparse(bquote(HR[adj]== .(blsa_surv_bw_50_all[[2]]))), size = 4, parse =TRUE, family = nytimes)
blsa_km_bw_75 <- blsa_surv_bw_75_all[[1]]$plot + ylim(-0.07, 1) + 
  annotate("text", x = -Inf, y = -0.07, hjust = -.09, label = deparse(bquote(HR[adj]== .(blsa_surv_bw_75_all[[2]]))), size = 4, parse =TRUE, family = nytimes)
blsa_km_bw_100 <- blsa_surv_bw_100_all[[1]]$plot + ylim(-0.07, 1) + 
  annotate("text", x = -Inf, y = -0.07, hjust = -.09, label = deparse(bquote(HR[adj]== .(blsa_surv_bw_100_all[[2]]))), size = 4, parse =TRUE, family = nytimes)

blsa_km_fat_50 <- blsa_surv_fat_50_all[[1]]$plot + ylim(-0.07, 1) + 
  annotate("text", x = -Inf, y = -0.07, hjust = -.09, label = deparse(bquote(HR[adj]== .(blsa_surv_fat_50_all[[2]]))), size = 4, parse =TRUE, family = nytimes)
blsa_km_fat_75 <- blsa_surv_fat_75_all[[1]]$plot + ylim(-0.07, 1) + 
  annotate("text", x = -Inf, y = -0.07, hjust = -.09, label = deparse(bquote(HR[adj]== .(blsa_surv_fat_75_all[[2]]))), size = 4, parse =TRUE, family = nytimes)
blsa_km_fat_100 <- blsa_surv_fat_100_all[[1]]$plot + ylim(-0.07, 1) + 
  annotate("text", x = -Inf, y = -0.07, hjust = -.09, label = deparse(bquote(HR[adj]== .(blsa_surv_fat_100_all[[2]]))), size = 4, parse =TRUE, family = nytimes)

km_all <- ggarrange(
  blsa_km_gluc_50, blsa_km_gluc_75, blsa_km_gluc_100, blsa_km_bw_50, blsa_km_bw_75, blsa_km_bw_100, blsa_km_fat_50, blsa_km_fat_75, blsa_km_fat_100,
  common.legend = TRUE, legend = "right") 

```

```{r blsa_surv_kable, fig.height = 12, fig.width = 12, warning=FALSE, message=FALSE, echo=FALSE}

km_all_output <- annotate_figure(km_all,
                                 top = text_grob("Survival Curve and Age Adjusted Mortality Hazard Ratio Comparing Highest and Lowest Metabolic Index Quartiles \n at Varying Timepoints\n", size = 12,  hjust = 0.52) 
                ) 

km_all_output 

```

*** 

#### cox proportional mortality hazard ratios 
- sex specific
- covariate adjustment
  - sex and age 
  - additional adjustment for body fat 

```{r blsa_cph_fun, warning=FALSE, message=FALSE, echo=FALSE}

cph_funct <- function(quart, exp_var, data, env = globalenv()){

surv_df <- as.data.frame(data) 

env$quart <- quart 
env$explanatory <- NULL 
env$explanatory$var <- exp_var 

formula_cph <- as.formula(paste0("Surv(per_le_wk, dead_censor) ~", env$quart, "+", paste(env$explanatory$var, collapse = " + "))) 
cph <- coxph(formula_cph, data = surv_df) 

cph_summary <- summary(cph) 
value <- format(round(cph_summary$conf.int[,"exp(coef)"][3], 2), nsmall = 2) 
le <- format(round(cph_summary$conf.int[,"lower .95"][3], 2), nsmall = 2) 
ue <- format(round(cph_summary$conf.int[,"upper .95"][3], 2), nsmall = 2) 
pval <- cph_summary$coefficients[,"Pr(>|z|)"][3] 

if(pval < 0.050){
  hr <- paste0("HR = ", value, " (", le, ", ", ue, ")*") 
}else{
  hr <- paste0("HR = ", value, " (", le, ", ", ue, ")") 
}
 
surv_results <- list(hr, pval, cph_summary) 

return(surv_results) 

} 

#function calls 
blsa_surv_gluc_50_all2 <- cph_funct(quart = "gluc_50_quart", exp_var = c("per_age_wk_50",  "sex"), data = main_cat_surv)
blsa_surv_gluc_75_all2 <- cph_funct(quart = "gluc_75_quart", exp_var = c("per_age_wk_75", "sex"), data = main_cat_surv)
blsa_surv_gluc_100_all2 <- cph_funct(quart = "gluc_100_quart", exp_var = c("per_age_wk_100", "sex"), data = main_cat_surv)

blsa_surv_bw_50_all2 <- cph_funct(quart = "bw_50_quart", exp_var = c("per_age_wk_50", "sex"), data = main_cat_surv)
blsa_surv_bw_75_all2 <- cph_funct(quart = "bw_75_quart", exp_var = c("per_age_wk_75", "sex"), data = main_cat_surv)
blsa_surv_bw_100_all2 <- cph_funct(quart = "bw_100_quart", exp_var = c("per_age_wk_100", "sex"), data = main_cat_surv)

blsa_surv_fat_50_all2 <- cph_funct(quart = "fat_50_quart", exp_var = c("per_age_wk_50", "sex"), data = main_cat_surv)
blsa_surv_fat_75_all2 <- cph_funct(quart = "fat_75_quart", exp_var = c("per_age_wk_75", "sex"), data = main_cat_surv)
blsa_surv_fat_100_all2 <- cph_funct(quart = "fat_100_quart", exp_var = c("per_age_wk_100", "sex"), data = main_cat_surv)

## by sex 
blsa_surv_gluc_50_f <- cph_funct(quart = "gluc_50_quart", exp_var = c("per_age_wk_50"), data = main_cat_surv_f) 
blsa_surv_gluc_75_f <- cph_funct(quart = "gluc_75_quart", exp_var = c("per_age_wk_75"), data = main_cat_surv_f) 
blsa_surv_gluc_100_f <- cph_funct(quart = "gluc_100_quart", exp_var = c("per_age_wk_100"), data = main_cat_surv_f) 

blsa_surv_bw_50_f <- cph_funct(quart = "bw_50_quart", exp_var = c("per_age_wk_50"), data = main_cat_surv_f) 
blsa_surv_bw_75_f <- cph_funct(quart = "bw_75_quart", exp_var = c("per_age_wk_75"), data = main_cat_surv_f) 
blsa_surv_bw_100_f <- cph_funct(quart = "bw_100_quart", exp_var = c("per_age_wk_100"), data = main_cat_surv_f) 

blsa_surv_fat_50_f <- cph_funct(quart = "fat_50_quart", exp_var = c("per_age_wk_50"), data = main_cat_surv_f) 
blsa_surv_fat_75_f <- cph_funct(quart = "fat_75_quart", exp_var = c("per_age_wk_75"), data = main_cat_surv_f) 
blsa_surv_fat_100_f <- cph_funct(quart = "fat_100_quart", exp_var = c("per_age_wk_100"), data = main_cat_surv_f) 

blsa_surv_gluc_50_m <- cph_funct(quart = "gluc_50_quart", exp_var = c("per_age_wk_50"), data = main_cat_surv_m)
blsa_surv_gluc_75_m <- cph_funct(quart = "gluc_75_quart", exp_var = c("per_age_wk_75"), data = main_cat_surv_m) 
blsa_surv_gluc_100_m <- cph_funct(quart = "gluc_100_quart", exp_var = c("per_age_wk_100"), data = main_cat_surv_m) 

blsa_surv_bw_50_m <- cph_funct(quart = "bw_50_quart", exp_var = c("per_age_wk_50"), data = main_cat_surv_m)
blsa_surv_bw_75_m <- cph_funct(quart = "bw_75_quart", exp_var = c("per_age_wk_75"), data = main_cat_surv_m) 
blsa_surv_bw_100_m <- cph_funct(quart = "bw_100_quart", exp_var = c("per_age_wk_100"), data = main_cat_surv_m) 

blsa_surv_fat_50_m <- cph_funct(quart = "fat_50_quart", exp_var = c("per_age_wk_50"), data = main_cat_surv_m)
blsa_surv_fat_75_m <- cph_funct(quart = "fat_75_quart", exp_var = c("per_age_wk_75"), data = main_cat_surv_m) 
blsa_surv_fat_100_m <- cph_funct(quart = "fat_100_quart", exp_var = c("per_age_wk_100"), data = main_cat_surv_m) 

## adjusted for body fat
blsa_surv_gluc_50_adj <- cph_funct(quart = "gluc_50_quart", exp_var = c("per_age_wk_50", "sex", "fat_50"), data = main_cat_surv)
blsa_surv_gluc_75_adj <- cph_funct(quart = "gluc_75_quart", exp_var = c("per_age_wk_75", "sex", "fat_75"), data = main_cat_surv)
blsa_surv_gluc_100_adj <- cph_funct(quart = "gluc_100_quart", exp_var = c("per_age_wk_100", "sex", "fat_100"), data = main_cat_surv)

## by sex
blsa_surv_gluc_50_adj_f <- cph_funct(quart = "gluc_50_quart", exp_var = c("per_age_wk_50", "fat_50"), data = main_cat_surv_f)
blsa_surv_gluc_75_adj_f <- cph_funct(quart = "gluc_75_quart", exp_var = c("per_age_wk_75", "fat_75"), data = main_cat_surv_f)
blsa_surv_gluc_100_adj_f <- cph_funct(quart = "gluc_100_quart", exp_var = c("per_age_wk_100", "fat_100"), data = main_cat_surv_f)

blsa_surv_gluc_50_adj_m <- cph_funct(quart = "gluc_50_quart", exp_var = c("per_age_wk_50", "fat_50"), data = main_cat_surv_m)
blsa_surv_gluc_75_adj_m <- cph_funct(quart = "gluc_75_quart", exp_var = c("per_age_wk_75", "fat_75"), data = main_cat_surv_m)
blsa_surv_gluc_100_adj_m <- cph_funct(quart = "gluc_100_quart", exp_var = c("per_age_wk_100", "fat_100"), data = main_cat_surv_m)

```

```{r blsa_km_by, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE, include=TRUE}

hr_toc_by <- NULL 
hr_toc_by$indices <- c("Glucose", "Glucose", "Glucose", "Body Weight", "Body Weight", "Body Weight", "Body Fat", "Body Fat", "Body Fat") 
hr_toc_by$timepoint <- c(50, 75, 100, 50, 75, 100, 50, 75, 100) 

##all
hr_toc_by$hr1_all <- c(blsa_surv_gluc_50_all2[[1]], blsa_surv_gluc_75_all2[[1]], blsa_surv_gluc_100_all2[[1]], 
                       blsa_surv_bw_50_all2[[1]], blsa_surv_bw_75_all2[[1]], blsa_surv_bw_100_all2[[1]], 
                       blsa_surv_fat_50_all2[[1]], blsa_surv_fat_75_all2[[1]], blsa_surv_fat_100_all2[[1]]) 
hr_toc_by$hr2_all <- c(blsa_surv_gluc_50_adj[[1]], blsa_surv_gluc_75_adj[[1]], blsa_surv_gluc_100_adj[[1]], 
                       NA, NA, NA, 
                       NA, NA, NA) 
##female
hr_toc_by$hr1_f <- c(blsa_surv_gluc_50_f[[1]], blsa_surv_gluc_75_f[[1]], blsa_surv_gluc_100_f[[1]], 
                     blsa_surv_bw_50_f[[1]], blsa_surv_bw_75_f[[1]], blsa_surv_bw_100_f[[1]], 
                     blsa_surv_fat_50_f[[1]], blsa_surv_fat_75_f[[1]], blsa_surv_fat_100_f[[1]]) 
hr_toc_by$hr2_f <- c(blsa_surv_gluc_50_adj_f[[1]], blsa_surv_gluc_75_adj_f[[1]], blsa_surv_gluc_100_adj_f[[1]], 
                     NA, NA, NA,
                     NA, NA, NA) 
##male
hr_toc_by$hr1_m <- c(blsa_surv_gluc_50_m[[1]], blsa_surv_gluc_75_m[[1]], blsa_surv_gluc_100_m[[1]], 
                     blsa_surv_bw_50_m[[1]], blsa_surv_bw_75_m[[1]], blsa_surv_bw_100_m[[1]], 
                     blsa_surv_fat_50_m[[1]], blsa_surv_fat_75_m[[1]], blsa_surv_fat_100_m[[1]]) 
hr_toc_by$hr2_m <- c(blsa_surv_gluc_50_adj_m[[1]], blsa_surv_gluc_75_adj_m[[1]], blsa_surv_gluc_100_adj_m[[1]], 
                     NA, NA, NA,
                     NA, NA, NA) 

hr_toc_by <- as.data.frame(hr_toc_by) 

##
hr_toc_by2 <- hr_toc_by %>% 
  mutate_all(~gsub("HR = ", "", .)) 

hr_toc_by3 <- hr_toc_by2 %>% 
  mutate(
    hr1_all = ifelse(grepl("Inf", hr1_all), "---", hr1_all),
    hr2_all = ifelse(grepl("Inf", hr2_all), "---", hr2_all), 
    hr1_m = ifelse(grepl("Inf", hr1_m), "---", hr1_m),
    hr2_m = ifelse(grepl("Inf", hr2_m), "---", hr2_m),
    hr1_f = ifelse(grepl("Inf", hr1_f), "---", hr1_f),
    hr2_f = ifelse(grepl("Inf", hr2_f), "---", hr2_f)
    )

blsa_hr_toc_by_3a <- hr_toc_by3 %>% 
  filter(indices == "Glucose") %>% 
  select(-indices)

blsa_hr_toc_by_3b <- hr_toc_by3 %>% 
  filter(indices != "Glucose") %>% 
  select(-contains("hr2")) 

```

```{r blsa_hr_gluc_kable, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE, include=TRUE}

tablehr1 <- kable(blsa_hr_toc_by_3a, row.names = FALSE, col.names = c("Timepoint$^$" , "Model 1", "Model 2", "Model 1", "Model 2", "Model 1", "Model 2"), escape = FALSE, align = "c", caption = "Mortality Hazard Ratios of Highest and Lowest Glucose Quartiles \n at Varying Timepoints Stratified by Sex") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left", font_size = 12) %>% 
  column_spec(1, bold = TRUE, border_right = T) %>% 
  column_spec(3, border_right = T) %>% 
  column_spec(5, border_right = T) %>% 
  column_spec(c(2:7), width = "30em") %>% 
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  add_header_above(c(" " = 1, "All" = 2, "Female" = 2, "Male" = 2)) %>% 
  footnote(symbol = c("p-value < 0.05", "Hazard ratios (95% confidence intervals) are adjusted for: Model 1) age and sex; Model 2) additionally adjusted for body fat", "Timepoint is percent of mean life expectancy")
           ) 

tablehr1 

```

***

```{r blsa_hr_bodyfat_kable, warning=FALSE, message=FALSE, echo=FALSE, eval=TRUE, include=TRUE}

tablehr2 <- kable(blsa_hr_toc_by_3b, row.names = FALSE, col.names = c("Metabolic Index", "Timepoint$^$" , "All", "Female", "Male"), escape = FALSE, align = "c", caption = "Mortality Hazard Ratios of Highest and Lowest Body Weight and Fat Quartiles \n at Varying Timepoints Stratified by Sex") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left", font_size = 12) %>% 
  column_spec(1, bold = TRUE, border_right = T) %>% 
  column_spec(2, border_right = T) %>% 
  collapse_rows(columns = 1:2, valign = "middle") %>% 
  footnote(symbol = c("p-value < 0.05", "Hazard ratios are adjusted for age and sex", "Timepoint is percent of mean life expectancy")
           ) 

tablehr2

```

## Author Questions 

#### determine assessments over lifespan 

```{r blsa_imp_summary, warning=FALSE, message=FALSE, echo=FALSE}
#create when entry dataset
entry <- main_all3 %>% 
  group_by(idno) %>% 
  summarise( 
    age_entry = min(age_wk, na.rm=T), 
    
    whenentry =  as.factor( 
                ifelse((age_entry >= (life_exp$life_exp_wk*0)) & (age_entry < (life_exp$life_exp_wk*0.375)), 25,
                ifelse((age_entry >= (life_exp$life_exp_wk*0.375)) & (age_entry < (life_exp$life_exp_wk*0.625)), 50,
                ifelse((age_entry >= (life_exp$life_exp_wk*0.625)) & (age_entry < (life_exp$life_exp_wk*0.875)), 75,
                ifelse((age_entry >= (life_exp$life_exp_wk*0.875)) & (age_entry < (life_exp$life_exp_wk*1.125)), 100,
                ifelse((age_entry >= (life_exp$life_exp_wk*1.125)), 125, NA)))))))

asmt_funct_n <- function(time, gluc, gluc2, bw, bw2, fat, fat2){ 
  
  gluc_a <- paste0("!is.na(", gluc, ")") 
  gluc_b <- paste0("is.na(", gluc2, ")") 
  gluc2_a <- paste0("!is.na(", gluc2, ")") 
  
  bw_a <- paste0("!is.na(", bw, ")") 
  bw_b <- paste0("is.na(", bw2, ")") 
  bw2_a <- paste0("!is.na(", bw2, ")") 
  
  fat_a <- paste0("!is.na(", fat, ")") 
  fat_b <- paste0("is.na(", fat2, ")") 
  fat2_a <- paste0("!is.na(", fat2, ")") 
  

main_all3 %>% 
  left_join(entry) %>% 
  left_join(census_c16_death) %>% 
  mutate(whendead = as.numeric(as.character(whendead))) %>% 
  mutate(whenvisit = as.numeric(as.character(whenvisit))) %>% 
  mutate(whenentry = as.numeric(as.character(whenentry))) %>% 
  ungroup() %>% 
  summarise( 

      timepoint = time, 
      new_participants = n_distinct(idno[whenentry == time], na.rm = T), 

      alive = n_distinct(idno[whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      dead = n_distinct(idno[whendead == time], na.rm = T),
      
      gluc_seen = n_distinct(idno[eval(parse(text = gluc_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      gluc_imputed = n_distinct(idno[eval(parse(text = gluc2_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T) - gluc_seen, 
      gluc_missing = n_distinct(idno[eval(parse(text = gluc_b)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      
      bw_seen = n_distinct(idno[eval(parse(text = bw_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      bw_imputed = n_distinct(idno[eval(parse(text = bw2_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T) - bw_seen, 
      bw_missing = n_distinct(idno[eval(parse(text = bw_b)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      
      fat_seen = n_distinct(idno[eval(parse(text = fat_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      fat_imputed = n_distinct(idno[eval(parse(text = fat2_a)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T) - fat_seen, 
      fat_missing = n_distinct(idno[eval(parse(text = fat_b)) & whenentry <= time & ((whendead >= time) | is.na(whendead))], na.rm = T), 
      ) 
} 

asmt_25 <- asmt_funct_n(25, "gluc_25", "gluc_25", "bw_25", "bw_25", "fat_25", "fat_25")
asmt_50 <- asmt_funct_n(50, "gluc_50", "gluc_50", "bw_50", "bw_50", "fat_50", "fat_50")
asmt_75 <- asmt_funct_n(75, "gluc_75", "gluc_75", "bw_75", "bw_75", "fat_75", "fat_75")
asmt_100 <- asmt_funct_n(100, "gluc_100", "gluc_100", "bw_100", "bw_100", "fat_100", "fat_100")
asmt_125 <- asmt_funct_n(125, "gluc_125", "gluc_125", "bw_125", "bw_125", "fat_125", "fat_125")

imputation <- rbind(asmt_25, asmt_50, asmt_75, asmt_100, asmt_125)

```

```{r imputation_kable, warning=FALSE, message=FALSE, echo=FALSE, include=TRUE}

kable(imputation, row.names = FALSE, align = "c", format = "html", escape = F,
      caption = "Assessments of Metabolic indices Over the Lifespan", col.names = c("Timepoint$^$", "New Participants", "Alive", "Dead", "Measured", "Imputed", "Missing", "Measured", "Imputed", "Missing", "Measured", "Imputed", "Missing")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE, position = "left") %>% 
  column_spec(1, bold = T, border_right = T) %>% 
  column_spec(4, border_right = T, width = c("250px"), include_thead =F) %>% 
  column_spec(7, border_right = T) %>% 
  column_spec(10, border_right = T) %>%
  collapse_rows(columns = 1, valign = "middle") %>% 
  add_header_above(c(" " = 1, "Population" = 3, "Glucose" = 3, "Body Weight" = 3, "Body Fat" = 3)) %>%  
  footnote(symbol = c("Imputation using last observed value carried forward", "Timepoint is percent of mean life expectancy"))

```

***

#### trajectories by varying survival groups

```{r blsa_lifespan_dist_fig, cache=TRUE}

annotate_figure(blsa_fig_d13, 
                top = text_grob("Trends in Metabolic Indices Over the Lifespan By Survival Group", size = 12, hjust = 0, x = unit(0, "lines"))
)

```
 
*** 
